### 2023-02-23 outputs/paper/paper3.qmd
Updated Quarto to read in clean data

---
title: "TBD"
title-block-banner: true
abstract: "TBD"
author: "Michaela Drouillard and Christina Wei"
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

```{r}
#| message: false
#| echo: false

#### Workspace set-up ####

# Load packages 
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

# read in cleaned data

cleaned_respondent_info = read_csv(
  file = here("outputs/data/cleaned_respondent_info.csv"),
  show_col_types = FALSE
)

cleaned_finance_survey = read_csv(
  file = here("outputs/data/cleaned_finance_survey.csv"),
  show_col_types = FALSE
)
```

### 2023-03-02 outputs/paper/paper3.qmd
Exploring relationships in data

@@
 abstract: "TBD"
-author: "Michaela Drouillard and Christina Wei"
+thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3104-3-US-Survey.git"
+author: 
+  - Michaela Drouillard
+  - Christina Wei
+date: today
 date-format: long
@@
 )
+
+combined_data = 
+  merge(
+    cleaned_respondent_info,
+    cleaned_finance_survey,
+    by = c("year", "id")
+  )
+
+rm(cleaned_respondent_info)
+rm(cleaned_finance_survey)
 ```
+
+## Introduction
+
+TBD
+
+## Data
+
+Data was cleaned and analyzed using the open source statistically programming language R [@r], using functionalities from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra] and `knitr` [@rKnitr].
+Details of the data extraction and cleaning processes are discussed in the subsections below.
+
+### Surveys Related to Financial Attitudes
+
+### Respondent Demographics
+
+### Measurement Bias
+
+## Findings
+
+```{r}
+#| message: false
+#| echo: false
+
+## Explore 2021 data
+
+survey_data_2021 =
+  combined_data |>
+  filter(year == 2021)
+
+#survey_data_2021 = merge(
+#  filter(cleaned_respondent_info, year == 2021),
+#  filter(cleaned_finance_survey, year == 2021),
+#  by = c("year", "id")
+#)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| 
+survey_data_2021$financial_satisfaction =
+  factor(survey_data_2021$financial_satisfaction,
+         levels = c(
+           "Not Satisfied At All",
+           "More or Less Satisfied",
+           "Pretty Well Satisfied"
+          )
+        )
+
+survey_data_2021$gender =
+  factor(survey_data_2021$gender,
+         levels = c(
+           "Male",
+           "Female"
+          )
+        )
+
+survey_data_2021$degree =
+  factor(survey_data_2021$degree,
+         levels = c(
+           "Less Than High School",
+           "High School",
+           "Associate/Junior College",
+           "Bachelors",
+           "Graduate"
+          )
+        )
+
+survey_data_2021$age_band =
+  factor(survey_data_2021$age_band,
+         levels = c(
+           "18-29",
+           "30-39",
+           "40-49",
+           "50-64",
+           "65-79",
+           ">=80"
+          )
+        )
+
+survey_data_2021$social_class =
+  factor(survey_data_2021$social_class,
+         levels = c(
+           "Lower Class",
+           "Working Class",
+           "Lower Middle Class",
+           "Middle Class",
+           "Upper Middle Class",
+           "Upper Class"
+          )
+        )
+```
+
+### Financial Satisfaction
+
+Overall spread
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  ggplot(aes(x = financial_satisfaction)) +
+  geom_bar(aes(y = (..count..) / sum(..count..))) +
+  scale_y_continuous(labels=scales::percent)
+
+```
+
+Female less satisfied
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(gender), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = gender)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~gender)
+```
+
+Higher degree - more financially satisfied
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = degree)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~degree)
+```
+
+Those born in US are less satisfied?
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(born), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = born)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~born)
+```
+
+Older cohorts are more satisfied with their financial situation
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = age_band)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~age_band)
+
+```
+
+Some differences based on working status
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(work_status), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = work_status)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~work_status)
+```
+
+Sexual orientation - not too much of a pattern?
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(sexual_orientation), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = sexual_orientation)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~sexual_orientation)
+```
+
+Social status - interestingly, upper class is less satisfied than upper middle class?
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = social_class)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~social_class)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+financial_satisfaction_across_years =
+  combined_data |>
+    select (year, financial_satisfaction) |>
+    group_by(year) |>
+    count(financial_satisfaction)
+
+financial_satisfaction_across_years
+```
+
+## Discussion
+
+## Conclusion
+
+## Reference

### 2023-03-02 scripts/01-download_data.R

@@ -11,18 +11,17 @@
 #### Workspace setup ####
 library(tidyverse)
 library(haven) # for reading DTA files
-library(here)
 
 #### Download data ####
 
 # Download & unzip all the GSS data across the years
-zip_file = here("inputs/data/large_files/GSS_stata.zip")
+zip_file = "inputs/data/large_files/GSS_stata.zip"
 
 download.file("https://gss.norc.org/documents/stata/GSS_stata.zip", zip_file)
-unzip(zip_file, exdir = here("inputs/data/large_files"))
+unzip(zip_file, exdir = "inputs/data/large_files")
 
 # read dta data and write to csv
-raw_gss_data = read_dta(here("inputs/data/large_files/gss7221_r3a.dta"))
+raw_gss_data = read_dta("inputs/data/large_files/gss7221_r3a.dta")
 
 
 ## Filter and save data related to respondent info and survey questions of interest
@@ -35,7 +34,7 @@ raw_respondent_info =
 
 write_csv(
   x = raw_respondent_info,
-  file = here("inputs/data/raw_respondent_info.csv")
+  file = "inputs/data/raw_respondent_info.csv"
 )
 
 # zip(
@@ -51,5 +50,5 @@ raw_finance_survey =
 
 write_csv(
   x = raw_finance_survey,
-  file = here("inputs/data/raw_finance_survey.csv")
+  file = "inputs/data/raw_finance_survey.csv"
 )

### 2023-03-02 scripts/02-data_cleaning.R

@@ -13,7 +13,7 @@ library(tidyverse)
 
 # unzip(
 #   "inputs/data/raw_respondent_info.zip",
-#   exdir = here("inputs/data/large_files")
+#   exdir = "inputs/data/large_files"
 # )
 
 raw_respondent_info =
@@ -246,10 +246,10 @@ cleaned_finance_survey =
 
 write_csv(
   x = cleaned_respondent_info,
-  file = here("outputs/data/cleaned_respondent_info.csv")
+  file = "outputs/data/cleaned_respondent_info.csv"
 )
 
 write_csv(
   x = cleaned_finance_survey,
-  file = here("outputs/data/cleaned_finance_survey.csv")
+  file = "outputs/data/cleaned_finance_survey.csv"
 )

### 2023-03-02 outputs/paper/paper3.qmd
updated data section

@@
 rm(cleaned_finance_survey)
+
 ```
@@
 Details of the data extraction and cleaning processes are discussed in the subsections below.
+
+## Data
+
+Since 1972, the General Social Survey (GSS) has been tracking trends in public opinion through in-person interviews.
+Due to the COVID-19 pandemic, the GSS had to change their data collection from in-person to address-based sampling and a push-to-web methodology, meaning that most of the interview was conducted with an online self-administered questionnaire.
+When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey.
+T
+
+#### Gaps
+
+This change in data collection methodology has limitations.
+First, data from the year 2020 is not recorded, which is inconsistent with the yearly GSS track record.
+Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
+Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
+*"To help adjust for differential non-response, the 2021 GSS Cross-section data include weights that are post- stratified to weighted proportions to equal 2019 American Community Survey Census Bureau control totals for region of the country, marital status, race, ethnicity, education, gender, and age.
+We recommend using these weights for all analysis of the 2021 GSS Cross-section data."*
+
+#### Translations
+
+Shifting from in-person interviewing to online data collection required the GSS to find a new approach to appropriately recording responses when respondents expressed uncertainty, indecision, or a refusal to answer.
+Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
+The 2021 GSS had no interviewers in the web mode to appropriately record these responses, so, in order to mitigate this change in mode, no item on the GSS shows "No Answer", unless the module sponsor requested that it be included.
+The exception to this was factual questions about occupation, income, or family background, since respondents could genuinely not know the answer, rather than being indecisive on the answer.
+Instead of "No answer" or "Don't know", users could skip the question.
+"Skipped on web" indicates that users read the questions but skipped it.
+
+#### Affordances: New Experiments
+
+The added element of having a user interface allowed for experiments in survey recording that, according to the GSS' initial review of the data, made a difference in the results.
+For instance, the 2021 GSS displayed questions that shared a theme in a grid together, so that respondents could answer three or four thematically linked questions at the same time.
+The GSS tested abortion items and suicide items in both the traditional and grid format, which items asked in a grid marked with a -G suffix.
+The 2021 GSS also conducted an experiment with volunteered responses.
+In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately.
+In the online format, the GSS included a volunteered response on the screen in some cases, where the variable is marked with a -V suffix, and removed the option entirely for other respondents, marking the variable with an -NV suffix.
+
+#### Impact on our study
+
+Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
+*\[ Here is where we can add how these methodology changes may impact specific variables in our study, when we come across them \]*
 

### 2023-03-02 outputs/paper/paper3.qmd
added graphs for financial compare and financial change

@@
 
+survey_data_2016 =
+  combined_data |>
+  filter(year == 2016)
+
 #survey_data_2021 = merge(
@@
         )
+
+survey_data_2021$financial_compare =
+  factor(survey_data_2021$financial_compare,
+         levels = c(
+           "Far Below Average",
+           "Below Average",
+           "Average",
+           "Above Average",
+           "Far Above Average"
+         ))
+
+survey_data_2021$financial_change =
+  factor(survey_data_2021$financial_change,
+         levels = c(
+           "Getting Worse",
+           "Stayed the Same",
+           "Getting Better"
+         ))
+
+survey_data_2016$financial_change =
+  factor(survey_data_2016$financial_change,
+         levels = c(
+           "Getting Worse",
+           "Stayed the Same",
+           "Getting Better"
+         ))
 ```
@@
 ```{r}
+
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(social_class), !is.na(financial_compare)) |>
+  ggplot(aes(x = financial_compare, group = social_class)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Compare") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~social_class)
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(social_class), !is.na(financial_change)) |>
+  ggplot(aes(x = financial_change, group = social_class)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Change") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~social_class) 
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2016 |>
+  filter(!is.na(social_class), !is.na(financial_change)) |>
+  ggplot(aes(x = financial_change, group = social_class)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Change") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~social_class) 
+```
+
+```{r}
 #| message: false

### 2023-03-05 outputs/paper/paper3.qmd
Added To Do section to keep track of tasks
data exploration for financial satisfaction & social status

@@
 
-#### Workspace set-up ####
+#### Workspace set-up and read in data####
 
@@
 rm(cleaned_finance_survey)
+```
 
+```{r}
+#| message: false
+#| echo: false
+
+#### Add factor defintions to define orders for columns in graphs ####
+
+combined_data$financial_satisfaction =
+  factor(combined_data$financial_satisfaction,
+         levels = c(
+           "Not Satisfied At All",
+           "More or Less Satisfied",
+           "Pretty Well Satisfied"
+          )
+        )
+
+combined_data$gender =
+  factor(combined_data$gender,
+         levels = c(
+           "Male",
+           "Female"
+          )
+        )
+
+combined_data$degree =
+  factor(combined_data$degree,
+         levels = c(
+           "Less Than High School",
+           "High School",
+           "Associate/Junior College",
+           "Bachelors",
+           "Graduate"
+          )
+        )
+
+combined_data$age_band =
+  factor(combined_data$age_band,
+         levels = c(
+           "18-29",
+           "30-39",
+           "40-49",
+           "50-64",
+           "65-79",
+           ">=80"
+          )
+        )
+
+combined_data$social_class =
+  factor(combined_data$social_class,
+         levels = c(
+           "Lower Class",
+           "Working Class",
+           "Middle Class",
+           "Upper Class"
+          )
+        )
+
+combined_data$social_class1 =
+  factor(combined_data$social_class1,
+         levels = c(
+           "Lower Class",
+           "Working Class",
+           "Lower Middle Class",
+           "Middle Class",
+           "Upper Middle Class",
+           "Upper Class"
+          )
+        )
+
+combined_data$cohort_band =
+  factor(combined_data$cohort_band,
+         levels = c(
+           "Greatest",
+           "Silent",
+           "Boomer",
+           "Gen X",
+           "Millenials",
+           "Gen Z"
+           )
+         )
+
+combined_data$financial_compare =
+  factor(combined_data$financial_compare,
+         levels = c(
+           "Far Below Average",
+           "Below Average",
+           "Average",
+           "Above Average",
+           "Far Above Average"
+         ))
+
+combined_data$financial_change =
+  factor(combined_data$financial_change,
+         levels = c(
+           "Getting Worse",
+           "Stayed the Same",
+           "Getting Better"
+         ))
 ```
+
+```{r}
+#| message: false
+#| echo: false
+
+## Save a slice of 2021 and 2016 data
+
+survey_data_2021 =
+  combined_data |>
+  filter(year == 2021)
+
+survey_data_2016 =
+  combined_data |>
+  filter(year == 2016)
+```
+
+## To Do
+
+Christina
+
+-   financial satisfaction
+
+-   Social status
+
+-   calculating % ... over the year for graph
+
+-   Looking into social class getting it over the years - DONE
+
+Michaela
+
+-   financial change
+
+-   financial compare
+
+Other
+
+-   Supplementary survey
+
+-   Simulated data
+
+-   Data validation tests
 
@@
 
-TBD
+The rich getting richer; poor getting poorer
+
+The super upper class are in a bubble and delusional
+
+Female and male differences
 
@@
 
-```{r}
-#| message: false
-#| echo: false
-
-## Explore 2021 data
-
-survey_data_2021 =
-  combined_data |>
-  filter(year == 2021)
-
-survey_data_2016 =
-  combined_data |>
-  filter(year == 2016)
-
-#survey_data_2021 = merge(
-#  filter(cleaned_respondent_info, year == 2021),
-#  filter(cleaned_finance_survey, year == 2021),
-#  by = c("year", "id")
-#)
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| 
-survey_data_2021$financial_satisfaction =
-  factor(survey_data_2021$financial_satisfaction,
-         levels = c(
-           "Not Satisfied At All",
-           "More or Less Satisfied",
-           "Pretty Well Satisfied"
-          )
-        )
-
-survey_data_2021$gender =
-  factor(survey_data_2021$gender,
-         levels = c(
-           "Male",
-           "Female"
-          )
-        )
-
-survey_data_2021$degree =
-  factor(survey_data_2021$degree,
-         levels = c(
-           "Less Than High School",
-           "High School",
-           "Associate/Junior College",
-           "Bachelors",
-           "Graduate"
-          )
-        )
-
-survey_data_2021$age_band =
-  factor(survey_data_2021$age_band,
-         levels = c(
-           "18-29",
-           "30-39",
-           "40-49",
-           "50-64",
-           "65-79",
-           ">=80"
-          )
-        )
-
-survey_data_2021$social_class =
-  factor(survey_data_2021$social_class,
-         levels = c(
-           "Lower Class",
-           "Working Class",
-           "Lower Middle Class",
-           "Middle Class",
-           "Upper Middle Class",
-           "Upper Class"
-          )
-        )
-
-survey_data_2021$financial_compare =
-  factor(survey_data_2021$financial_compare,
-         levels = c(
-           "Far Below Average",
-           "Below Average",
-           "Average",
-           "Above Average",
-           "Far Above Average"
-         ))
-
-survey_data_2021$financial_change =
-  factor(survey_data_2021$financial_change,
-         levels = c(
-           "Getting Worse",
-           "Stayed the Same",
-           "Getting Better"
-         ))
-
-survey_data_2016$financial_change =
-  factor(survey_data_2016$financial_change,
-         levels = c(
-           "Getting Worse",
-           "Stayed the Same",
-           "Getting Better"
-         ))
-```
-
 ### Financial Satisfaction
 
-Overall spread
+Overall spread in 2021
 
@@
 survey_data_2021 |>
+  filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~social_class1)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
   filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
@@
 ```
+
+Financial Satisfaction Over the Years
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+## Create variable to track financial satisfaction across years ##
+
+financial_satisfaction_across_years =
+  combined_data |>
+    select(year, financial_satisfaction) |>
+    group_by(year) |>
+    count(financial_satisfaction)
+```
+
+Financial satisfaction - 2002, 2004, 2006 years have a lot of NAs!
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+financial_satisfaction_across_years |>
+  filter(!is.na(financial_satisfaction)) |>
+  #filter(!(year %in% c(2002, 2004, 2006))) |>
+  mutate(percent = n / sum(n)) |>
+  ggplot(aes(x = year, y = percent, color = financial_satisfaction)) +
+  geom_smooth(method=lm, se=FALSE) + 
+  #geom_bar(stat="identity") +
+  geom_line() +
+  scale_y_continuous(labels=scales::percent)
+
+```
+
+Financial satisfaction - tracking cohort over time
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+financial_satisfaction_across_years_by_cohort |>
+  filter(!is.na(financial_satisfaction)) |>
+  filter(!(is.na(cohort_band))) |>
+  mutate(percent = n / sum(n)) |>
+  #filter(financial_satisfaction == "Not Satisfied At All") |>
+  #mutate(satisfied = 1 - percent) |>
+  ggplot(aes(x = year, y = percent, color = financial_satisfaction )) +
+  geom_line() +
+  geom_smooth(method=lm, se=FALSE) + 
+  scale_y_continuous(labels=scales::percent) +
+  facet_wrap(vars(cohort_band), nrow = 3, ncol = 3) +
+  theme(legend.position="bottom")
+```
+
+Social class over the years
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+## Create variable to track financial satisfaction across years ##
+
+social_class_across_years =
+  combined_data |>
+  select(year, social_class) |>
+  group_by(year) |>
+  count(social_class)
+
+social_class_across_years |>
+  filter(!is.na(social_class)) |>
+  mutate(percent = n / sum(n)) |>
+  ggplot(aes(x = year, y = percent, color = social_class)) +
+  #geom_bar(stat="identity") +
+  geom_line() +
+  scale_y_continuous(labels=scales::percent) +
+  geom_smooth(method=lm, se=FALSE)
+
+```
+
+### Financial Compare
 
@@
 # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2016 |>
-  filter(!is.na(social_class), !is.na(financial_change)) |>
-  ggplot(aes(x = financial_change, group = social_class)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Change") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class) 
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-financial_satisfaction_across_years =
-  combined_data |>
-    select (year, financial_satisfaction) |>
-    group_by(year) |>
-    count(financial_satisfaction)
-
-financial_satisfaction_across_years
+# survey_data_2016 |>
+#   filter(!is.na(social_class), !is.na(financial_change)) |>
+#   ggplot(aes(x = financial_change, group = social_class)) +
+#   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+#   scale_y_continuous(labels=scales::percent) +
+#   ylab("Percentage of response") +
+#   xlab("Financial Change") +
+#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+#   theme(legend.position="none") +
+#   facet_grid(~social_class) 
 ```

### 2023-03-05 scripts/01-download_data.R

@@ -30,7 +30,7 @@ raw_gss_data = read_dta("inputs/data/large_files/gss7221_r3a.dta")
 raw_respondent_info =
   raw_gss_data |>
   #select (year,id,wrkstat,hrs1,hrs2,evwork,occ,prestige,wrkslf,wrkgovt,commute,industry,occ80,prestg80,indus80,indus07,occonet,found,occ10,occindv,occstatus,occtag,prestg10,prestg105plus,indus10,indstatus,indtag,marital,martype,agewed,divorce,widowed,spwrksta,sphrs1,sphrs2,spevwork,cowrksta,cowrkslf,coevwork,cohrs1,cohrs2,spocc,sppres,spwrkslf,spind,spocc80,sppres80,spind80,spocc10,spoccindv,spoccstatus,spocctag,sppres10,sppres105plus,spind10,spindstatus,spindtag,coocc10,coind10,paocc16,papres16,pawrkslf,paind16,paocc80,papres80,paind80,paocc10,paoccindv,paoccstatus,paocctag,papres10,papres105plus,paind10,paindstatus,paindtag,maocc80,mapres80,mawrkslf,maind80,maocc10,maoccindv,maoccstatus,maocctag,mapres10,mapres105plus,maind10,maindstatus,maindtag,sibs,childs,age,agekdbrn,educ,paeduc,maeduc,speduc,coeduc,codeg,degree,padeg,madeg,spdeg,major1,major2,dipged,spdipged,codipged,cosector,whenhs,whencol,sector,eftotlt,barate,gradtounder,voedcol,voednme1,voedncol,voednme2,spsector,speftotlt,spbarate,spgradtounder,cobarate,cogradtounder,coeftotlt,sex,race,res16,reg16,mobile16,family16,famdif16,mawork,mawkbaby,mawkborn,mawk16,mawrkgrw,incom16,born,parborn,granborn,hompop,babies,preteen,teens,adults,unrelat,earnrs,income,rincome,income72,income77,rincom77,income82,rincom82,income86,rincom86,income91,rincom91,income98,rincom98,income06,rincom06,income16,rincom16)
-  select(year, id, wrkstat, marital, cohort, satjob, occ10, age, degree, born, class1, sexornt, sexbirth1, realinc, realrinc) 
+  select(year, id, wrkstat, marital, cohort, satjob, occ10, age, degree, born, class, class1, sexornt, sexbirth1, realinc, realrinc) 
 
 write_csv(
   x = raw_respondent_info,

### 2023-03-05 scripts/02-data_cleaning.R

@@ -69,17 +69,16 @@ cleaned_respondent_info =
   select (-satjob)
 
 # [cohort] Birth cohort of respondent.
+# Defined based on generations https://caregiversofamerica.com/2022-generation-names-explained/
 cleaned_respondent_info =
   cleaned_respondent_info |>
-  mutate("cohort" = case_when(
-    cohort == 1932 ~ '1930-1939',
-    cohort == 1940 ~ '1940-1949',
-    cohort == 1950 ~ '1950-1959',
-    cohort == 1960 ~ '1960-1969',
-    cohort == 1970 ~ '1970-1979',
-    cohort == 1980 ~ '1980-1989',
-    cohort == 1990 ~ '1990-1999',
-    cohort == 2000 ~ '2000-2009',
+  mutate("cohort_band" = case_when(
+    cohort <= 1924 ~ 'Greatest',
+    cohort >= 1925 & cohort <= 1945 ~ 'Silent',
+    cohort >= 1946 & cohort <= 1964 ~ 'Boomer',
+    cohort >= 1965 & cohort <= 1979 ~ 'Gen X',
+    cohort >= 1980 & cohort <= 1994 ~ 'Millenials',
+    cohort >= 1995 & cohort <= 2012 ~ 'Gen Z'
   ))
 
 # [occupation] Respondent's occupation
@@ -118,6 +117,17 @@ cleaned_respondent_info =
 cleaned_respondent_info =
   cleaned_respondent_info |>
   mutate("social_class" = case_when(
+    class == 1 ~ 'Lower Class',
+    class == 2 ~ 'Working Class',
+    class == 3 ~ 'Middle Class',
+    class == 4 ~ 'Upper Class',
+  )) |>
+  select(-class)
+
+# [social_class] Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
+cleaned_respondent_info =
+  cleaned_respondent_info |>
+  mutate("social_class1" = case_when(
     class1 == 1 ~ 'Lower Class',
     class1 == 2 ~ 'Working Class',
     class1 == 3 ~ 'Lower Middle Class',

### 2023-03-05 outputs/paper/paper3.qmd
minor addition to TODO items
Explore relationships with income

@@
 
+combined_data$family_income_bands =
+  factor(combined_data$family_income_bands,
+         levels = c(
+           "<$15K",
+           "$15K-$25K",
+           "$25K-$35K",
+           "$35K-$50K",
+           "$50K-$75K",
+           "$75K-$100K",
+           ">$100K"
+           )
+         )
+
 combined_data$financial_compare =
@@
 
-## To Do
+# To Do
 
@@
 
--   calculating % ... over the year for graph
+-   calculating % ... over the year for graph - DONE
 
@@
 
-## Introduction
+# Introduction
 
@@
 
-## Data
+# Data
 
@@
 
-#### Gaps
+## Gaps
 
@@
 
-#### Translations
+## Translations
 
@@
 
-#### Affordances: New Experiments
+## Affordances: New Experiments
 
@@
 
-#### Impact on our study
+## Impact on our study
 
@@
 
-## Findings
+# Findings
 
-### Financial Satisfaction
+## Financial Satisfaction
 
@@
 
+Financial satisfaction - does money make you more financially satisfied? oh yea.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| 
+survey_data_2021 |>
+  filter(!is.na(family_income_bands), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = family_income_bands)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~family_income_bands)
+```
+
 Financial Satisfaction Over the Years
@@
 
+financial_satisfaction_across_years_by_cohort =
+  combined_data |>
+    select(year, cohort_band, financial_satisfaction) |>
+    group_by(year, cohort_band) |>
+    count(financial_satisfaction)
+
 financial_satisfaction_across_years_by_cohort |>
@@
   mutate(percent = n / sum(n)) |>
-  #filter(financial_satisfaction == "Not Satisfied At All") |>
-  #mutate(satisfied = 1 - percent) |>
   ggplot(aes(x = year, y = percent, color = financial_satisfaction )) +
@@
 ```
+
+## Social Class
 
@@
 
-### Financial Compare
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+survey_data_2021 |>
+  filter(!is.na(family_income_bands), !is.na(social_class)) |>
+  ggplot(aes(x = social_class, group = family_income_bands)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~family_income_bands)
+```
+
+The variable *realinc* has only \~30 values - probably something to do with bands that are in the raw survey data
 
 ```{r}
+#| message: false
+#| echo: false
+#| warning: false
 
+survey_data_2021 |>
+  filter(!is.na(social_class)) |>
+  ggplot(aes(x = social_class, y = realinc)) +
+  geom_boxplot() +
+  geom_jitter(alpha = 0.3, width = 0.10, height = 0)
+  #geom_point(alpha = 0.2) +
+  #geom_violin()
+```
+
+## Financial Compare
+
+```{r}
 #| message: false
@@
 
-## Discussion
+# Discussion
 
-## Conclusion
+# Conclusion
 
-## Reference
+# Reference

### 2023-03-06 outputs/paper/paper3.qmd

@@
 # Findings
+
+## Story
+
+Introduction - news, inflation, millennials and younger not doing better than their parents
+
+\% of financial compare & % financial satisfaction - in 2021
+
+Let's dig deeper into 2021
+
+-   gender
+
+-   education
+
+-   Social class
+
+-   Age
+
+How financial compare / financial satisfaction changed over the years - follow cohort and age
+
+How does social class impact these perceptions? - link to Michaela's literature
 
@@
 survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(financial_compare)) |>
-  ggplot(aes(x = financial_compare, group = social_class)) +
+  filter(!is.na(social_class1), !is.na(financial_compare)) |>
+  ggplot(aes(x = financial_compare, group = social_class1)) +
   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
@@
   theme(legend.position="none") +
-  facet_grid(~social_class)
+  facet_grid(~social_class1)
 

### 2023-03-08 scripts/01-download_data.R
update download and cleaning to include income16 variable; added helper function for graphing

@@ -30,7 +30,7 @@ raw_gss_data = read_dta("inputs/data/large_files/gss7221_r3a.dta")
 raw_respondent_info =
   raw_gss_data |>
   #select (year,id,wrkstat,hrs1,hrs2,evwork,occ,prestige,wrkslf,wrkgovt,commute,industry,occ80,prestg80,indus80,indus07,occonet,found,occ10,occindv,occstatus,occtag,prestg10,prestg105plus,indus10,indstatus,indtag,marital,martype,agewed,divorce,widowed,spwrksta,sphrs1,sphrs2,spevwork,cowrksta,cowrkslf,coevwork,cohrs1,cohrs2,spocc,sppres,spwrkslf,spind,spocc80,sppres80,spind80,spocc10,spoccindv,spoccstatus,spocctag,sppres10,sppres105plus,spind10,spindstatus,spindtag,coocc10,coind10,paocc16,papres16,pawrkslf,paind16,paocc80,papres80,paind80,paocc10,paoccindv,paoccstatus,paocctag,papres10,papres105plus,paind10,paindstatus,paindtag,maocc80,mapres80,mawrkslf,maind80,maocc10,maoccindv,maoccstatus,maocctag,mapres10,mapres105plus,maind10,maindstatus,maindtag,sibs,childs,age,agekdbrn,educ,paeduc,maeduc,speduc,coeduc,codeg,degree,padeg,madeg,spdeg,major1,major2,dipged,spdipged,codipged,cosector,whenhs,whencol,sector,eftotlt,barate,gradtounder,voedcol,voednme1,voedncol,voednme2,spsector,speftotlt,spbarate,spgradtounder,cobarate,cogradtounder,coeftotlt,sex,race,res16,reg16,mobile16,family16,famdif16,mawork,mawkbaby,mawkborn,mawk16,mawrkgrw,incom16,born,parborn,granborn,hompop,babies,preteen,teens,adults,unrelat,earnrs,income,rincome,income72,income77,rincom77,income82,rincom82,income86,rincom86,income91,rincom91,income98,rincom98,income06,rincom06,income16,rincom16)
-  select(year, id, wrkstat, marital, cohort, satjob, occ10, age, degree, born, class, class1, sexornt, sexbirth1, realinc, realrinc) 
+  select(year, id, wrkstat, cohort, satjob, occ10, age, degree, class, class1, sexbirth1, marital, realinc, income16) 
 
 write_csv(
   x = raw_respondent_info,

### 2023-03-08 scripts/02-data_cleaning.R

@@ -105,14 +105,6 @@ cleaned_respondent_info =
     degree == 4 ~ 'Graduate',
   ))
 
-# [born] Were you born in this country?
-cleaned_respondent_info =
-  cleaned_respondent_info |>
-  mutate("born" = case_when(
-    born == 1 ~ 'Yes',
-    born == 2 ~ 'No',
-  ))
-
 # [social_class] Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
 cleaned_respondent_info =
   cleaned_respondent_info |>
@@ -124,7 +116,7 @@ cleaned_respondent_info =
   )) |>
   select(-class)
 
-# [social_class] Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
+# [social_class1] Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
 cleaned_respondent_info =
   cleaned_respondent_info |>
   mutate("social_class1" = case_when(
@@ -137,15 +129,15 @@ cleaned_respondent_info =
   )) |>
   select(-class1)
 
-# [sexual_orientation] 
-cleaned_respondent_info =
-  cleaned_respondent_info |>
-  mutate("sexual_orientation" = case_when(
-    sexornt == 1 ~ 'Homosexual',
-    sexornt == 2 ~ 'Bisexual',
-    sexornt == 3 ~ 'Heterosexual',
-  )) |>
-  select(-sexornt)
+# # [sexual_orientation] 
+# cleaned_respondent_info =
+#   cleaned_respondent_info |>
+#   mutate("sexual_orientation" = case_when(
+#     sexornt == 1 ~ 'Homosexual',
+#     sexornt == 2 ~ 'Bisexual',
+#     sexornt == 3 ~ 'Heterosexual',
+#   )) |>
+#   select(-sexornt)
 
 # [gender] 
 cleaned_respondent_info =
@@ -155,6 +147,25 @@ cleaned_respondent_info =
     sexbirth1 == 2 ~ 'Female',
   )) |>
   select(-sexbirth1)
+
+# [income16]
+cleaned_respondent_info =
+  cleaned_respondent_info |>
+  mutate("income" = case_when(
+    income16 <= 8 ~ '<$10K',
[…]

### 2023-03-08 scripts/04-helper_functions.R

@@ -1,0 +1,23 @@
+#### Preamble ####
+# Purpose: Collection of helper functions
+# Author: Christina Wei
+# Date: 8 March 2023
+# Contact: christina.wei@mail.utoronto.ca
+# License: MIT
+# Prerequisites: none
+
+library(tidyverse)
+
+generate_percentage_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = "Percentage of Responses", angle = NULL, hjust = NULL, vjust = NULL) {
+
+ggplot_data + 
+geom_bar(aes(y = (..count..) / sum(..count..))) +
+scale_y_continuous(labels=scales::percent) +
+theme_minimal() +
+theme(axis.text.x = element_text(angle = angle, hjust = hjust, vjust = vjust)) + #rotate x-axis & smaller font
+labs (
+    x = xlabel,
+    y = ylabel
+  )
+
+}

### 2023-03-08 scripts/02-data_cleaning.R
minor updates to data cleaning routine + cleaned file

@@ -165,7 +165,8 @@ cleaned_respondent_info =
     income16 == 24 ~ '$130K-$150K',
     income16 == 25 ~ '$150K-$170K',
     income16 == 26 ~ '>$170K',
-    ))
+    )) |>
+    select(-income16)
 
 # [family_income_band]
 cleaned_respondent_info =
@@ -178,8 +179,8 @@ cleaned_respondent_info =
     realinc >= 50000 & realinc < 75000 ~ '$50K-$75K',
     realinc >= 75000 & realinc < 100000 ~ '$75K-$100K',
     realinc >= 100000 ~ '>$100K',
-  ))
-
+  )) |>
+  select(-realinc)
 
 # [age_band]
 cleaned_respondent_info = 

### 2023-03-08 outputs/paper/paper3.qmd
Added content for respondent demographics in 2021  survey results

@@
 library(kableExtra)
+library(formattable)
 library(here)
+source(here("scripts/04-helper_functions.R"))
 
@@
 
+combined_data$marital =
+  factor(combined_data$marital,
+         levels = c(
+           "Never Married",
+           "Married",
+           "Widowed",
+           "Divorced",
+           "Separated"
+          )
+        )
+
 combined_data$degree =
@@
            "Upper Class"
+          )
+        )
+
+combined_data$work_status = 
+  factor(combined_data$work_status,
+         levels = c(
+           "Full Time",
+           "Part Time",
+           "In School",
+           "Keeping House",
+           "Unemployed",
+           "Temporarily Not Working",
+           "Retired",
+           "Other"
          )
@@
 
+combined_data$income =
+  factor(combined_data$income,
+         levels = c(
+           "<$10K",
+           "$10K-$20K",
+           "$20K-$30K",
+           "$30K-$40K",
+           "$40K-$50K",
+           "$50K-$60K",
+           "$60K-$75K",
+           "$75K-$90K",
+           "$90K-$110K",
+           "$110K-$130K",
+           "$130K-$150K",
+           "$150K-$17K",
+           ">$170K"
+           )
+         )
+
 combined_data$financial_compare =
@@
 
-survey_data_2016 =
-  combined_data |>
-  filter(year == 2016)
+# survey_data_2016 =
+#   combined_data |>
+#   filter(year == 2016)
 ```
@@
 -   Social status
-
--   calculating % ... over the year for graph - DONE
-
--   Looking into social class getting it over the years - DONE
 
@@
 
-# Findings
+# Results
+
+## Respondent Demographics in 2021 Survey
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-2021-gender
+#| tbl-cap: Number of respondents by gender for 2021 survey
+
+survey_data_2021 |>
+  group_by(gender) |>
+  count() |>
+  ungroup() |>
+  mutate (pct = percent(n / sum(n), digit = 1)) |>
+  kable(
+    col.names = c("Gender", "Number of Responses", "Percentage"),
+    booktabs = TRUE
+  )
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-2021-demographics
+#| fig-cap: "Distribution of Respondents' Demgraphics"
+#| fig-subcap: ["Age","Marital Status", "Level of Education", "Family Income"]
+#| layout-ncol: 2
+#| layout-nrow: 2
+
+#Age
+# survey_data_2021 |>
+#   ggplot(mapping = aes(x = age)) +
+#   geom_histogram(bins = 20, color="black", fill="grey") + 
+#   theme_minimal() +
+#   labs(
+#     x = "Age",
+#     y = "Number of responses"
+#   )
+
+#source(here("scripts/04-helper_functions.R"))
+
+#Age
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$age_band)),
+  xlabel = "Age",
+  angle = 45,
+  hjust = 1
+)
+
+#Marital status
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$marital)),
+  xlabel = "Marital Status",
+  angle = 45,
+  hjust = 1
+)
+
+#Degree
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$degree)),
+  xlabel = "Degree of Education",
+  angle = 45,
+  hjust = 1
+)
+
+#Family Income
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$income)),
+  xlabel = "Family Income in 2016",
+  angle = 45,
+  hjust = 1
+)
+```
+
+In 2021, there are 4032 responses recorded for the US General Social Survey.
+Out of these responses, 1730 (42.9%) of the respondents identified as male, 2198 (54.5%) identified as female, with 104 (2.6%) of them did not identify their **gender** in the survey (@tbl-2021-gender).
+Other than gender, we also looked at four different dimensions of respondents' demographics: age, marital status, degree of education and family income.
+
+The average **age** of respondents who responded in 2021 is 52 years old.
+Looking at the age distribution in @fig-2021-demographics, we see a good spread of values across the different categories, but with noticeably smaller number of respondents less than 29 years old, and those who are over 80 years old.
+It would be interesting to compare this to the demographic data across United States to see if this is a representative sample of the general population.
+There are also a large number of `NA` records for age responses, accounting for 8.3% of our data.
+Based on GSS codebook, 5.6% respondents chose "Not Applicable" when asked about their age, while 2.7% of the participants did not provide any answers.
+
+Most of the respondent provided an answer to their **marital status**.
+About half (49.6%) of the respondents are married as of the time of completing the survey, followed by 2.41% who were never married.
+The rest of the respondents selected either divorced (16.2%), widowed (7.5%), or separated (2.4%).
+
+Respondents' **level of education** was also well answered by respondents, with only 23 `NA` responses (0.6%).
+A significant portion of the respondents reported high school as their highest level of education (39.6%), followed by those who have a junior college or bachelors degree (34.9% combined).
+There are 18.8% of the respondents with a graduate degree, and 6.1% of them reported having less than high school education.
+
+Looking at the respondents' **family income** as of 2016, the first thing that jumps out is the number of `NA` values, accounting for 12.9% of the data.
+Based on GSS codebook, 8.3% respondents chose "Don't Know" for their income, 2.0% refused to answer the question, and 2.7% skipped the question all together.
+Otherwise, there seems to be a good spread of the number of respondents in each of the income bands, with half of the respondents reporting family income less than \$75,000.
+The last income band "\>\$170K" groups all respondents making over \$170,000 into one bucket.
+It reduces the insights we can glean from the data, for example, how much more than \$170,000 are there respondents making per year.
+
+## Overall Financial Satisfaction
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  ggplot(aes(x = financial_satisfaction)) +
+  geom_bar(aes(y = (..count..) / sum(..count..))) +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal()
+
+survey_data_2021 |>
+  ggplot(aes(x = financial_compare)) +
+  geom_bar(aes(y = (..count..) / sum(..count..))) +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal()
+```
 
@@
 
--   Age
+-   age
 
@@
 
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(born), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = born)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~born)
+# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+# survey_data_2021 |>
+#   filter(!is.na(born), !is.na(financial_satisfaction)) |>
+#   ggplot(aes(x = financial_satisfaction, group = born)) +
+#   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+#   scale_y_continuous(labels=scales::percent) +
+#   ylab("Percentage of response") +
+#   xlab("Financial Satisfaction") +
+#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+#   theme(legend.position="none") +
+#   facet_grid(~born)
 ```
@@
 
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
 survey_data_2021 |>
-  filter(!is.na(sexual_orientation), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = sexual_orientation)) +
+  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = marital)) +
   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
@@
   theme(legend.position="none") +
-  facet_grid(~sexual_orientation)
+  facet_grid(~marital)
 ```
@@
 
-survey_data_2021 |>
-  filter(!is.na(social_class)) |>
-  ggplot(aes(x = social_class, y = realinc)) +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.3, width = 0.10, height = 0)
-  #geom_point(alpha = 0.2) +
+# survey_data_2021 |>
+#   filter(!is.na(social_class)) |>
+#   ggplot(aes(x = social_class, y = realinc)) +
+#   geom_boxplot() +
+#   geom_jitter(alpha = 0.3, width = 0.10, height = 0)
+#   #geom_point(alpha = 0.2) +
   #geom_violin()

### 2023-03-08 outputs/paper/paper3.qmd
Added financial attitude section to papper

@@
 
+# Story
+
+Introduction - news, inflation, millennials and younger not doing better than their parents
+
+\% of financial compare & % financial satisfaction - in 2021
+
+Let's dig deeper into 2021
+
+-   gender
+
+-   education
+
+-   Social class
+
+-   age
+
+How financial compare / financial satisfaction changed over the years - follow cohort and age
+
+How does social class impact these perceptions? - link to Michaela's literature
+
 # To Do
@@
 
+-   Simulated data - Christina
+
+-   Data validation tests - Christina
+
 Michaela
@@
 
-Other
-
 -   Supplementary survey
-
--   Simulated data
-
--   Data validation tests
 
@@
 The average **age** of respondents who responded in 2021 is 52 years old.
-Looking at the age distribution in @fig-2021-demographics, we see a good spread of values across the different categories, but with noticeably smaller number of respondents less than 29 years old, and those who are over 80 years old.
+Looking at the age distribution in @fig-2021-demographics-1, we see a good spread of values across the different categories, but with noticeably smaller number of respondents less than 29 years old, and those who are over 80 years old.
 It would be interesting to compare this to the demographic data across United States to see if this is a representative sample of the general population.
@@
 
-Most of the respondent provided an answer to their **marital status**.
+Most of the respondent provided an answer to their **marital status** (@fig-2021-demographics-2).
 About half (49.6%) of the respondents are married as of the time of completing the survey, followed by 2.41% who were never married.
@@
 
-Respondents' **level of education** was also well answered by respondents, with only 23 `NA` responses (0.6%).
+Respondents' **level of education** was also well answered by respondents (@fig-2021-demographics-3), with only 23 `NA` responses (0.6%).
 A significant portion of the respondents reported high school as their highest level of education (39.6%), followed by those who have a junior college or bachelors degree (34.9% combined).
@@
 
-Looking at the respondents' **family income** as of 2016, the first thing that jumps out is the number of `NA` values, accounting for 12.9% of the data.
+Looking at the respondents' **family income** as of 2016 (@fig-2021-demographics-4), the first thing that jumps out is the number of `NA` values, accounting for 12.9% of the data.
 Based on GSS codebook, 8.3% respondents chose "Don't Know" for their income, 2.0% refused to answer the question, and 2.7% skipped the question all together.
@@
 
-## Overall Financial Satisfaction
+## Attitudes Towards Financial Wellness
+
+As estimands to assess societal attitudes towards financial wellness in United States, we looked at the results of following survey questions:
+
+-   **Financial Satisfaction** - We are interested in how people are getting along financially these days.
+So far as you and your family are concerned, would you say that you are pretty well satisfied with your present financial situation, more or less satisfied, or not satisfied at all?
+-   **Comparing to Others** - Compared with American families in general, would you say your family income is far below average, below average, average, above average, or far above average?
+-   **Change in Financial Situation** - During the last few years, has your financial situation been getting better, worse, or has it stayed the same?
+-   **Self-Identified Social Class** - Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
 
@@
 #| warning: false
+#| label: fig-2021-financial
+#| fig-cap: "Distribution of respondents' attitudes towards their financial wellness"
+#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation", "Comparing to Others", "Self-Identified Social Class"]
+#| layout-ncol: 2
+#| layout-nrow: 2
 
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  ggplot(aes(x = financial_satisfaction)) +
-  geom_bar(aes(y = (..count..) / sum(..count..))) +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal()
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_satisfaction)),
+  xlabel = "Financial Satisfaction",
+  angle = 45,
+  hjust = 1
+)
 
-survey_data_2021 |>
-  ggplot(aes(x = financial_compare)) +
-  geom_bar(aes(y = (..count..) / sum(..count..))) +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal()
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_change)),
+  xlabel = "Change in Financial Situation",
+  vjust = 5
+)
+
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_compare)),
+  xlabel = "Comparing Family Income with American Families in General",
+  angle = 45,
+  hjust = 1
+)
+
+generate_percentage_bar_graph(
+  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$social_class)),
+  xlabel = "Self Identified Social Class",
+  vjust = 5
+)
 ```
 
-## Story
+Based on @fig-2021-financial-1, it looks like our society is not doing too badly with their current financial situation.
+Only 23.9% reporting "not satisfied at all" on the question asking whether they are satisfied with their financial situation.
+Similarly, most respondents reported that they are staying the same or getting better in their financial situation, with only 20.1% feeling they are getting worse financially (@fig-2021-financial-2).
 
-Introduction - news, inflation, millennials and younger not doing better than their parents
+To analyze how respondents compare themselves to others, we look at @fig-2021-financial-3 on how they ranked themselves against American families in general.
+This graphs shows that most individuals feel like they are on par with others (39.5%), with small number of respondents feeling above average (24.8%) and below average (24.3%).
+However, 7.1% of respondent felt that they are far below average compared to other families in general, and 3.8% felt they are far above average.
 
-\% of financial compare & % financial satisfaction - in 2021
+Looking at how respondents self identified their social classes (@fig-2021-financial-4), we noticed an interesting pattern that there are similar percentages for respondents who rated themselves as far below average (7.1%) and those who classified themselves as lower class (8.7%).
+Also, the percentages are similar for respondents who rated themselves as far above average (3.8%) and those who classified themselves as upper class (4.2%).
+Most respondents identified themselves ass middle class (49.6%) or working class (37.2%).
 
-Let's dig deeper into 2021
-
--   gender
-
--   education
-
--   Social class
-
--   age
-
-How financial compare / financial satisfaction changed over the years - follow cohort and age
-
-How does social class impact these perceptions? - link to Michaela's literature
-
-## Financial Satisfaction
-
-Overall spread in 2021
+Digging further into the relationship between self identified social class and how respondents compare their financial situation with others, we visualized their correlation using @fig-2021-class-compare.
+There is a clear trend shown in this heatmap: as individuals rank themselves higher in social class, they also perceive their financial situation to be better than the general public.
 
@@
 #| warning: false
+#| label: fig-2021-class-compare
+#| fig-cap: "Correlation between social class and comparing financial situation with others "
 
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
 survey_data_2021 |>
-  ggplot(aes(x = financial_satisfaction)) +
-  geom_bar(aes(y = (..count..) / sum(..count..))) +
-  scale_y_continuous(labels=scales::percent)
+  filter(!is.na(financial_compare), !is.na(social_class)) |>
+  group_by(social_class, financial_compare) |>
+  count() |>
+  ungroup() |>
+  group_by(social_class) |>
+  mutate(Percentage = n / sum(n)) |>
+  ggplot(aes(x = financial_compare, y = social_class, fill = Percentage)) +
+  geom_tile() + 
+  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
+  labs (
+    x = "Comparing Financial Situation to Others",
+    y = "Self-Identified Social Class",
+  )
+```
 
-```
+## Financial Satisfaction
 
@@
 
+# Limitations and Future Direction
+
 # Conclusion

### 2023-03-08 scripts/04-helper_functions.R

@@ -10,6 +10,7 @@ library(tidyverse)
 
 generate_percentage_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = "Percentage of Responses", angle = NULL, hjust = NULL, vjust = NULL) {
 
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
 ggplot_data + 
 geom_bar(aes(y = (..count..) / sum(..count..))) +
 scale_y_continuous(labels=scales::percent) +

### 2023-03-09 outputs/paper/paper3.qmd
added age writeup

@@
 
-## Financial Satisfaction
+## Slicing Financial Attitudes by Demographic Factors
 
-Female less satisfied
+Now that we have a general understanding of the respondent demographics and their attitudes towards financial wellness, let's take a look at the intersections between the attitudes and the demographics.
+
+### Gender
 
@@
 #| warning: false
+#| label: fig-2021-gender-financial
+#| fig-cap: "Intersection of financial attitudes with gender"
+#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation", "Comparing to Others"]
+#| layout-ncol: 3
 
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
 survey_data_2021 |>
   filter(!is.na(gender), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = gender)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  group_by(gender, financial_satisfaction) |>
+  count() |>
+  ungroup() |>
+  group_by(gender) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = financial_satisfaction, y = pct, fill = gender)) +
+  geom_bar(stat="identity", position="dodge") +
   scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~gender)
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Financial Satisfaction",
+    y = "Percentage of Responses",
+    fill = "Gender"
+  )
+
+survey_data_2021 |>
+  filter(!is.na(gender), !is.na(financial_change)) |>
+  group_by(gender, financial_change) |>
+  count() |>
+  ungroup() |>
+  group_by(gender) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = financial_change, y = pct, fill = gender)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Change in Financial Situation",
+    y = "Percentage of Responses",
+    fill = "Gender"
+  )
+
+survey_data_2021 |>
+  filter(!is.na(gender), !is.na(financial_compare)) |>
+  group_by(gender, financial_compare) |>
+  count() |>
+  ungroup() |>
+  group_by(gender) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = financial_compare, y = pct, fill = gender)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Comparing to Others",
+    y = "Percentage of Responses",
+    fill = "Gender"
+  )
 ```
+
+Compare across different genders (defined as sex at birth), we noticed in @fig-2021-gender-financial-1 that there are more female respondents who are not satisfied with their financial situations (26.3%) as compared to male respondents (20.5%).
+Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
+When comparing their financial situations to others, more proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
+These results triangulated together shows that women in our society may be more disadvantaged in their financial wellness as compared to men.
+
+### Age
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-2021-age-financial
+#| fig-cap: "Intersection of financial attitudes with age"
+#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
+#| layout-ncol: 2
+
+survey_data_2021 |>
+  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
+  group_by(age_band, financial_satisfaction) |>
+  count() |>
+  ungroup() |>
+  group_by(age_band) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Financial Satisfaction",
+    y = "Percentage of Responses",
+    fill = "Age Group"
+  )
+
+survey_data_2021 |>
+  filter(!is.na(age_band), !is.na(financial_change)) |>
+  group_by(age_band, financial_change) |>
+  count() |>
+  ungroup() |>
+  group_by(age_band) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = age_band, y = pct, fill = financial_change)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Change in Financial Situation",
+    y = "Percentage of Responses",
+    fill = "Age Group"
+  )
+```
+
+Are we achieving better financial wellness as we get older?
+We found evidence to support this in the 2021 GSS data, as @fig-2021-age-financial-1 demonstrates that the percentage of respondents who selected "pretty well satisfied" over time increases as age increases.
+However, it is interesting to note that the percentage of responses with "not satisfied at all" within an age band increases in the early adulthood, then decreases starting with the 40-49 age band.
+
+Looking at how respondents assessed changes in their financial situation, we see in @fig-2021-age-financial-2 the concerning trend over the age groups that the percentage of those selecting "getting worse" as an answer stays roughly constant at around 20%, showing no reductions over the age bands.
+On the other hand, the responses for "getting better" starts off high at around 50% at younger ages, then drops off starting with the 40-49 age band.
+Staring with the 50-64 age band, most users rate their financial situation to be "staying the same".
+
+### Level of Education
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-2021-degree-financial
+#| fig-cap: "Intersection of financial attitudes with level of education"
+#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
+#| layout-ncol: 2
+
+survey_data_2021 |>
+  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
+  group_by(age_band, financial_satisfaction) |>
+  count() |>
+  ungroup() |>
+  group_by(age_band) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Financial Satisfaction",
+    y = "Percentage of Responses",
+    fill = "Age Group"
+  )
+
+survey_data_2021 |>
+  filter(!is.na(age_band), !is.na(financial_change)) |>
+  group_by(age_band, financial_change) |>
+  count() |>
+  ungroup() |>
+  group_by(age_band) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = age_band, y = pct, fill = financial_change)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Change in Financial Situation",
+    y = "Percentage of Responses",
+    fill = "Age Group"
+  )
+```
+
+### Family Income
 
@@
 survey_data_2021 |>
-  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = degree)) +
+  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = marital)) +
   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
@@
   theme(legend.position="none") +
-  facet_grid(~degree)
+  facet_grid(~marital)
 ```
@@
 
+## Trend Over Time
+
 ## Social Class

### 2023-03-09 outputs/paper/paper3.qmd
Completed first pass at financial wellness x demographic factors; reorganized paper structure

@@
 
-## Data
+## General Social Survey
 
@@
 
-## Gaps
-
-This change in data collection methodology has limitations.
-First, data from the year 2020 is not recorded, which is inconsistent with the yearly GSS track record.
-Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
-Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
-*"To help adjust for differential non-response, the 2021 GSS Cross-section data include weights that are post- stratified to weighted proportions to equal 2019 American Community Survey Census Bureau control totals for region of the country, marital status, race, ethnicity, education, gender, and age.
-We recommend using these weights for all analysis of the 2021 GSS Cross-section data."*
-
-## Translations
-
-Shifting from in-person interviewing to online data collection required the GSS to find a new approach to appropriately recording responses when respondents expressed uncertainty, indecision, or a refusal to answer.
-Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
-The 2021 GSS had no interviewers in the web mode to appropriately record these responses, so, in order to mitigate this change in mode, no item on the GSS shows "No Answer", unless the module sponsor requested that it be included.
-The exception to this was factual questions about occupation, income, or family background, since respondents could genuinely not know the answer, rather than being indecisive on the answer.
-Instead of "No answer" or "Don't know", users could skip the question.
-"Skipped on web" indicates that users read the questions but skipped it.
-
-## Affordances: New Experiments
-
-The added element of having a user interface allowed for experiments in survey recording that, according to the GSS' initial review of the data, made a difference in the results.
-For instance, the 2021 GSS displayed questions that shared a theme in a grid together, so that respondents could answer three or four thematically linked questions at the same time.
-The GSS tested abortion items and suicide items in both the traditional and grid format, which items asked in a grid marked with a -G suffix.
-The 2021 GSS also conducted an experiment with volunteered responses.
-In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately.
-In the online format, the GSS included a volunteered response on the screen in some cases, where the variable is marked with a -V suffix, and removed the option entirely for other respondents, marking the variable with an -NV suffix.
-
-## Impact on our study
-
-Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
-*\[ Here is where we can add how these methodology changes may impact specific variables in our study, when we come across them \]*
-
 ### Surveys Related to Financial Attitudes
@@
 ### Respondent Demographics
-
-### Measurement Bias
 
@@
 survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
-  group_by(age_band, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
+  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
+  group_by(degree, financial_satisfaction) |>
+  count() |>
+  ungroup() |>
+  group_by(degree) |>
   mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)) +
+  ggplot(aes(x = degree, y = pct, fill = financial_satisfaction)) +
   geom_bar(stat="identity", position="dodge") +
@@
 survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_change)) |>
-  group_by(age_band, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
+  filter(!is.na(degree), !is.na(financial_change)) |>
+  group_by(degree, financial_change) |>
+  count() |>
+  ungroup() |>
+  group_by(degree) |>
   mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_change)) +
+  ggplot(aes(x = degree, y = pct, fill = financial_change)) +
   geom_bar(stat="identity", position="dodge") +
@@
 
+The 2021 GSS data shows that having a higher level of education is correlated with higher financial satisfaction.
+in @fig-2021-degree-financial-1 there is a trend of decreasing percentages of answers for "not satisfied at all" as level of education increases, as well as an increasing percentages of responses for "pretty well satisfied" as the level of education increases.
+This finding is supported by the survey results for respondents' perceptions on the changes in the financial situation (@fig-2021-degree-financial-1).
+As the level of education increases, there is a noticeable increase in the percentages of answers for "getting better" in their financial situation.
+Stay in school kids, it will help you financially in the long run.
+
 ### Family Income
 
-Higher degree - more financially satisfied
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = marital)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~marital)
-```
-
-Those born in US are less satisfied?
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-# survey_data_2021 |>
-#   filter(!is.na(born), !is.na(financial_satisfaction)) |>
-#   ggplot(aes(x = financial_satisfaction, group = born)) +
-#   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-#   scale_y_continuous(labels=scales::percent) +
-#   ylab("Percentage of response") +
-#   xlab("Financial Satisfaction") +
-#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-#   theme(legend.position="none") +
-#   facet_grid(~born)
-```
-
-Older cohorts are more satisfied with their financial situation
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = age_band)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~age_band)
-
-```
-
-Some differences based on working status
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(work_status), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = work_status)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~work_status)
-```
-
-Sexual orientation - not too much of a pattern?
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = marital)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~marital)
-```
-
-Social status - interestingly, upper class is less satisfied than upper middle class?
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class1)
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = social_class)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class)
-```
-
-Financial satisfaction - does money make you more financially satisfied? oh yea.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| 
-survey_data_2021 |>
-  filter(!is.na(family_income_bands), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = family_income_bands)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~family_income_bands)
-```
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-2021-income-financial
+#| fig-cap: "Intersection of financial attitudes with family income (2016)"
+#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
+#| layout-ncol: 2
+
+survey_data_2021 |>
+  filter(!is.na(financial_satisfaction), !is.na(income)) |>
+  group_by(financial_satisfaction, income) |>
+  count() |>
+  ungroup() |>
+  group_by(income) |>
+  mutate(Percentage = n / sum(n)) |>
+  ggplot(aes(x = financial_satisfaction, y = income, fill = Percentage)) +
+  geom_tile() + 
+  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
+  labs (
+    x = "Financial Satisfaction",
+    y = "Family Income in 2016",
+  )
+
+survey_data_2021 |>
+  filter(!is.na(financial_change), !is.na(income)) |>
+  group_by(financial_change, income) |>
+  count() |>
+  ungroup() |>
+  group_by(income) |>
+  mutate(Percentage = n / sum(n)) |>
+  ggplot(aes(x = financial_change, y = income, fill = Percentage)) +
+  geom_tile() + 
+  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
+  labs (
+    x = "Comparing Financial Situation to Others",
+    y = "Self-Identified Social Class",
+  )
+
+```
+
+As a hypothesis, we would expect attitudes towards financial wellness is highly correlated with respondents' financial income.
+This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation.
+On the flip side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
+There is also the general trend that as the family income increases, the percentage of being satisfied with their current situation also increases.
+Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).
+
+## Trend Over Time
 
@@
 
-## Trend Over Time
-
 ## Social Class
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~social_class1)
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = social_class)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~social_class)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+survey_data_2021 |>
+  filter(!is.na(social_class), !is.na(income)) |>
+  group_by(social_class, income) |>
+  count() |>
+  ungroup() |>
+  group_by(income) |>
+  mutate(Percentage = n / sum(n)) |>
+  ggplot(aes(x = social_class, y = income, fill = Percentage)) +
+  geom_tile() + 
+  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
+  labs (
+    x = "Comparing Financial Situation to Others",
+    y = "Self-Identified Social Class",
+  )
+```
 
@@
 
+## Marginalization against segments of the population
+
+Rich and richer; poor getting poorer
+
+## Newer generations not doing as well financially
+
+Not trending well over time
+
+##
+
+## Fleishman Effect
+
+Upper class in their own bubble
+
 # Limitations and Future Direction
 
+## Gaps
+
+This change in data collection methodology has limitations.
+First, data from the year 2020 is not recorded, which is inconsistent with the yearly GSS track record.
+Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
+Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
+*"To help adjust for differential non-response, the 2021 GSS Cross-section data include weights that are post- stratified to weighted proportions to equal 2019 American Community Survey Census Bureau control totals for region of the country, marital status, race, ethnicity, education, gender, and age.
+We recommend using these weights for all analysis of the 2021 GSS Cross-section data."*
+
+## Translations
+
+Shifting from in-person interviewing to online data collection required the GSS to find a new approach to appropriately recording responses when respondents expressed uncertainty, indecision, or a refusal to answer.
+Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
+The 2021 GSS had no interviewers in the web mode to appropriately record these responses, so, in order to mitigate this change in mode, no item on the GSS shows "No Answer", unless the module sponsor requested that it be included.
+The exception to this was factual questions about occupation, income, or family background, since respondents could genuinely not know the answer, rather than being indecisive on the answer.
+Instead of "No answer" or "Don't know", users could skip the question.
+"Skipped on web" indicates that users read the questions but skipped it.
+
+## Affordances: New Experiments
+
+The added element of having a user interface allowed for experiments in survey recording that, according to the GSS' initial review of the data, made a difference in the results.
+For instance, the 2021 GSS displayed questions that shared a theme in a grid together, so that respondents could answer three or four thematically linked questions at the same time.
+The GSS tested abortion items and suicide items in both the traditional and grid format, which items asked in a grid marked with a -G suffix.
+The 2021 GSS also conducted an experiment with volunteered responses.
+In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately.
+In the online format, the GSS included a volunteered response on the screen in some cases, where the variable is marked with a -V suffix, and removed the option entirely for other respondents, marking the variable with an -NV suffix.
+
+## Impact on our study
+
+Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
+*\[ Here is where we can add how these methodology changes may impact specific variables in our study, when we come across them \]*
+
+Future studies - incorporate more survey responses
+
+More correlations between factors
+
 # Conclusion

### 2023-03-09 outputs/paper/paper3.qmd
added trend over time findings

@@
 ---
-title: "TBD"
+title: "The Decline of Financial Wellness"
 title-block-banner: true
@@
 
+Compare across different genders (defined as sex at birth), we noticed in @fig-2021-gender-financial-1 that there are more female respondents who are not satisfied with their financial situations (26.3%) as compared to male respondents (20.5%).
+Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
+When comparing their financial situations to others, more proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
+These results triangulated together shows that women in our society may be more disadvantaged in their financial wellness as compared to men.
+
 ```{r}
@@
 
-Compare across different genders (defined as sex at birth), we noticed in @fig-2021-gender-financial-1 that there are more female respondents who are not satisfied with their financial situations (26.3%) as compared to male respondents (20.5%).
-Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
-When comparing their financial situations to others, more proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
-These results triangulated together shows that women in our society may be more disadvantaged in their financial wellness as compared to men.
-
 ### Age
+
+Are we achieving better financial wellness as we get older?
+We found evidence to support this in the 2021 GSS data, as @fig-2021-age-financial-1 demonstrates that the percentage of respondents who selected "pretty well satisfied" over time increases as age increases.
+However, it is interesting to note that the percentage of responses with "not satisfied at all" within an age band increases in the early adulthood, then decreases starting with the 40-49 age band.
+
+Looking at how respondents assessed changes in their financial situation, we see in @fig-2021-age-financial-2 the concerning trend over the age groups that the percentage of those selecting "getting worse" as an answer stays roughly constant at around 20%, showing no reductions over the age bands.
+On the other hand, the responses for "getting better" starts off high at around 50% at younger ages, then drops off starting with the 40-49 age band.
+Staring with the 50-64 age band, most users rate their financial situation to be "staying the same".
 
@@
 
-Are we achieving better financial wellness as we get older?
-We found evidence to support this in the 2021 GSS data, as @fig-2021-age-financial-1 demonstrates that the percentage of respondents who selected "pretty well satisfied" over time increases as age increases.
-However, it is interesting to note that the percentage of responses with "not satisfied at all" within an age band increases in the early adulthood, then decreases starting with the 40-49 age band.
-
-Looking at how respondents assessed changes in their financial situation, we see in @fig-2021-age-financial-2 the concerning trend over the age groups that the percentage of those selecting "getting worse" as an answer stays roughly constant at around 20%, showing no reductions over the age bands.
-On the other hand, the responses for "getting better" starts off high at around 50% at younger ages, then drops off starting with the 40-49 age band.
-Staring with the 50-64 age band, most users rate their financial situation to be "staying the same".
-
 ### Level of Education
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-degree-financial
-#| fig-cap: "Intersection of financial attitudes with level of education"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
-#| layout-ncol: 2
-
-survey_data_2021 |>
-  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
-  group_by(degree, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(degree) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = degree, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(degree), !is.na(financial_change)) |>
-  group_by(degree, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(degree) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = degree, y = pct, fill = financial_change)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-```
 
@@
 
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-2021-degree-financial
+#| fig-cap: "Intersection of financial attitudes with level of education"
+#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
+#| layout-ncol: 2
+
+survey_data_2021 |>
+  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
+  group_by(degree, financial_satisfaction) |>
+  count() |>
+  ungroup() |>
+  group_by(degree) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = degree, y = pct, fill = financial_satisfaction)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Financial Satisfaction",
+    y = "Percentage of Responses",
+    fill = "Age Group"
+  )
+
+survey_data_2021 |>
+  filter(!is.na(degree), !is.na(financial_change)) |>
+  group_by(degree, financial_change) |>
+  count() |>
+  ungroup() |>
+  group_by(degree) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = degree, y = pct, fill = financial_change)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Change in Financial Situation",
+    y = "Percentage of Responses",
+    fill = "Age Group"
+  )
+```
+
 ### Family Income
+
+As a hypothesis, we would expect attitudes towards financial wellness is highly correlated with respondents' financial income.
+This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation.
+On the flip side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
+There is also the general trend that as the family income increases, the percentage of being satisfied with their current situation also increases.
+Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).
 
@@
 
-As a hypothesis, we would expect attitudes towards financial wellness is highly correlated with respondents' financial income.
-This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation.
-On the flip side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
-There is also the general trend that as the family income increases, the percentage of being satisfied with their current situation also increases.
-Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).
-
-## Trend Over Time
-
-Financial Satisfaction Over the Years
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-## Create variable to track financial satisfaction across years ##
-
-financial_satisfaction_across_years =
+## Attitude Towards Financial Wellness Over Time
+
+After investigating different demographic factors that are correlated to respondents' attitudes towards financial wellness, we want to understand how these attitudes evolve over time.
+
+### Financial Satisfaction & Comparison with Others
+
+In @fig-overtime-1, responses for financial satisfaction are graphed between 1972 to 2021.
+Based on this graph, we noticed that the percentage of respondents feeling "more or less satisfied" with their financial situations stays flat over time at around 44%.
+There is a downward trend for the percentage of responses for "pretty well satisfied", as well as an upward trend for the percentage of responses for "not satisfied at all".
+
+Looking at how respondents feel about their financial situation as compared to the general public in @fig-overtime-2, there is a noticeable decline in the percentage of responses for "average", from 57% in 1972 to 40% in 2021.
+The percentage of the other responses all have a similarly slightly positive upward trend.
+This graph shows indications that financial satisfaction has become more divided in society between the rich and the poor.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-overtime
+#| fig-cap: "Attitude towards financial wellness over the years"
+#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
+#| layout-ncol: 2
+
+#Legend wrap line reference: https://datavizpyr.com/fold-legend-into-two-rows-in-ggplot2/
+
 combined_data |>
@@
   group_by(year) |>
-    count(financial_satisfaction)
-```
-
-Financial satisfaction - 2002, 2004, 2006 years have a lot of NAs!
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-financial_satisfaction_across_years |>
+  count(financial_satisfaction) |>
   filter(!is.na(financial_satisfaction)) |>
-  #filter(!(year %in% c(2002, 2004, 2006))) |>
   mutate(percent = n / sum(n)) |>
@@
   geom_smooth(method=lm, se=FALSE) + 
-  #geom_bar(stat="identity") +
   geom_line() +
-  scale_y_continuous(labels=scales::percent)
-
-```
-
-Financial satisfaction - tracking cohort over time
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-financial_satisfaction_across_years_by_cohort =
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() + 
+  theme(legend.position="bottom") + 
+  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
+  labs(
+    x = "Year",
+    y = "Percentage of Responses",
+    color = "Financial Satisfaction"
+  )
+
+combined_data |>
+  select(year, financial_compare) |>
+  group_by(year) |>
+  count(financial_compare) |>
+  filter(!is.na(financial_compare)) |>
+  mutate(percent = n / sum(n)) |>
+  ggplot(aes(x = year, y = percent, color = financial_compare)) +
+  geom_smooth(method=lm, se=FALSE) + 
+  geom_line() +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() + 
+  theme(legend.position="bottom") +
+  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
+  labs(
+    x = "Year",
+    y = "Percentage of Responses",
+    color = "Financial Compare"
+  )
+```
+
+### Cohort
+
+Next we looked at how each generation cohort is doing over time.
+We defined the generations based on the commonly accepted guidelines [@generations].
+Based on @fig-overtime-cohort we found the following trends:
+
+-   **The Greatest Generation (Greatest)**, defined as those born between 1910 and 1924, are between the ages 48 and 64 when the survey started to collect data in 1972.
+They are generally getting more satisfied with their financial situation over time, as shown by increasing "pretty well satisfied" and decreasing "not satisfied at all" response percentages.
+
+-   **The Silent Generation (Silent)**, defined as those born between 1925 and 1945, are between the ages 27 and 47 when the survey started to collect data in 1972.
+They were more dissatisfied with their financial situation to start with, but has increased their financial satisfaction over time, with close to 50% responded with "pretty well satisfied" as of 2021.
+
+-   **Baby Boomer Generation (Boomer),** defined as those born between 1946 and 1964, with some of them as young adults when the survey started to collect data in 1972.
+They were dissatisfied with their financial situation at the beginning, with more than 30% responding that they are "not satisfied at all".
+We have seen improvements over time, with increasing percentages of "pretty well satisified" responses and decreasing percentages of "not satisfied at all" responses.
+
+-   **Generation X (Gen X)**, defined as those born between 1965 and 1979, show up for the first time in survey data in 1983 when the eldest in this generation turned 18.
+This group has a high percentage of respondents dissatisfied with their financial situation (\~30%), while about 25% of them are pertty well satisfied.
+Their attitudes towards financial satisfaction stay pretty much the same throughout the years.
+
+-   **Millennials,** defined as those born between 1980 and 1994, show up for the first time in survey data around year 2000.
+This group starts with a higher satisfaction with their financial situation (\~30%), but it decreases over the years with only \~20% responding with "pretty well satisfied" in 2021.
+
+-   **Generation Z (Gen Z)**, defined as those born between 1995 and 2012, show up for the first time in survey data in year 2013 when the eldest in this generation turned 18.
+This group starts with the highest satisfaction with their financial situation (\~45%), but then it drastically declines to only \~20% responding with "pretty well satisfied" in 2021.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-overtime-cohort
+#| fig-cap: "Financial satisfaction results over the years for each generation"
+
 combined_data |>
@@
   group_by(year, cohort_band) |>
-    count(financial_satisfaction)
-
-financial_satisfaction_across_years_by_cohort |>
-  filter(!is.na(financial_satisfaction)) |>
-  filter(!(is.na(cohort_band))) |>
+  count(financial_satisfaction) |>
+  filter(!is.na(financial_satisfaction), !(is.na(cohort_band))) |>
   mutate(percent = n / sum(n)) |>
@@
   scale_y_continuous(labels=scales::percent) +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 90)) +
   facet_wrap(vars(cohort_band), nrow = 3, ncol = 3) +
-  theme(legend.position="bottom")
+  theme(legend.position="bottom") +
+  theme(legend.text = element_text(size = 5)) +
+  labs(
+    x = "Year",
+    y = "Percentage of Responses",
+    color = "Financial Satisfaction"
+  )
+```
+
+### Age
+
+In addition to looking at cohort over time, we also want to know how respondents at the same age groups are perceiving their financial situation over time.
+Looking at @fig-overtime-age, we see a discouraging picture where "not satisfied at all" with their financial situations is trending upwards for almost all age groups, while being "pretty well satisfied" is trending downwards.
+Out of the bands, 50-64 age group had the highest percentage increase for "not satisfied at all", and most amount of percentage decrease for "pretty well satisfied".
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-overtime-age
+#| fig-cap: "Financial satisfaction results over the years for each age band"
+
+combined_data |>
+  select(year, age_band, financial_satisfaction) |>
+  group_by(year, age_band) |>
+  count(financial_satisfaction) |>
+  filter(!is.na(financial_satisfaction), !(is.na(age_band))) |>
+  mutate(percent = n / sum(n)) |>
+  filter(financial_satisfaction != "More or Less Satisfied") |>
+  ggplot(aes(x = year, y = percent, color = age_band )) +
+  geom_line() +
+  geom_smooth(method=lm, se=FALSE) + 
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() + 
+  #theme(axis.text.x = element_text(angle = 90)) +
+  facet_wrap(vars(financial_satisfaction), ncol = 2) +
+  theme(legend.position="bottom") +
+  #theme(legend.text = element_text(size = 5)) +
+  labs(
+    x = "Year",
+    y = "Percentage of Responses",
+    color = "Age Band"
+  )
 ```
@@
 
-## Create variable to track financial satisfaction across years ##
-
-social_class_across_years =
 combined_data |>
@@
   group_by(year) |>
-  count(social_class)
-
-social_class_across_years |>
+  count(social_class) |>
   filter(!is.na(social_class)) |>
@@
   ggplot(aes(x = year, y = percent, color = social_class)) +
-  #geom_bar(stat="identity") +
+  geom_smooth(method=lm, se=FALSE) + 
   geom_line() +
   scale_y_continuous(labels=scales::percent) +
-  geom_smooth(method=lm, se=FALSE)
+  labs(
+    xlabel = "Year",
+    ylabel = "Percentage of Responses",
+    color = "Social Class"
+  )
 
@@
 
-## Newer generations not doing as well financially
-
-Not trending well over time
+## Getting Worse with Financial Wellness Over Time
+
+Not trending well over time @fig-overtime,
 
@@
 
+Understand whether the demographics in GSS survey is representative of US population --\> gender split, education etc.
+
 # Conclusion

### 2023-03-09 outputs/paper/paper3.qmd
plotted financial compare by income. in the >100k facet virtually no ppl see themselves as far below average. ties back to our convo about cultural capital. also the data could be fudged. also people could be misintepreting what "upper class" means

--- 2023-03-09 outputs/paper/paper3.qmd
@@
 library(kableExtra)
-library(formattable)
 library(here)
-source(here("scripts/04-helper_functions.R"))
 
@@
 
-combined_data$marital =
-  factor(combined_data$marital,
-         levels = c(
-           "Never Married",
-           "Married",
-           "Widowed",
-           "Divorced",
-           "Separated"
-          )
-        )
-
 combined_data$degree =
@@
            "Upper Class"
-          )
-        )
-
-combined_data$work_status = 
-  factor(combined_data$work_status,
-         levels = c(
-           "Full Time",
-           "Part Time",
-           "In School",
-           "Keeping House",
-           "Unemployed",
-           "Temporarily Not Working",
-           "Retired",
-           "Other"
           )
@@
 
-combined_data$income =
-  factor(combined_data$income,
-         levels = c(
-           "<$10K",
-           "$10K-$20K",
-           "$20K-$30K",
-           "$30K-$40K",
-           "$40K-$50K",
-           "$50K-$60K",
-           "$60K-$75K",
-           "$75K-$90K",
-           "$90K-$110K",
-           "$110K-$130K",
-           "$130K-$150K",
-           "$150K-$17K",
-           ">$170K"
-           )
-         )
-
 combined_data$financial_compare =
@@
 
-# survey_data_2016 =
-#   combined_data |>
-#   filter(year == 2016)
+survey_data_2016 =
+  combined_data |>
+  filter(year == 2016)
 ```
-
-# Story
-
-Introduction - news, inflation, millennials and younger not doing better than their parents
-
-\% of financial compare & % financial satisfaction - in 2021
-
-Let's dig deeper into 2021
-
--   gender
-
--   education
-
--   Social class
-
--   age
-
-How financial compare / financial satisfaction changed over the years - follow cohort and age
-
-How does social class impact these perceptions? - link to Michaela's literature
 
@@
 
--   Simulated data - Christina
+-   calculating % ... over the year for graph - DONE
 
--   Data validation tests - Christina
+-   Looking into social class getting it over the years - DONE
 
@@
 
+Other
+
 -   Supplementary survey
+
+-   Simulated data
+
+-   Data validation tests
 
@@
 
-# Results
+# Findings
 
-## Respondent Demographics in 2021 Survey
+## Story
+
+Introduction - news, inflation, millennials and younger not doing better than their parents
+
+\% of financial compare & % financial satisfaction - in 2021
+
+Let's dig deeper into 2021
+
+-   gender
+
+-   education
+
+-   Social class
+
+-   Age
+
+How financial compare / financial satisfaction changed over the years - follow cohort and age
+
+How does social class impact these perceptions? - link to Michaela's literature
+
+## Financial Satisfaction
+
+Overall spread in 2021
 
@@
 #| warning: false
-#| label: tbl-2021-gender
-#| tbl-cap: Number of respondents by gender for 2021 survey
 
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
 survey_data_2021 |>
-  group_by(gender) |>
-  count() |>
-  ungroup() |>
-  mutate (pct = percent(n / sum(n), digit = 1)) |>
-  kable(
-    col.names = c("Gender", "Number of Responses", "Percentage"),
-    booktabs = TRUE
-  )
+  ggplot(aes(x = financial_satisfaction)) +
+  geom_bar(aes(y = (..count..) / sum(..count..))) +
+  scale_y_continuous(labels=scales::percent)
+
 ```
+
+Female less satisfied
 
@@
 #| warning: false
-#| label: fig-2021-demographics
-#| fig-cap: "Distribution of Respondents' Demgraphics"
-#| fig-subcap: ["Age","Marital Status", "Level of Education", "Family Income"]
-#| layout-ncol: 2
-#| layout-nrow: 2
 
-#Age
-# survey_data_2021 |>
-#   ggplot(mapping = aes(x = age)) +
-#   geom_histogram(bins = 20, color="black", fill="grey") + 
-#   theme_minimal() +
-#   labs(
-#     x = "Age",
-#     y = "Number of responses"
-#   )
-
-#source(here("scripts/04-helper_functions.R"))
-
-#Age
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$age_band)),
-  xlabel = "Age",
-  angle = 45,
-  hjust = 1
-)
-
-#Marital status
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$marital)),
-  xlabel = "Marital Status",
-  angle = 45,
-  hjust = 1
-)
-
-#Degree
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$degree)),
-  xlabel = "Degree of Education",
-  angle = 45,
-  hjust = 1
-)
-
-#Family Income
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$income)),
-  xlabel = "Family Income in 2016",
-  angle = 45,
-  hjust = 1
-)
-```
-
-In 2021, there are 4032 responses recorded for the US General Social Survey.
-Out of these responses, 1730 (42.9%) of the respondents identified as male, 2198 (54.5%) identified as female, with 104 (2.6%) of them did not identify their **gender** in the survey (@tbl-2021-gender).
-Other than gender, we also looked at four different dimensions of respondents' demographics: age, marital status, degree of education and family income.
-
-The average **age** of respondents who responded in 2021 is 52 years old.
-Looking at the age distribution in @fig-2021-demographics-1, we see a good spread of values across the different categories, but with noticeably smaller number of respondents less than 29 years old, and those who are over 80 years old.
-It would be interesting to compare this to the demographic data across United States to see if this is a representative sample of the general population.
-There are also a large number of `NA` records for age responses, accounting for 8.3% of our data.
-Based on GSS codebook, 5.6% respondents chose "Not Applicable" when asked about their age, while 2.7% of the participants did not provide any answers.
-
-Most of the respondent provided an answer to their **marital status** (@fig-2021-demographics-2).
-About half (49.6%) of the respondents are married as of the time of completing the survey, followed by 2.41% who were never married.
-The rest of the respondents selected either divorced (16.2%), widowed (7.5%), or separated (2.4%).
-
-Respondents' **level of education** was also well answered by respondents (@fig-2021-demographics-3), with only 23 `NA` responses (0.6%).
-A significant portion of the respondents reported high school as their highest level of education (39.6%), followed by those who have a junior college or bachelors degree (34.9% combined).
-There are 18.8% of the respondents with a graduate degree, and 6.1% of them reported having less than high school education.
-
-Looking at the respondents' **family income** as of 2016 (@fig-2021-demographics-4), the first thing that jumps out is the number of `NA` values, accounting for 12.9% of the data.
-Based on GSS codebook, 8.3% respondents chose "Don't Know" for their income, 2.0% refused to answer the question, and 2.7% skipped the question all together.
-Otherwise, there seems to be a good spread of the number of respondents in each of the income bands, with half of the respondents reporting family income less than \$75,000.
-The last income band "\>\$170K" groups all respondents making over \$170,000 into one bucket.
-It reduces the insights we can glean from the data, for example, how much more than \$170,000 are there respondents making per year.
-
-## Attitudes Towards Financial Wellness
-
-As estimands to assess societal attitudes towards financial wellness in United States, we looked at the results of following survey questions:
-
--   **Financial Satisfaction** - We are interested in how people are getting along financially these days.
-So far as you and your family are concerned, would you say that you are pretty well satisfied with your present financial situation, more or less satisfied, or not satisfied at all?
--   **Comparing to Others** - Compared with American families in general, would you say your family income is far below average, below average, average, above average, or far above average?
--   **Change in Financial Situation** - During the last few years, has your financial situation been getting better, worse, or has it stayed the same?
--   **Self-Identified Social Class** - Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-financial
-#| fig-cap: "Distribution of respondents' attitudes towards their financial wellness"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation", "Comparing to Others", "Self-Identified Social Class"]
-#| layout-ncol: 2
-#| layout-nrow: 2
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_satisfaction)),
-  xlabel = "Financial Satisfaction",
-  angle = 45,
-  hjust = 1
-)
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_change)),
-  xlabel = "Change in Financial Situation",
-  vjust = 5
-)
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_compare)),
-  xlabel = "Comparing Family Income with American Families in General",
-  angle = 45,
-  hjust = 1
-)
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$social_class)),
-  xlabel = "Self Identified Social Class",
-  vjust = 5
-)
-```
-
-Based on @fig-2021-financial-1, it looks like our society is not doing too badly with their current financial situation.
-Only 23.9% reporting "not satisfied at all" on the question asking whether they are satisfied with their financial situation.
-Similarly, most respondents reported that they are staying the same or getting better in their financial situation, with only 20.1% feeling they are getting worse financially (@fig-2021-financial-2).
-
-To analyze how respondents compare themselves to others, we look at @fig-2021-financial-3 on how they ranked themselves against American families in general.
-This graphs shows that most individuals feel like they are on par with others (39.5%), with small number of respondents feeling above average (24.8%) and below average (24.3%).
-However, 7.1% of respondent felt that they are far below average compared to other families in general, and 3.8% felt they are far above average.
-
-Looking at how respondents self identified their social classes (@fig-2021-financial-4), we noticed an interesting pattern that there are similar percentages for respondents who rated themselves as far below average (7.1%) and those who classified themselves as lower class (8.7%).
-Also, the percentages are similar for respondents who rated themselves as far above average (3.8%) and those who classified themselves as upper class (4.2%).
-Most respondents identified themselves ass middle class (49.6%) or working class (37.2%).
-
-Digging further into the relationship between self identified social class and how respondents compare their financial situation with others, we visualized their correlation using @fig-2021-class-compare.
-There is a clear trend shown in this heatmap: as individuals rank themselves higher in social class, they also perceive their financial situation to be better than the general public.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-class-compare
-#| fig-cap: "Correlation between social class and comparing financial situation with others "
-
-survey_data_2021 |>
-  filter(!is.na(financial_compare), !is.na(social_class)) |>
-  group_by(social_class, financial_compare) |>
-  count() |>
-  ungroup() |>
-  group_by(social_class) |>
-  mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = financial_compare, y = social_class, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Comparing Financial Situation to Others",
-    y = "Self-Identified Social Class",
-  )
-```
-
-## Slicing Financial Attitudes by Demographic Factors
-
-Now that we have a general understanding of the respondent demographics and their attitudes towards financial wellness, let's take a look at the intersections between the attitudes and the demographics.
-
-### Gender
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-gender-financial
-#| fig-cap: "Intersection of financial attitudes with gender"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation", "Comparing to Others"]
-#| layout-ncol: 3
-
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
 survey_data_2021 |>
   filter(!is.na(gender), !is.na(financial_satisfaction)) |>
-  group_by(gender, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(gender) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_satisfaction, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
+  ggplot(aes(x = financial_satisfaction, group = gender)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
   scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(gender), !is.na(financial_change)) |>
-  group_by(gender, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(gender) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_change, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(gender), !is.na(financial_compare)) |>
-  group_by(gender, financial_compare) |>
-  count() |>
-  ungroup() |>
-  group_by(gender) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_compare, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Comparing to Others",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~gender)
 ```
-
-Compare across different genders (defined as sex at birth), we noticed in @fig-2021-gender-financial-1 that there are more female respondents who are not satisfied with their financial situations (26.3%) as compared to male respondents (20.5%).
-Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
-When comparing their financial situations to others, more proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
-These results triangulated together shows that women in our society may be more disadvantaged in their financial wellness as compared to men.
-
-### Age
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-age-financial
-#| fig-cap: "Intersection of financial attitudes with age"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
-#| layout-ncol: 2
-
-survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
-  group_by(age_band, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_change)) |>
-  group_by(age_band, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_change)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-```
-
-Are we achieving better financial wellness as we get older?
-We found evidence to support this in the 2021 GSS data, as @fig-2021-age-financial-1 demonstrates that the percentage of respondents who selected "pretty well satisfied" over time increases as age increases.
-However, it is interesting to note that the percentage of responses with "not satisfied at all" within an age band increases in the early adulthood, then decreases starting with the 40-49 age band.
-
-Looking at how respondents assessed changes in their financial situation, we see in @fig-2021-age-financial-2 the concerning trend over the age groups that the percentage of those selecting "getting worse" as an answer stays roughly constant at around 20%, showing no reductions over the age bands.
-On the other hand, the responses for "getting better" starts off high at around 50% at younger ages, then drops off starting with the 40-49 age band.
-Staring with the 50-64 age band, most users rate their financial situation to be "staying the same".
-
-### Level of Education
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-degree-financial
-#| fig-cap: "Intersection of financial attitudes with level of education"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
-#| layout-ncol: 2
-
-survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
-  group_by(age_band, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_change)) |>
-  group_by(age_band, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_change)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-```
-
-### Family Income
 
@@
 survey_data_2021 |>
-  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = marital)) +
+  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = degree)) +
   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
@@
   theme(legend.position="none") +
-  facet_grid(~marital)
+  facet_grid(~degree)
 ```
@@
 
-# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-# survey_data_2021 |>
-#   filter(!is.na(born), !is.na(financial_satisfaction)) |>
-#   ggplot(aes(x = financial_satisfaction, group = born)) +
-#   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-#   scale_y_continuous(labels=scales::percent) +
-#   ylab("Percentage of response") +
-#   xlab("Financial Satisfaction") +
-#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-#   theme(legend.position="none") +
-#   facet_grid(~born)
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(born), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = born)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Satisfaction") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~born)
 ```
@@
 
-# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
 survey_data_2021 |>
-  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = marital)) +
+  filter(!is.na(sexual_orientation), !is.na(financial_satisfaction)) |>
+  ggplot(aes(x = financial_satisfaction, group = sexual_orientation)) +
   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
@@
   theme(legend.position="none") +
-  facet_grid(~marital)
+  facet_grid(~sexual_orientation)
 ```
@@
 
-## Trend Over Time
-
 ## Social Class
@@
 
-# survey_data_2021 |>
-#   filter(!is.na(social_class)) |>
-#   ggplot(aes(x = social_class, y = realinc)) +
-#   geom_boxplot() +
-#   geom_jitter(alpha = 0.3, width = 0.10, height = 0)
-#   #geom_point(alpha = 0.2) +
+survey_data_2021 |>
+  filter(!is.na(social_class)) |>
+  ggplot(aes(x = social_class, y = realinc)) +
+  geom_boxplot() +
+  geom_jitter(alpha = 0.3, width = 0.10, height = 0)
+  #geom_point(alpha = 0.2) +
   #geom_violin()
@@
   facet_grid(~social_class1)
+
+```
+
+```{r}
+
+{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(family_income_bands), !is.na(financial_compare)) |>
+  ggplot(aes(x = financial_compare, group = family_income_bands)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Compare") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~family_income_bands)
 
@@
 
-# Limitations and Future Direction
-
 # Conclusion

### 2023-03-09 outputs/paper/paper3.qmd

--- 2023-03-09 outputs/paper/paper3.qmd
@@
 ```{r}
+
+{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+survey_data_2021 |>
+  filter(!is.na(family_income_bands), !is.na(financial_compare)) |>
+  ggplot(aes(x = financial_compare, group = family_income_bands)) +
+  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  scale_y_continuous(labels=scales::percent) +
+  ylab("Percentage of response") +
+  xlab("Financial Compare") +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  theme(legend.position="none") +
+  facet_grid(~family_income_bands)
+
+```
+
+```{r}
 #| message: false

### 2023-03-09 outputs/paper/paper3.qmd
draft data section complete

@@
 author: 
+  - Christina Wei
   - Michaela Drouillard
-  - Christina Wei
 date: today
@@
 
--   Simulated data - Christina
-
--   Data validation tests - Christina
+-   Simulated data
+
+-   Data validation tests
+
+-   README
 
@@
 
--   financial change
-
--   financial compare
-
--   Supplementary survey
+-   Supplementary survey: The survey should have an introductory section and include the details of a contact person.
+The survey questions should be well constructed and appropriate to the task.
+The questions should have an appropriate ordering.
+A final section should thank the respondent
+
+-   Introduction
+
+-   Discussions
+
+-   Limitations & Future Work
+
+To Do
+
+-   Abstract
+
+-   Conclusion
+
+-   Title
 
@@
 
-The super upper class are in a bubble and delusional
+The upper class are in a bubble and delusional
 
@@
 
-Data was cleaned and analyzed using the open source statistically programming language R [@r], using functionalities from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra] and `knitr` [@rKnitr].
-Details of the data extraction and cleaning processes are discussed in the subsections below.
-
-## General Social Survey
+Data used in this pare are retrieved from the US General Social Survey (GSS) from NORC at the University of Chicago [@gss].
+We retrieved demographic data as well as survey questions related to financial wellness from 1972 to 2021.
+
+## Source Data
+
+Specifically, we retrieved the following demographic data (TODO: explain these values)
+
+-   cohort
+
+-   age
+
+-   degree
+
+-   class
+
+-   class1
+
+-   sexbirth1
+
+-   martial
+
+-   realinc
+
+-   income16
+
+For survey questions related to financial wellness, we retrieved the following data points (TODO: explain these values):
+
+-   confinan
+
+-   satfin
+
+-   finalter
+
+-   finrela
+
+## Data Limitations
 
@@
 When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey.
-T
-
-### Surveys Related to Financial Attitudes
-
-### Respondent Demographics
+
+### Gaps
+
+This change in data collection methodology has limitations.
+First, data from the year 2020 is not recorded, which is inconsistent with the yearly GSS track record.
+Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
+Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
+*"To help adjust for differential non-response, the 2021 GSS Cross-section data include weights that are post- stratified to weighted proportions to equal 2019 American Community Survey Census Bureau control totals for region of the country, marital status, race, ethnicity, education, gender, and age.
+We recommend using these weights for all analysis of the 2021 GSS Cross-section data."*
+
+### Translations
+
+Shifting from in-person interviewing to online data collection required the GSS to find a new approach to appropriately recording responses when respondents expressed uncertainty, indecision, or a refusal to answer.
+Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
+The 2021 GSS had no interviewers in the web mode to appropriately record these responses, so, in order to mitigate this change in mode, no item on the GSS shows "No Answer", unless the module sponsor requested that it be included.
+The exception to this was factual questions about occupation, income, or family background, since respondents could genuinely not know the answer, rather than being indecisive on the answer.
+Instead of "No answer" or "Don't know", users could skip the question.
+"Skipped on web" indicates that users read the questions but skipped it.
+
+### Affordances: New Experiments
+
+The added element of having a user interface allowed for experiments in survey recording that, according to the GSS' initial review of the data, made a difference in the results.
+For instance, the 2021 GSS displayed questions that shared a theme in a grid together, so that respondents could answer three or four thematically linked questions at the same time.
+The GSS tested abortion items and suicide items in both the traditional and grid format, which items asked in a grid marked with a -G suffix.
+The 2021 GSS also conducted an experiment with volunteered responses.
+In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately.
+In the online format, the GSS included a volunteered response on the screen in some cases, where the variable is marked with a -V suffix, and removed the option entirely for other respondents, marking the variable with an -NV suffix.
+
+### Impact on our study
+
+Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
+*\[ Here is where we can add how these methodology changes may impact specific variables in our study, when we come across them \]*
+
+Future studies - incorporate more survey responses
+
+More correlations between factors
+
+Understand whether the demographics in GSS survey is representative of US population --\> gender split, education etc.
+
+## Data Cleaning
+
+Data was cleaned and analyzed using the open source statistically programming language R [@r], using functionalities from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `haven` [@rHaven], `here` [@rHere], `formattable` [@rFormattable], `kableExtra` [@rKableExtra] and `knitr` [@rKnitr].
+
+After downloading and filtering for the desired source data values from GSS[^1], we performed data cleaning on each one of the columns based on value definitions as defined in GSS codebook [@gssCodebook], as well as cleaning up column names.
+For example, the column `sexbirth1` has been renamed to `gender`, with the value 1 updated to "Male", and the value 2 updated to "Female" based on mapping in the codebook.
+
+[^1]: https://gss.norc.org/documents/stata/GSS_stata.zip
+
+Overall, there are 68,846 observations collected across 33 surveys administered over the years.
+On average there are 2086 observations per survey, with the maximum number of 4510 observations in year 2006, and minimum number of 1372 observations in year 1990.
+The standard deviation for the number of observations per survey is 788 (see @tbl-statistics).
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-statistics
+#| tbl-cap: Summary statistics for number of observations in surveys
+combined_data |>
+  group_by(year) |>
+  count() |>
+  ungroup() |>
+  summarise(
+    mean = mean(n),
+    max = max(n),
+    min = min(n),
+    sd = sd(n)
+  ) |>
+  kable(
+    booktabs = TRUE,
+    digits = 1
+  )
+```
 
@@
 
-## Social Class
+## Social Class (still need to fix up)
 
@@
 
-## Financial Compare
+## Financial Compare (old)
 
@@
 
-##
-
 ## Fleishman Effect
@@
 
-## Gaps
-
-This change in data collection methodology has limitations.
-First, data from the year 2020 is not recorded, which is inconsistent with the yearly GSS track record.
-Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
-Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
-*"To help adjust for differential non-response, the 2021 GSS Cross-section data include weights that are post- stratified to weighted proportions to equal 2019 American Community Survey Census Bureau control totals for region of the country, marital status, race, ethnicity, education, gender, and age.
-We recommend using these weights for all analysis of the 2021 GSS Cross-section data."*
-
-## Translations
-
-Shifting from in-person interviewing to online data collection required the GSS to find a new approach to appropriately recording responses when respondents expressed uncertainty, indecision, or a refusal to answer.
-Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
-The 2021 GSS had no interviewers in the web mode to appropriately record these responses, so, in order to mitigate this change in mode, no item on the GSS shows "No Answer", unless the module sponsor requested that it be included.
-The exception to this was factual questions about occupation, income, or family background, since respondents could genuinely not know the answer, rather than being indecisive on the answer.
-Instead of "No answer" or "Don't know", users could skip the question.
-"Skipped on web" indicates that users read the questions but skipped it.
-
-## Affordances: New Experiments
-
-The added element of having a user interface allowed for experiments in survey recording that, according to the GSS' initial review of the data, made a difference in the results.
-For instance, the 2021 GSS displayed questions that shared a theme in a grid together, so that respondents could answer three or four thematically linked questions at the same time.
-The GSS tested abortion items and suicide items in both the traditional and grid format, which items asked in a grid marked with a -G suffix.
-The 2021 GSS also conducted an experiment with volunteered responses.
-In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately.
-In the online format, the GSS included a volunteered response on the screen in some cases, where the variable is marked with a -V suffix, and removed the option entirely for other respondents, marking the variable with an -NV suffix.
-
-## Impact on our study
-
-Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
-*\[ Here is where we can add how these methodology changes may impact specific variables in our study, when we come across them \]*
-
-Future studies - incorporate more survey responses
-
-More correlations between factors
-
-Understand whether the demographics in GSS survey is representative of US population --\> gender split, education etc.
+Additional financial variables in survey for analysis
+
+Connect with US census data to check representativeness
+
+Connect with other macroeconomic indicators - e.g. inflation, stock markets
+
+Collect more data to supplement our findings -\> link to survey <https://docs.google.com/forms/d/e/1FAIpQLSdqe87Bet1PyaCezPxN_t7A1smQCpSap6EYjG2LrTN0K_CX9A/viewform>
 

### 2023-03-09 outputs/paper/paper3.qmd
updated data section - adding variable definitions

@@
 
-Specifically, we retrieved the following demographic data (TODO: explain these values)
-
--   cohort
-
--   age
-
--   degree
-
--   class
-
--   class1
-
--   sexbirth1
-
--   martial
-
--   realinc
-
--   income16
-
-For survey questions related to financial wellness, we retrieved the following data points (TODO: explain these values):
-
--   confinan
-
--   satfin
-
--   finalter
-
--   finrela
+Specifically, we retrieved the following demographic data.
+
+| Variable  | New Name | Description                    | Example              |
+|-----------|----------|--------------------------------|----------------------|
+| age       | age      | Respondent's age               | 56                   |
+| cohort    | cohort   | Birth cohort of respondent     | 1965                 |
+| sexbirth1 | gender   | sex recorded at birth          | Female               |
+| marital   | marital  | Respondent's martial status    | Divorced             |
+| degree    | degree   | Respondent's degree            | Bachelors            |
+| income16  | income   | family income recorded in 2016 | \$50,000 TO \$59,999 |
+
+For survey questions related to financial wellness, we retrieved the following data points:
+
+| Variable | New Name               | Description                                                                             | Example               |
+|---------|-----------------|----------------------------|------------------|
+| satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
+| finalter | financial_change       | Changes in respondent's financial situation                                             | Stayed the same       |
+| finrela  | financial_compare      | Comparing respondent's family income with American families in general                  | Above Average         |
+| class    | social_class           | self identified social class the respondent belong in                                   | Middle Class          |
+| class1   | social_class1          | self identified social class the respondent belong in (more granular than social_class) | Upper Middle Class    |
 
@@
 ```{r}
-
-{r}
 #| message: false

### 2023-03-09 outputs/paper/paper3.qmd
changed subtitles

@@
 | Variable | New Name               | Description                                                                             | Example               |
-|---------|-----------------|----------------------------|------------------|
+|----------------|----------------|-------------------------|----------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 
-### Gaps
+### Gap in 2020 data
 
@@
 
-### Translations
+### "Don't know" responses
 
@@
 
-### Affordances: New Experiments
+### Afordances: New Experiments
 

### 2023-03-09 outputs/paper/paper3.qmd
added discussion or limitation

@@
 
-## Marginalization against segments of the population
+Our results point to two trends: first, evidence to support the hypothesis that the rich are getting richer, and that the poor are getting poorer.
+And second, evidence to support the Fleishman effect \-- the idea that upper classes have a distorted sense of their financial well-being, since they compare themselves to the upper class and opt into wealthy lifestyles, rather than comparing themselves to the general population.
+
+### (Economic) The Rich are getting Richer, and the Poor are getting Poorer
 
@@
 
-## Getting Worse with Financial Wellness Over Time
-
-Not trending well over time @fig-overtime,
-
-## Fleishman Effect
-
-Upper class in their own bubble
-
-# Limitations and Future Direction
+### (Psychological) Fleishman Effect
+
+During World War II, sociologist Samuel Stouffer led the American Soldier study, which aimed to examine the attitudes and motivations of American soldiers during war time.
+The study collected data using both surveys and interviews from over 500,00 soldiers, and covered topics such as combat experiences, political beliefs, and racial attitudes.
+
+One important insight from this study is the contribution it made to our understanding of reference groups.
+The American Soldier study found that soldiers tend to identify more strongly with fellow soldiers than civilians.
+This identification was found to be a significant factor in shaping soldiers\' attitudes in relation to morale and motivation.
+
+In 1957, the sociologist Robert Merton expanded on this idea of reference groups to further develop the concept of relative deprivation, which refers to feelings of discontent that arise when individuals perceive a discrepancy between their own status, and the status of others in their social comparison group.
+Merton argues that social mobility leads to an increase in relative deprivation, since individuals who achieve higher incomes and status become more aware of the gap between their status, and those who are more affluent in their reference group.
+
+Our data on financial satisfaction based on income supports this hypothesis \[figure about financial satisfaction\].
+When faceted by class instead, the question becomes more interesting.
+When asked to compare themselves to the general population, a significant portion of self-identified upper-classes see themselves as \"below average\".
+
+Is the upper class simply delusional?
+Could survey respondents have misinterpreted what \"upper class\" means?
+Or, are there more factors at play when people consider their own financial situations \-- for instance, the child of millionaires who themselves make \$40k a year might see themselves as culturally and socially upper class, yet below average in terms of financial comparison.
+
+In her 2019 memoir We Need To Talk About Money, Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of struggling millennials.
+After learning that a friend had concealed that she, in fact, didn\'t have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that political liberal individuals can experience as  anti-capitalist sentiment have entered the mainstream conversations:
+
+\"Shame as it relates to wealth is somewhat more unexpected because -- despite occasional pockets of resistance and the growing anti-capitalist sentiment that has entered mainstream cultural conversations -- on the whole, we tend to venerate the rich, assigning enormous social currency to traditional markers of wealth: nice cars, massive houses, designer clothes.
+But that\'s the thing about shame -- it\'s an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to \'fit in\'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
+For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited.\"
+
+Concealing wealth out of shame, to Uwagba, is \"actively unhelpful\" for others in the same reference group, and exacerbates the frustration of realizing that achieving personal finance goals is even more out of reach than she initially thought.	In \[\@fig-overtime\], we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z\'s self-reporting does correlate with a pandemic).
+This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions that distort people\'s understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
+Millennials aren\'t experiencing the same financial satisfaction that Baby Boomers did when they were the same age.
+
+Meanwhile, the lure of upward mobility is addictive, as even the upper class reports financial dissatisfaction.
+An article by The Cut titled \"The Fleishman Effect\" explores how upper-class, New York women are responding to the HBO show Fleishman Is In Trouble, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where \"the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present\".
+
+## Limitations and Future Direction
+
+The study would be strengthened with data that could provide a fuller profile of respondent\'s reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
+
+	For instance, geography can have a major influence on personal finance and reference groups.
+A respondent making \$75,000 a year living in Hattiesburg, Mississippi might feel more financially satisfied than a respondent with the same income living in San Francisco, California.  However, this could sacrifice anonymity.
+This question would still have to be broad enough to provide information on their location (i.e.
+by city), rather than something that could de-anonymize respondents when triangulated with other data (i.e.
+postal code).
+
+Another key factor in how people interpret their own wealth is the way that they narrativize their success in comparison to their upbringing.
+In our supplemental survey in Appendix X, we suggest the questions \"Have you moved up a full socio-economic class compared to your upbringing?
+(Y/N/Don\'t Know\", \"Do you have at least one parent who moved up a full socio-economic class?\" (Y/N/Don\'t Know), and \"Is your current career in the same field as your parents?\" (Y/N/Don\'t Know) to better understand respondent\'s family histories.
+Asking \"Are your friends in the same socio-economic class, higher, or lower?
+(Higher/About the same/Lower)\" would also provide more granular detail into how someone perceives their financial status in comparison to the people closest to them.
+The Social Capital Atlas (<https://www.socialcapital.org/?dimension=EconomicConnectednessIndividual&geoLevel=county&selectedId=06037&dim1=EconomicConnectednessIndividual&dim2=CohesivenessSupportRatio&dim3=CivicEngagementVolunteeringRates&bigModalSection=&bigModalChart=scatterplot&showOutliers=false&colorBy=>) conducted a survey pursuing a similar estimand using Facebook data to explore \"economic connectedness\" in friend groups, which refers to the percentage of friends a respondent has who earn higher or lower incomes than them.
+
+	(fig-2021-degree-financial) shows that respondents with post-secondary education experience better changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is \"getting better\".
+Subsetting these reference groups by their area of study, and whether or not they regret their major, could provide more information on how the different returns-on-investment on different degrees change over time, as shown by The Washington Post\'s reporting on the Federal Reserve Survey of Household Economics and Decisionmaking (<https://www.washingtonpost.com/business/2022/09/02/college-major-regrets/>).
+
+We propose adding the questions \"If you went to college or university, what was your major?\" \"Do you regret your major?
+(Y/N/Don\'t know)\" to the survey.
+Adding this question to our study would provide more detailed insight on reference groups, since we could see how people who enrolled in each major compare themselves to others, and, with more information on family history, we could also gather insight on which fields individuals from different social classes pursue in higher education.
+We could also see how people across different income ranges and classes experience regret about their decisions.
+
+\
 

### 2023-03-09 outputs/paper/paper3.qmd
editing to do
added intro, conclusion, abstract, title

@@
 ---
-title: "The Decline of Financial Wellness"
+title: "Subjective Evaluations of Financial Satisfaction and Well-being: A Study of the US General Social Survey"
 title-block-banner: true
-abstract: "TBD"
+abstract: "This paper examines the subjective evaluations of financial satisfaction and well-being, and how they are influenced by various factors.
+Using data from the US General Social Survey, we explore how people understand their financial situations in comparison to others, subsetting by different demographic factors, against the backdrop of an increasing wealth gap between high and low earning individuals.
+The paper discusses the sociological dynamics and cognitive biases that affect these subjective evaluations, and highlights the importance of considering these factors in understanding financial satisfaction and well-being."
 thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3104-3-US-Survey.git"
@@
 
--   Abstract
-
--   Conclusion
-
--   Title
+-   How the data affects us
 
@@
 
-The rich getting richer; poor getting poorer
-
-The upper class are in a bubble and delusional
-
-Female and male differences
+Financial satisfaction and well-being are subjective self-evaluations that can be influenced by a wide variety of different factors.
+However, the COVID-19 pandemic and the ensuing inflation rates have shown that pressures that economic trends bring on affect different demographics disproportionately.
+While the pandemic was a shared global experience, individual experiences of its implications were not.
+
+In this paper, we use data from the US General Social Survey (GSS) from NORC at the University of Chicago to better understand how people understand their own financial situations in comparison to others, against the background an increasing wealth gap between higher and lower earning individuals.
+We subset and visualize the data by a variety of different demographic factors, such as gender, age, level of education, along with views of how different generational cohorts report their financial well-being over time.
+
+Overall, the data supports general hypotheses that over time, the rich are getting richer, the poor are getting poorer, and that Millennials and Gen Z are financial worse off than preceding generations.
+When faceting by social class, the data provides evidence that individuals understandings of their financial well-being and satisfaction are influenced by their reference groups.
+A significant number of respondents who self-reported as upper class still reported dissatisfaction, and described their financial situation as being "below average".
+In our discussion, we explore the different ways in which sociological dynamics and cognitive biases can affect how someone might understand their own financial situation, along with other factors specific to the survey's limitations that shape this statistic.
 
@@
 
+In conclusion, financial satisfaction and well-being are subjective and influenced by various factors.
+This paper used data from the US General Social Survey to examine how people perceive their financial situations in comparison to others, particularly in light of the increasing wealth gap between high and low earning individuals.
+
+The results support the hypothesis that the rich are getting richer, the poor are getting poorer, and that younger generations are financially worse off than preceding cohorts.
+The data also suggests that people's reference groups influence their understanding of their financial well-being and satisfaction.
+Overall, this paper highlights the complexity of financial satisfaction and well-being, and the importance of considering sociological dynamics and cognitive biases in understanding these subjective evaluations.
+
 # Reference

### 2023-03-09 outputs/paper/paper3.qmd
added how data impacts our study

@@
 
-To Do
-
--   How the data affects us
-
 # Introduction
@@
 
-Data used in this pare are retrieved from the US General Social Survey (GSS) from NORC at the University of Chicago [@gss].
+Data used in this paper are retrieved from the US General Social Survey (GSS) from NORC at the University of Chicago [@gss].
 We retrieved demographic data as well as survey questions related to financial wellness from 1972 to 2021.
@@
 Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
-*"To help adjust for differential non-response, the 2021 GSS Cross-section data include weights that are post- stratified to weighted proportions to equal 2019 American Community Survey Census Bureau control totals for region of the country, marital status, race, ethnicity, education, gender, and age.
-We recommend using these weights for all analysis of the 2021 GSS Cross-section data."*
 
@@
 
-### Afordances: New Experiments
+### Survey Experimentation
 
@@
 Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
-*\[ Here is where we can add how these methodology changes may impact specific variables in our study, when we come across them \]*
-
-Future studies - incorporate more survey responses
-
-More correlations between factors
-
-Understand whether the demographics in GSS survey is representative of US population --\> gender split, education etc.
+While the incomes reported are from 2019, the emotions recorded (i.e.
+financial satisfaction, financial comparison), are from 2021, after a year when many people lost their jobs, moved cities, or had other changes in household arrangements and income.
+While this has less of an impact on long-term visualizations of reported satisfaction over time, analyses of reported satisfaction based on income levels using 2021 data will be influenced by this gap.
 

### 2023-03-09 outputs/paper/paper3.qmd
rendered it

@@
 | Variable | New Name               | Description                                                                             | Example               |
-|----------------|----------------|-------------------------|----------------|
+|----------|------------------------|-----------------------------------------------------------------------------------------|-----------------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 The American Soldier study found that soldiers tend to identify more strongly with fellow soldiers than civilians.
-This identification was found to be a significant factor in shaping soldiers\' attitudes in relation to morale and motivation.
+This identification was found to be a significant factor in shaping soldiers' attitudes in relation to morale and motivation.
 
@@
 When faceted by class instead, the question becomes more interesting.
-When asked to compare themselves to the general population, a significant portion of self-identified upper-classes see themselves as \"below average\".
+When asked to compare themselves to the general population, a significant portion of self-identified upper-classes see themselves as "below average".
 
 Is the upper class simply delusional?
-Could survey respondents have misinterpreted what \"upper class\" means?
+Could survey respondents have misinterpreted what "upper class" means?
 Or, are there more factors at play when people consider their own financial situations \-- for instance, the child of millionaires who themselves make \$40k a year might see themselves as culturally and socially upper class, yet below average in terms of financial comparison.
@@
 In her 2019 memoir We Need To Talk About Money, Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of struggling millennials.
-After learning that a friend had concealed that she, in fact, didn\'t have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that political liberal individuals can experience as  anti-capitalist sentiment have entered the mainstream conversations:
-
-\"Shame as it relates to wealth is somewhat more unexpected because -- despite occasional pockets of resistance and the growing anti-capitalist sentiment that has entered mainstream cultural conversations -- on the whole, we tend to venerate the rich, assigning enormous social currency to traditional markers of wealth: nice cars, massive houses, designer clothes.
-But that\'s the thing about shame -- it\'s an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to \'fit in\'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
-For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited.\"
-
-Concealing wealth out of shame, to Uwagba, is \"actively unhelpful\" for others in the same reference group, and exacerbates the frustration of realizing that achieving personal finance goals is even more out of reach than she initially thought.	In \[\@fig-overtime\], we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z\'s self-reporting does correlate with a pandemic).
-This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions that distort people\'s understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
-Millennials aren\'t experiencing the same financial satisfaction that Baby Boomers did when they were the same age.
+After learning that a friend had concealed that she, in fact, didn't have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that political liberal individuals can experience as  anti-capitalist sentiment have entered the mainstream conversations:
+
+"Shame as it relates to wealth is somewhat more unexpected because -- despite occasional pockets of resistance and the growing anti-capitalist sentiment that has entered mainstream cultural conversations -- on the whole, we tend to venerate the rich, assigning enormous social currency to traditional markers of wealth: nice cars, massive houses, designer clothes.
+But that's the thing about shame -- it's an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to 'fit in'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
+For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited."
+
+Concealing wealth out of shame, to Uwagba, is "actively unhelpful" for others in the same reference group, and exacerbates the frustration of realizing that achieving personal finance goals is even more out of reach than she initially thought.
+In \[\@fig-overtime\], we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z's self-reporting does correlate with a pandemic).
+This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions that distort people's understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
+Millennials aren't experiencing the same financial satisfaction that Baby Boomers did when they were the same age.
 
 Meanwhile, the lure of upward mobility is addictive, as even the upper class reports financial dissatisfaction.
-An article by The Cut titled \"The Fleishman Effect\" explores how upper-class, New York women are responding to the HBO show Fleishman Is In Trouble, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where \"the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present\".
+An article by The Cut titled "The Fleishman Effect" explores how upper-class, New York women are responding to the HBO show Fleishman Is In Trouble, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
 
@@
 
-The study would be strengthened with data that could provide a fuller profile of respondent\'s reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
+The study would be strengthened with data that could provide a fuller profile of respondent's reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
 
@@
 Another key factor in how people interpret their own wealth is the way that they narrativize their success in comparison to their upbringing.
-In our supplemental survey in Appendix X, we suggest the questions \"Have you moved up a full socio-economic class compared to your upbringing?
-(Y/N/Don\'t Know\", \"Do you have at least one parent who moved up a full socio-economic class?\" (Y/N/Don\'t Know), and \"Is your current career in the same field as your parents?\" (Y/N/Don\'t Know) to better understand respondent\'s family histories.
-Asking \"Are your friends in the same socio-economic class, higher, or lower?
-(Higher/About the same/Lower)\" would also provide more granular detail into how someone perceives their financial status in comparison to the people closest to them.
-The Social Capital Atlas (<https://www.socialcapital.org/?dimension=EconomicConnectednessIndividual&geoLevel=county&selectedId=06037&dim1=EconomicConnectednessIndividual&dim2=CohesivenessSupportRatio&dim3=CivicEngagementVolunteeringRates&bigModalSection=&bigModalChart=scatterplot&showOutliers=false&colorBy=>) conducted a survey pursuing a similar estimand using Facebook data to explore \"economic connectedness\" in friend groups, which refers to the percentage of friends a respondent has who earn higher or lower incomes than them.
+In our supplemental survey in Appendix X, we suggest the questions "Have you moved up a full socio-economic class compared to your upbringing?
+(Y/N/Don't Know", "Do you have at least one parent who moved up a full socio-economic class?" (Y/N/Don't Know), and "Is your current career in the same field as your parents?" (Y/N/Don't Know) to better understand respondent's family histories.
+Asking "Are your friends in the same socio-economic class, higher, or lower?
+(Higher/About the same/Lower)" would also provide more granular detail into how someone perceives their financial status in comparison to the people closest to them.
+The Social Capital Atlas (<https://www.socialcapital.org/?dimension=EconomicConnectednessIndividual&geoLevel=county&selectedId=06037&dim1=EconomicConnectednessIndividual&dim2=CohesivenessSupportRatio&dim3=CivicEngagementVolunteeringRates&bigModalSection=&bigModalChart=scatterplot&showOutliers=false&colorBy=>) conducted a survey pursuing a similar estimand using Facebook data to explore "economic connectedness" in friend groups, which refers to the percentage of friends a respondent has who earn higher or lower incomes than them.
 
@@
 
-We propose adding the questions \"If you went to college or university, what was your major?\" \"Do you regret your major?
-(Y/N/Don\'t know)\" to the survey.
+We propose adding the questions "If you went to college or university, what was your major?" "Do you regret your major?
+(Y/N/Don't know)" to the survey.
 Adding this question to our study would provide more detailed insight on reference groups, since we could see how people who enrolled in each major compare themselves to others, and, with more information on family history, we could also gather insight on which fields individuals from different social classes pursue in higher education.

### 2023-03-10 outputs/paper/MDCopyOfpaper3.qmd
Fixed up formatting and references

--- 2023-03-09 outputs/paper/paper3.qmd
@@
 ---
-title: "Subjective Evaluations of Financial Satisfaction and Well-being: A Study of the US General Social Survey"
+title: "TBD"
 title-block-banner: true
-abstract: "This paper examines the subjective evaluations of financial satisfaction and well-being, and how they are influenced by various factors.
-Using data from the US General Social Survey, we explore how people understand their financial situations in comparison to others, subsetting by different demographic factors, against the backdrop of an increasing wealth gap between high and low earning individuals.
-The paper discusses the sociological dynamics and cognitive biases that affect these subjective evaluations, and highlights the importance of considering these factors in understanding financial satisfaction and well-being."
+abstract: "TBD"
 thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3104-3-US-Survey.git"
 author: 
+  - Michaela Drouillard
   - Christina Wei
-  - Michaela Drouillard
 date: today
@@
 ---
-
-```{r}
-#| message: false
-#| echo: false
-
-#### Workspace set-up and read in data####
-
-# Load packages 
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-library(formattable)
-library(here)
-source(here("scripts/04-helper_functions.R"))
-
-# read in cleaned data
-
-cleaned_respondent_info = read_csv(
-  file = here("outputs/data/cleaned_respondent_info.csv"),
-  show_col_types = FALSE
-)
-
-cleaned_finance_survey = read_csv(
-  file = here("outputs/data/cleaned_finance_survey.csv"),
-  show_col_types = FALSE
-)
-
-combined_data = 
-  merge(
-    cleaned_respondent_info,
-    cleaned_finance_survey,
-    by = c("year", "id")
-  )
-
-rm(cleaned_respondent_info)
-rm(cleaned_finance_survey)
-```
-
-```{r}
-#| message: false
-#| echo: false
-
-#### Add factor defintions to define orders for columns in graphs ####
-
-combined_data$financial_satisfaction =
-  factor(combined_data$financial_satisfaction,
-         levels = c(
-           "Not Satisfied At All",
-           "More or Less Satisfied",
-           "Pretty Well Satisfied"
-          )
-        )
-
-combined_data$gender =
-  factor(combined_data$gender,
-         levels = c(
-           "Male",
-           "Female"
-          )
-        )
-
-combined_data$marital =
-  factor(combined_data$marital,
-         levels = c(
-           "Never Married",
-           "Married",
-           "Widowed",
-           "Divorced",
-           "Separated"
-          )
-        )
-
-combined_data$degree =
-  factor(combined_data$degree,
-         levels = c(
-           "Less Than High School",
-           "High School",
-           "Associate/Junior College",
-           "Bachelors",
-           "Graduate"
-          )
-        )
-
-combined_data$age_band =
-  factor(combined_data$age_band,
-         levels = c(
-           "18-29",
-           "30-39",
-           "40-49",
-           "50-64",
-           "65-79",
-           ">=80"
-          )
-        )
-
-combined_data$social_class =
-  factor(combined_data$social_class,
-         levels = c(
-           "Lower Class",
-           "Working Class",
-           "Middle Class",
-           "Upper Class"
-          )
-        )
-
-combined_data$work_status = 
-  factor(combined_data$work_status,
-         levels = c(
-           "Full Time",
-           "Part Time",
-           "In School",
-           "Keeping House",
-           "Unemployed",
-           "Temporarily Not Working",
-           "Retired",
-           "Other"
-         )
-        )
-
-combined_data$social_class1 =
-  factor(combined_data$social_class1,
-         levels = c(
-           "Lower Class",
-           "Working Class",
-           "Lower Middle Class",
-           "Middle Class",
-           "Upper Middle Class",
-           "Upper Class"
-          )
-        )
-
-combined_data$cohort_band =
-  factor(combined_data$cohort_band,
-         levels = c(
-           "Greatest",
-           "Silent",
-           "Boomer",
-           "Gen X",
-           "Millenials",
-           "Gen Z"
-           )
-         )
-
-combined_data$family_income_bands =
-  factor(combined_data$family_income_bands,
-         levels = c(
-           "<$15K",
-           "$15K-$25K",
-           "$25K-$35K",
-           "$35K-$50K",
-           "$50K-$75K",
-           "$75K-$100K",
-           ">$100K"
-           )
-         )
-
-combined_data$income =
-  factor(combined_data$income,
-         levels = c(
-           "<$10K",
-           "$10K-$20K",
-           "$20K-$30K",
-           "$30K-$40K",
-           "$40K-$50K",
-           "$50K-$60K",
-           "$60K-$75K",
-           "$75K-$90K",
-           "$90K-$110K",
-           "$110K-$130K",
-           "$130K-$150K",
-           "$150K-$17K",
-           ">$170K"
-           )
-         )
-
-combined_data$financial_compare =
-  factor(combined_data$financial_compare,
-         levels = c(
-           "Far Below Average",
-           "Below Average",
-           "Average",
-           "Above Average",
-           "Far Above Average"
-         ))
-
-combined_data$financial_change =
-  factor(combined_data$financial_change,
-         levels = c(
-           "Getting Worse",
-           "Stayed the Same",
-           "Getting Better"
-         ))
-```
-
-```{r}
-#| message: false
-#| echo: false
-
-## Save a slice of 2021 and 2016 data
-
-survey_data_2021 =
-  combined_data |>
-  filter(year == 2021)
-
-# survey_data_2016 =
-#   combined_data |>
-#   filter(year == 2016)
-```
 
@@
 -   Limitations & Future Work
-
-# Introduction
-
-Financial satisfaction and well-being are subjective self-evaluations that can be influenced by a wide variety of different factors.
-However, the COVID-19 pandemic and the ensuing inflation rates have shown that pressures that economic trends bring on affect different demographics disproportionately.
-While the pandemic was a shared global experience, individual experiences of its implications were not.
-
-In this paper, we use data from the US General Social Survey (GSS) from NORC at the University of Chicago to better understand how people understand their own financial situations in comparison to others, against the background an increasing wealth gap between higher and lower earning individuals.
-We subset and visualize the data by a variety of different demographic factors, such as gender, age, level of education, along with views of how different generational cohorts report their financial well-being over time.
-
-Overall, the data supports general hypotheses that over time, the rich are getting richer, the poor are getting poorer, and that Millennials and Gen Z are financial worse off than preceding generations.
-When faceting by social class, the data provides evidence that individuals understandings of their financial well-being and satisfaction are influenced by their reference groups.
-A significant number of respondents who self-reported as upper class still reported dissatisfaction, and described their financial situation as being "below average".
-In our discussion, we explore the different ways in which sociological dynamics and cognitive biases can affect how someone might understand their own financial situation, along with other factors specific to the survey's limitations that shape this statistic.
-
-# Data
-
-Data used in this paper are retrieved from the US General Social Survey (GSS) from NORC at the University of Chicago [@gss].
-We retrieved demographic data as well as survey questions related to financial wellness from 1972 to 2021.
-
-## Source Data
-
-Specifically, we retrieved the following demographic data.
-
-| Variable  | New Name | Description                    | Example              |
-|-----------|----------|--------------------------------|----------------------|
-| age       | age      | Respondent's age               | 56                   |
-| cohort    | cohort   | Birth cohort of respondent     | 1965                 |
-| sexbirth1 | gender   | sex recorded at birth          | Female               |
-| marital   | marital  | Respondent's martial status    | Divorced             |
-| degree    | degree   | Respondent's degree            | Bachelors            |
-| income16  | income   | family income recorded in 2016 | \$50,000 TO \$59,999 |
-
-For survey questions related to financial wellness, we retrieved the following data points:
-
-| Variable | New Name               | Description                                                                             | Example               |
-|----------------|----------------|-------------------------|----------------|
-| satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
-| finalter | financial_change       | Changes in respondent's financial situation                                             | Stayed the same       |
-| finrela  | financial_compare      | Comparing respondent's family income with American families in general                  | Above Average         |
-| class    | social_class           | self identified social class the respondent belong in                                   | Middle Class          |
-| class1   | social_class1          | self identified social class the respondent belong in (more granular than social_class) | Upper Middle Class    |
-
-## Data Limitations
-
-Since 1972, the General Social Survey (GSS) has been tracking trends in public opinion through in-person interviews.
-Due to the COVID-19 pandemic, the GSS had to change their data collection from in-person to address-based sampling and a push-to-web methodology, meaning that most of the interview was conducted with an online self-administered questionnaire.
-When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey.
-
-### Gap in 2020 data
-
-This change in data collection methodology has limitations.
-First, data from the year 2020 is not recorded, which is inconsistent with the yearly GSS track record.
-Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
-Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
-
-### "Don't know" responses
-
-Shifting from in-person interviewing to online data collection required the GSS to find a new approach to appropriately recording responses when respondents expressed uncertainty, indecision, or a refusal to answer.
-Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
-The 2021 GSS had no interviewers in the web mode to appropriately record these responses, so, in order to mitigate this change in mode, no item on the GSS shows "No Answer", unless the module sponsor requested that it be included.
-The exception to this was factual questions about occupation, income, or family background, since respondents could genuinely not know the answer, rather than being indecisive on the answer.
-Instead of "No answer" or "Don't know", users could skip the question.
-"Skipped on web" indicates that users read the questions but skipped it.
-
-### Survey Experimentation
-
-The added element of having a user interface allowed for experiments in survey recording that, according to the GSS' initial review of the data, made a difference in the results.
-For instance, the 2021 GSS displayed questions that shared a theme in a grid together, so that respondents could answer three or four thematically linked questions at the same time.
-The GSS tested abortion items and suicide items in both the traditional and grid format, which items asked in a grid marked with a -G suffix.
-The 2021 GSS also conducted an experiment with volunteered responses.
-In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately.
-In the online format, the GSS included a volunteered response on the screen in some cases, where the variable is marked with a -V suffix, and removed the option entirely for other respondents, marking the variable with an -NV suffix.
-
-### Impact on our study
-
-Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
-While the incomes reported are from 2019, the emotions recorded (i.e.
-financial satisfaction, financial comparison), are from 2021, after a year when many people lost their jobs, moved cities, or had other changes in household arrangements and income.
-While this has less of an impact on long-term visualizations of reported satisfaction over time, analyses of reported satisfaction based on income levels using 2021 data will be influenced by this gap.
-
-## Data Cleaning
-
-Data was cleaned and analyzed using the open source statistically programming language R [@r], using functionalities from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `haven` [@rHaven], `here` [@rHere], `formattable` [@rFormattable], `kableExtra` [@rKableExtra] and `knitr` [@rKnitr].
-
-After downloading and filtering for the desired source data values from GSS[^1], we performed data cleaning on each one of the columns based on value definitions as defined in GSS codebook [@gssCodebook], as well as cleaning up column names.
-For example, the column `sexbirth1` has been renamed to `gender`, with the value 1 updated to "Male", and the value 2 updated to "Female" based on mapping in the codebook.
-
-[^1]: https://gss.norc.org/documents/stata/GSS_stata.zip
-
-Overall, there are 68,846 observations collected across 33 surveys administered over the years.
-On average there are 2086 observations per survey, with the maximum number of 4510 observations in year 2006, and minimum number of 1372 observations in year 1990.
-The standard deviation for the number of observations per survey is 788 (see @tbl-statistics).
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: tbl-statistics
-#| tbl-cap: Summary statistics for number of observations in surveys
-combined_data |>
-  group_by(year) |>
-  count() |>
-  ungroup() |>
-  summarise(
-    mean = mean(n),
-    max = max(n),
-    min = min(n),
-    sd = sd(n)
-  ) |>
-  kable(
-    booktabs = TRUE,
-    digits = 1
-  )
-```
-
-# Results
-
-## Respondent Demographics in 2021 Survey
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: tbl-2021-gender
-#| tbl-cap: Number of respondents by gender for 2021 survey
-
-survey_data_2021 |>
-  group_by(gender) |>
-  count() |>
-  ungroup() |>
-  mutate (pct = percent(n / sum(n), digit = 1)) |>
-  kable(
-    col.names = c("Gender", "Number of Responses", "Percentage"),
-    booktabs = TRUE
-  )
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-demographics
-#| fig-cap: "Distribution of Respondents' Demgraphics"
-#| fig-subcap: ["Age","Marital Status", "Level of Education", "Family Income"]
-#| layout-ncol: 2
-#| layout-nrow: 2
-
-#Age
-# survey_data_2021 |>
-#   ggplot(mapping = aes(x = age)) +
-#   geom_histogram(bins = 20, color="black", fill="grey") + 
-#   theme_minimal() +
-#   labs(
-#     x = "Age",
-#     y = "Number of responses"
-#   )
-
-#source(here("scripts/04-helper_functions.R"))
-
-#Age
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$age_band)),
-  xlabel = "Age",
-  angle = 45,
-  hjust = 1
-)
-
-#Marital status
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$marital)),
-  xlabel = "Marital Status",
-  angle = 45,
-  hjust = 1
-)
-
-#Degree
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$degree)),
-  xlabel = "Degree of Education",
-  angle = 45,
-  hjust = 1
-)
-
-#Family Income
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$income)),
-  xlabel = "Family Income in 2016",
-  angle = 45,
-  hjust = 1
-)
-```
-
-In 2021, there are 4032 responses recorded for the US General Social Survey.
-Out of these responses, 1730 (42.9%) of the respondents identified as male, 2198 (54.5%) identified as female, with 104 (2.6%) of them did not identify their **gender** in the survey (@tbl-2021-gender).
-Other than gender, we also looked at four different dimensions of respondents' demographics: age, marital status, degree of education and family income.
-
-The average **age** of respondents who responded in 2021 is 52 years old.
-Looking at the age distribution in @fig-2021-demographics-1, we see a good spread of values across the different categories, but with noticeably smaller number of respondents less than 29 years old, and those who are over 80 years old.
-It would be interesting to compare this to the demographic data across United States to see if this is a representative sample of the general population.
-There are also a large number of `NA` records for age responses, accounting for 8.3% of our data.
-Based on GSS codebook, 5.6% respondents chose "Not Applicable" when asked about their age, while 2.7% of the participants did not provide any answers.
-
-Most of the respondent provided an answer to their **marital status** (@fig-2021-demographics-2).
-About half (49.6%) of the respondents are married as of the time of completing the survey, followed by 2.41% who were never married.
-The rest of the respondents selected either divorced (16.2%), widowed (7.5%), or separated (2.4%).
-
-Respondents' **level of education** was also well answered by respondents (@fig-2021-demographics-3), with only 23 `NA` responses (0.6%).
-A significant portion of the respondents reported high school as their highest level of education (39.6%), followed by those who have a junior college or bachelors degree (34.9% combined).
-There are 18.8% of the respondents with a graduate degree, and 6.1% of them reported having less than high school education.
-
-Looking at the respondents' **family income** as of 2016 (@fig-2021-demographics-4), the first thing that jumps out is the number of `NA` values, accounting for 12.9% of the data.
-Based on GSS codebook, 8.3% respondents chose "Don't Know" for their income, 2.0% refused to answer the question, and 2.7% skipped the question all together.
-Otherwise, there seems to be a good spread of the number of respondents in each of the income bands, with half of the respondents reporting family income less than \$75,000.
-The last income band "\>\$170K" groups all respondents making over \$170,000 into one bucket.
-It reduces the insights we can glean from the data, for example, how much more than \$170,000 are there respondents making per year.
-
-## Attitudes Towards Financial Wellness
-
-As estimands to assess societal attitudes towards financial wellness in United States, we looked at the results of following survey questions:
-
--   **Financial Satisfaction** - We are interested in how people are getting along financially these days.
-So far as you and your family are concerned, would you say that you are pretty well satisfied with your present financial situation, more or less satisfied, or not satisfied at all?
--   **Comparing to Others** - Compared with American families in general, would you say your family income is far below average, below average, average, above average, or far above average?
--   **Change in Financial Situation** - During the last few years, has your financial situation been getting better, worse, or has it stayed the same?
--   **Self-Identified Social Class** - Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-financial
-#| fig-cap: "Distribution of respondents' attitudes towards their financial wellness"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation", "Comparing to Others", "Self-Identified Social Class"]
-#| layout-ncol: 2
-#| layout-nrow: 2
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_satisfaction)),
-  xlabel = "Financial Satisfaction",
-  angle = 45,
-  hjust = 1
-)
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_change)),
-  xlabel = "Change in Financial Situation",
-  vjust = 5
-)
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$financial_compare)),
-  xlabel = "Comparing Family Income with American Families in General",
-  angle = 45,
-  hjust = 1
-)
-
-generate_percentage_bar_graph(
-  ggplot_data = ggplot(mapping = aes(x = survey_data_2021$social_class)),
-  xlabel = "Self Identified Social Class",
-  vjust = 5
-)
-```
-
-Based on @fig-2021-financial-1, it looks like our society is not doing too badly with their current financial situation.
-Only 23.9% reporting "not satisfied at all" on the question asking whether they are satisfied with their financial situation.
-Similarly, most respondents reported that they are staying the same or getting better in their financial situation, with only 20.1% feeling they are getting worse financially (@fig-2021-financial-2).
-
-To analyze how respondents compare themselves to others, we look at @fig-2021-financial-3 on how they ranked themselves against American families in general.
-This graphs shows that most individuals feel like they are on par with others (39.5%), with small number of respondents feeling above average (24.8%) and below average (24.3%).
-However, 7.1% of respondent felt that they are far below average compared to other families in general, and 3.8% felt they are far above average.
-
-Looking at how respondents self identified their social classes (@fig-2021-financial-4), we noticed an interesting pattern that there are similar percentages for respondents who rated themselves as far below average (7.1%) and those who classified themselves as lower class (8.7%).
-Also, the percentages are similar for respondents who rated themselves as far above average (3.8%) and those who classified themselves as upper class (4.2%).
-Most respondents identified themselves ass middle class (49.6%) or working class (37.2%).
-
-Digging further into the relationship between self identified social class and how respondents compare their financial situation with others, we visualized their correlation using @fig-2021-class-compare.
-There is a clear trend shown in this heatmap: as individuals rank themselves higher in social class, they also perceive their financial situation to be better than the general public.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-class-compare
-#| fig-cap: "Correlation between social class and comparing financial situation with others "
-
-survey_data_2021 |>
-  filter(!is.na(financial_compare), !is.na(social_class)) |>
-  group_by(social_class, financial_compare) |>
-  count() |>
-  ungroup() |>
-  group_by(social_class) |>
-  mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = financial_compare, y = social_class, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Comparing Financial Situation to Others",
-    y = "Self-Identified Social Class",
-  )
-```
-
-## Slicing Financial Attitudes by Demographic Factors
-
-Now that we have a general understanding of the respondent demographics and their attitudes towards financial wellness, let's take a look at the intersections between the attitudes and the demographics.
-
-### Gender
-
-Compare across different genders (defined as sex at birth), we noticed in @fig-2021-gender-financial-1 that there are more female respondents who are not satisfied with their financial situations (26.3%) as compared to male respondents (20.5%).
-Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
-When comparing their financial situations to others, more proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
-These results triangulated together shows that women in our society may be more disadvantaged in their financial wellness as compared to men.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-gender-financial
-#| fig-cap: "Intersection of financial attitudes with gender"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation", "Comparing to Others"]
-#| layout-ncol: 3
-
-survey_data_2021 |>
-  filter(!is.na(gender), !is.na(financial_satisfaction)) |>
-  group_by(gender, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(gender) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_satisfaction, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(gender), !is.na(financial_change)) |>
-  group_by(gender, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(gender) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_change, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(gender), !is.na(financial_compare)) |>
-  group_by(gender, financial_compare) |>
-  count() |>
-  ungroup() |>
-  group_by(gender) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_compare, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Comparing to Others",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
-```
-
-### Age
-
-Are we achieving better financial wellness as we get older?
-We found evidence to support this in the 2021 GSS data, as @fig-2021-age-financial-1 demonstrates that the percentage of respondents who selected "pretty well satisfied" over time increases as age increases.
-However, it is interesting to note that the percentage of responses with "not satisfied at all" within an age band increases in the early adulthood, then decreases starting with the 40-49 age band.
-
-Looking at how respondents assessed changes in their financial situation, we see in @fig-2021-age-financial-2 the concerning trend over the age groups that the percentage of those selecting "getting worse" as an answer stays roughly constant at around 20%, showing no reductions over the age bands.
-On the other hand, the responses for "getting better" starts off high at around 50% at younger ages, then drops off starting with the 40-49 age band.
-Staring with the 50-64 age band, most users rate their financial situation to be "staying the same".
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-age-financial
-#| fig-cap: "Intersection of financial attitudes with age"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
-#| layout-ncol: 2
-
-survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
-  group_by(age_band, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(age_band), !is.na(financial_change)) |>
-  group_by(age_band, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(age_band) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_change)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-```
-
-### Level of Education
-
-The 2021 GSS data shows that having a higher level of education is correlated with higher financial satisfaction.
-in @fig-2021-degree-financial-1 there is a trend of decreasing percentages of answers for "not satisfied at all" as level of education increases, as well as an increasing percentages of responses for "pretty well satisfied" as the level of education increases.
-This finding is supported by the survey results for respondents' perceptions on the changes in the financial situation (@fig-2021-degree-financial-1).
-As the level of education increases, there is a noticeable increase in the percentages of answers for "getting better" in their financial situation.
-Stay in school kids, it will help you financially in the long run.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-degree-financial
-#| fig-cap: "Intersection of financial attitudes with level of education"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
-#| layout-ncol: 2
-
-survey_data_2021 |>
-  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
-  group_by(degree, financial_satisfaction) |>
-  count() |>
-  ungroup() |>
-  group_by(degree) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = degree, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-
-survey_data_2021 |>
-  filter(!is.na(degree), !is.na(financial_change)) |>
-  group_by(degree, financial_change) |>
-  count() |>
-  ungroup() |>
-  group_by(degree) |>
-  mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = degree, y = pct, fill = financial_change)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-```
-
-### Family Income
-
-As a hypothesis, we would expect attitudes towards financial wellness is highly correlated with respondents' financial income.
-This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation.
-On the flip side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
-There is also the general trend that as the family income increases, the percentage of being satisfied with their current situation also increases.
-Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-2021-income-financial
-#| fig-cap: "Intersection of financial attitudes with family income (2016)"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
-#| layout-ncol: 2
-
-survey_data_2021 |>
-  filter(!is.na(financial_satisfaction), !is.na(income)) |>
-  group_by(financial_satisfaction, income) |>
-  count() |>
-  ungroup() |>
-  group_by(income) |>
-  mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = financial_satisfaction, y = income, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Financial Satisfaction",
-    y = "Family Income in 2016",
-  )
-
-survey_data_2021 |>
-  filter(!is.na(financial_change), !is.na(income)) |>
-  group_by(financial_change, income) |>
-  count() |>
-  ungroup() |>
-  group_by(income) |>
-  mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = financial_change, y = income, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Comparing Financial Situation to Others",
-    y = "Self-Identified Social Class",
-  )
-
-```
-
-## Attitude Towards Financial Wellness Over Time
-
-After investigating different demographic factors that are correlated to respondents' attitudes towards financial wellness, we want to understand how these attitudes evolve over time.
-
-### Financial Satisfaction & Comparison with Others
-
-In @fig-overtime-1, responses for financial satisfaction are graphed between 1972 to 2021.
-Based on this graph, we noticed that the percentage of respondents feeling "more or less satisfied" with their financial situations stays flat over time at around 44%.
-There is a downward trend for the percentage of responses for "pretty well satisfied", as well as an upward trend for the percentage of responses for "not satisfied at all".
-
-Looking at how respondents feel about their financial situation as compared to the general public in @fig-overtime-2, there is a noticeable decline in the percentage of responses for "average", from 57% in 1972 to 40% in 2021.
-The percentage of the other responses all have a similarly slightly positive upward trend.
-This graph shows indications that financial satisfaction has become more divided in society between the rich and the poor.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-overtime
-#| fig-cap: "Attitude towards financial wellness over the years"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
-#| layout-ncol: 2
-
-#Legend wrap line reference: https://datavizpyr.com/fold-legend-into-two-rows-in-ggplot2/
-
-combined_data |>
-  select(year, financial_satisfaction) |>
-  group_by(year) |>
-  count(financial_satisfaction) |>
-  filter(!is.na(financial_satisfaction)) |>
-  mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = financial_satisfaction)) +
-  geom_smooth(method=lm, se=FALSE) + 
-  geom_line() +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() + 
-  theme(legend.position="bottom") + 
-  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
-  labs(
-    x = "Year",
-    y = "Percentage of Responses",
-    color = "Financial Satisfaction"
-  )
-
-combined_data |>
-  select(year, financial_compare) |>
-  group_by(year) |>
-  count(financial_compare) |>
-  filter(!is.na(financial_compare)) |>
-  mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = financial_compare)) +
-  geom_smooth(method=lm, se=FALSE) + 
-  geom_line() +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() + 
-  theme(legend.position="bottom") +
-  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
-  labs(
-    x = "Year",
-    y = "Percentage of Responses",
-    color = "Financial Compare"
-  )
-```
-
-### Cohort
-
-Next we looked at how each generation cohort is doing over time.
-We defined the generations based on the commonly accepted guidelines [@generations].
-Based on @fig-overtime-cohort we found the following trends:
-
--   **The Greatest Generation (Greatest)**, defined as those born between 1910 and 1924, are between the ages 48 and 64 when the survey started to collect data in 1972.
-They are generally getting more satisfied with their financial situation over time, as shown by increasing "pretty well satisfied" and decreasing "not satisfied at all" response percentages.
-
--   **The Silent Generation (Silent)**, defined as those born between 1925 and 1945, are between the ages 27 and 47 when the survey started to collect data in 1972.
-They were more dissatisfied with their financial situation to start with, but has increased their financial satisfaction over time, with close to 50% responded with "pretty well satisfied" as of 2021.
-
--   **Baby Boomer Generation (Boomer),** defined as those born between 1946 and 1964, with some of them as young adults when the survey started to collect data in 1972.
-They were dissatisfied with their financial situation at the beginning, with more than 30% responding that they are "not satisfied at all".
-We have seen improvements over time, with increasing percentages of "pretty well satisified" responses and decreasing percentages of "not satisfied at all" responses.
-
--   **Generation X (Gen X)**, defined as those born between 1965 and 1979, show up for the first time in survey data in 1983 when the eldest in this generation turned 18.
-This group has a high percentage of respondents dissatisfied with their financial situation (\~30%), while about 25% of them are pertty well satisfied.
-Their attitudes towards financial satisfaction stay pretty much the same throughout the years.
-
--   **Millennials,** defined as those born between 1980 and 1994, show up for the first time in survey data around year 2000.
-This group starts with a higher satisfaction with their financial situation (\~30%), but it decreases over the years with only \~20% responding with "pretty well satisfied" in 2021.
-
--   **Generation Z (Gen Z)**, defined as those born between 1995 and 2012, show up for the first time in survey data in year 2013 when the eldest in this generation turned 18.
-This group starts with the highest satisfaction with their financial situation (\~45%), but then it drastically declines to only \~20% responding with "pretty well satisfied" in 2021.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-overtime-cohort
-#| fig-cap: "Financial satisfaction results over the years for each generation"
-
-combined_data |>
-  select(year, cohort_band, financial_satisfaction) |>
-  group_by(year, cohort_band) |>
-  count(financial_satisfaction) |>
-  filter(!is.na(financial_satisfaction), !(is.na(cohort_band))) |>
-  mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = financial_satisfaction )) +
-  geom_line() +
-  geom_smooth(method=lm, se=FALSE) + 
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() + 
-  theme(axis.text.x = element_text(angle = 90)) +
-  facet_wrap(vars(cohort_band), nrow = 3, ncol = 3) +
-  theme(legend.position="bottom") +
-  theme(legend.text = element_text(size = 5)) +
-  labs(
-    x = "Year",
-    y = "Percentage of Responses",
-    color = "Financial Satisfaction"
-  )
-```
-
-### Age
-
-In addition to looking at cohort over time, we also want to know how respondents at the same age groups are perceiving their financial situation over time.
-Looking at @fig-overtime-age, we see a discouraging picture where "not satisfied at all" with their financial situations is trending upwards for almost all age groups, while being "pretty well satisfied" is trending downwards.
-Out of the bands, 50-64 age group had the highest percentage increase for "not satisfied at all", and most amount of percentage decrease for "pretty well satisfied".
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-overtime-age
-#| fig-cap: "Financial satisfaction results over the years for each age band"
-
-combined_data |>
-  select(year, age_band, financial_satisfaction) |>
-  group_by(year, age_band) |>
-  count(financial_satisfaction) |>
-  filter(!is.na(financial_satisfaction), !(is.na(age_band))) |>
-  mutate(percent = n / sum(n)) |>
-  filter(financial_satisfaction != "More or Less Satisfied") |>
-  ggplot(aes(x = year, y = percent, color = age_band )) +
-  geom_line() +
-  geom_smooth(method=lm, se=FALSE) + 
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() + 
-  #theme(axis.text.x = element_text(angle = 90)) +
-  facet_wrap(vars(financial_satisfaction), ncol = 2) +
-  theme(legend.position="bottom") +
-  #theme(legend.text = element_text(size = 5)) +
-  labs(
-    x = "Year",
-    y = "Percentage of Responses",
-    color = "Age Band"
-  )
-```
-
-## Social Class (still need to fix up)
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class1)
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = social_class)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class)
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(income)) |>
-  group_by(social_class, income) |>
-  count() |>
-  ungroup() |>
-  group_by(income) |>
-  mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = social_class, y = income, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Comparing Financial Situation to Others",
-    y = "Self-Identified Social Class",
-  )
-```
-
-Social class over the years
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-combined_data |>
-  select(year, social_class) |>
-  group_by(year) |>
-  count(social_class) |>
-  filter(!is.na(social_class)) |>
-  mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = social_class)) +
-  geom_smooth(method=lm, se=FALSE) + 
-  geom_line() +
-  scale_y_continuous(labels=scales::percent) +
-  labs(
-    xlabel = "Year",
-    ylabel = "Percentage of Responses",
-    color = "Social Class"
-  )
-
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-survey_data_2021 |>
-  filter(!is.na(family_income_bands), !is.na(social_class)) |>
-  ggplot(aes(x = social_class, group = family_income_bands)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~family_income_bands)
-```
-
-The variable *realinc* has only \~30 values - probably something to do with bands that are in the raw survey data
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# survey_data_2021 |>
-#   filter(!is.na(social_class)) |>
-#   ggplot(aes(x = social_class, y = realinc)) +
-#   geom_boxplot() +
-#   geom_jitter(alpha = 0.3, width = 0.10, height = 0)
-#   #geom_point(alpha = 0.2) +
-  #geom_violin()
-```
 
@@
 ```
-
-# Discussion
-
-Our results point to two trends: first, evidence to support the hypothesis that the rich are getting richer, and that the poor are getting poorer.
-And second, evidence to support the Fleishman effect \-- the idea that upper classes have a distorted sense of their financial well-being, since they compare themselves to the upper class and opt into wealthy lifestyles, rather than comparing themselves to the general population.
-
-### (Economic) The Rich are getting Richer, and the Poor are getting Poorer
-
-Rich and richer; poor getting poorer
-
-### (Psychological) Fleishman Effect
-
-During World War II, sociologist Samuel Stouffer led the American Soldier study, which aimed to examine the attitudes and motivations of American soldiers during war time.
-The study collected data using both surveys and interviews from over 500,00 soldiers, and covered topics such as combat experiences, political beliefs, and racial attitudes.
-
-One important insight from this study is the contribution it made to our understanding of reference groups.
-The American Soldier study found that soldiers tend to identify more strongly with fellow soldiers than civilians.
-This identification was found to be a significant factor in shaping soldiers\' attitudes in relation to morale and motivation.
-
-In 1957, the sociologist Robert Merton expanded on this idea of reference groups to further develop the concept of relative deprivation, which refers to feelings of discontent that arise when individuals perceive a discrepancy between their own status, and the status of others in their social comparison group.
-Merton argues that social mobility leads to an increase in relative deprivation, since individuals who achieve higher incomes and status become more aware of the gap between their status, and those who are more affluent in their reference group.
-
-Our data on financial satisfaction based on income supports this hypothesis \[figure about financial satisfaction\].
-When faceted by class instead, the question becomes more interesting.
-When asked to compare themselves to the general population, a significant portion of self-identified upper-classes see themselves as \"below average\".
-
-Is the upper class simply delusional?
-Could survey respondents have misinterpreted what \"upper class\" means?
-Or, are there more factors at play when people consider their own financial situations \-- for instance, the child of millionaires who themselves make \$40k a year might see themselves as culturally and socially upper class, yet below average in terms of financial comparison.
-
-In her 2019 memoir We Need To Talk About Money, Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of struggling millennials.
-After learning that a friend had concealed that she, in fact, didn\'t have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that political liberal individuals can experience as  anti-capitalist sentiment have entered the mainstream conversations:
-
-\"Shame as it relates to wealth is somewhat more unexpected because -- despite occasional pockets of resistance and the growing anti-capitalist sentiment that has entered mainstream cultural conversations -- on the whole, we tend to venerate the rich, assigning enormous social currency to traditional markers of wealth: nice cars, massive houses, designer clothes.
-But that\'s the thing about shame -- it\'s an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to \'fit in\'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
-For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited.\"
-
-Concealing wealth out of shame, to Uwagba, is \"actively unhelpful\" for others in the same reference group, and exacerbates the frustration of realizing that achieving personal finance goals is even more out of reach than she initially thought.	In \[\@fig-overtime\], we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z\'s self-reporting does correlate with a pandemic).
-This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions that distort people\'s understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
-Millennials aren\'t experiencing the same financial satisfaction that Baby Boomers did when they were the same age.
-
-Meanwhile, the lure of upward mobility is addictive, as even the upper class reports financial dissatisfaction.
-An article by The Cut titled \"The Fleishman Effect\" explores how upper-class, New York women are responding to the HBO show Fleishman Is In Trouble, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where \"the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present\".
-
-## Limitations and Future Direction
-
-The study would be strengthened with data that could provide a fuller profile of respondent\'s reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
-
-	For instance, geography can have a major influence on personal finance and reference groups.
-A respondent making \$75,000 a year living in Hattiesburg, Mississippi might feel more financially satisfied than a respondent with the same income living in San Francisco, California.  However, this could sacrifice anonymity.
-This question would still have to be broad enough to provide information on their location (i.e.
-by city), rather than something that could de-anonymize respondents when triangulated with other data (i.e.
-postal code).
-
-Another key factor in how people interpret their own wealth is the way that they narrativize their success in comparison to their upbringing.
-In our supplemental survey in Appendix X, we suggest the questions \"Have you moved up a full socio-economic class compared to your upbringing?
-(Y/N/Don\'t Know\", \"Do you have at least one parent who moved up a full socio-economic class?\" (Y/N/Don\'t Know), and \"Is your current career in the same field as your parents?\" (Y/N/Don\'t Know) to better understand respondent\'s family histories.
-Asking \"Are your friends in the same socio-economic class, higher, or lower?
-(Higher/About the same/Lower)\" would also provide more granular detail into how someone perceives their financial status in comparison to the people closest to them.
-The Social Capital Atlas (<https://www.socialcapital.org/?dimension=EconomicConnectednessIndividual&geoLevel=county&selectedId=06037&dim1=EconomicConnectednessIndividual&dim2=CohesivenessSupportRatio&dim3=CivicEngagementVolunteeringRates&bigModalSection=&bigModalChart=scatterplot&showOutliers=false&colorBy=>) conducted a survey pursuing a similar estimand using Facebook data to explore \"economic connectedness\" in friend groups, which refers to the percentage of friends a respondent has who earn higher or lower incomes than them.
-
-	(fig-2021-degree-financial) shows that respondents with post-secondary education experience better changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is \"getting better\".
-Subsetting these reference groups by their area of study, and whether or not they regret their major, could provide more information on how the different returns-on-investment on different degrees change over time, as shown by The Washington Post\'s reporting on the Federal Reserve Survey of Household Economics and Decisionmaking (<https://www.washingtonpost.com/business/2022/09/02/college-major-regrets/>).
-
-We propose adding the questions \"If you went to college or university, what was your major?\" \"Do you regret your major?
-(Y/N/Don\'t know)\" to the survey.
-Adding this question to our study would provide more detailed insight on reference groups, since we could see how people who enrolled in each major compare themselves to others, and, with more information on family history, we could also gather insight on which fields individuals from different social classes pursue in higher education.
-We could also see how people across different income ranges and classes experience regret about their decisions.
-
-\
-
-Additional financial variables in survey for analysis
-
-Connect with US census data to check representativeness
-
-Connect with other macroeconomic indicators - e.g. inflation, stock markets
-
-Collect more data to supplement our findings -\> link to survey <https://docs.google.com/forms/d/e/1FAIpQLSdqe87Bet1PyaCezPxN_t7A1smQCpSap6EYjG2LrTN0K_CX9A/viewform>
-
-# Conclusion
-
-In conclusion, financial satisfaction and well-being are subjective and influenced by various factors.
-This paper used data from the US General Social Survey to examine how people perceive their financial situations in comparison to others, particularly in light of the increasing wealth gap between high and low earning individuals.
-
-The results support the hypothesis that the rich are getting richer, the poor are getting poorer, and that younger generations are financially worse off than preceding cohorts.
-The data also suggests that people's reference groups influence their understanding of their financial well-being and satisfaction.
-Overall, this paper highlights the complexity of financial satisfaction and well-being, and the importance of considering sociological dynamics and cognitive biases in understanding these subjective evaluations.
-
-# Reference

### 2023-03-10 outputs/paper/paper3.qmd
Fixed up formatting and references

@@
 
-# Story
-
-Introduction - news, inflation, millennials and younger not doing better than their parents
-
-\% of financial compare & % financial satisfaction - in 2021
-
-Let's dig deeper into 2021
-
--   gender
-
--   education
-
--   Social class
-
--   age
-
-How financial compare / financial satisfaction changed over the years - follow cohort and age
-
-How does social class impact these perceptions? - link to Michaela's literature
-
-# To Do
-
-Christina
-
--   financial satisfaction
-
--   Social status
-
--   Simulated data
-
--   Data validation tests
-
--   README
-
-Michaela
-
--   Supplementary survey: The survey should have an introductory section and include the details of a contact person.
-The survey questions should be well constructed and appropriate to the task.
-The questions should have an appropriate ordering.
-A final section should thank the respondent
-
--   Introduction
-
--   Discussions
-
--   Limitations & Future Work
-
 # Introduction
@@
 
-In this paper, we use data from the US General Social Survey (GSS) from NORC at the University of Chicago to better understand how people understand their own financial situations in comparison to others, against the background an increasing wealth gap between higher and lower earning individuals.
+In this paper, we use data from the US General Social Survey (GSS) from NORC [@gss] at the University of Chicago to better understand how people understand their own financial situations in comparison to others, against the background an increasing wealth gap between higher and lower earning individuals.
 We subset and visualize the data by a variety of different demographic factors, such as gender, age, level of education, along with views of how different generational cohorts report their financial well-being over time.
@@
 | Variable | New Name               | Description                                                                             | Example               |
-|----------|------------------------|-----------------------------------------------------------------------------------------|-----------------------|
+|----------------|----------------|------------------------|----------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 
+Let's look specifically at social class.
+
 ```{r}
@@
 
-## Financial Compare (old)
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(social_class1), !is.na(financial_compare)) |>
-  ggplot(aes(x = financial_compare, group = social_class1)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Compare") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class1)
-
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(family_income_bands), !is.na(financial_compare)) |>
-  ggplot(aes(x = financial_compare, group = family_income_bands)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Compare") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~family_income_bands)
-
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(financial_change)) |>
-  ggplot(aes(x = financial_change, group = social_class)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Change") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class) 
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-# survey_data_2016 |>
-#   filter(!is.na(social_class), !is.na(financial_change)) |>
-#   ggplot(aes(x = financial_change, group = social_class)) +
-#   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-#   scale_y_continuous(labels=scales::percent) +
-#   ylab("Percentage of response") +
-#   xlab("Financial Change") +
-#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-#   theme(legend.position="none") +
-#   facet_grid(~social_class) 
-```
-
 # Discussion
@@
 
-Rich and richer; poor getting poorer
+(to do) Rich and richer; poor getting poorer
 
@@
 
-During World War II, sociologist Samuel Stouffer led the American Soldier study, which aimed to examine the attitudes and motivations of American soldiers during war time.
+During World War II, sociologist Samuel Stouffer led the American Soldier study, which aimed to examine the attitudes and motivations of American soldiers during war time [@samuelStouffer].
 The study collected data using both surveys and interviews from over 500,00 soldiers, and covered topics such as combat experiences, political beliefs, and racial attitudes.
-
 One important insight from this study is the contribution it made to our understanding of reference groups.
@@
 
-In 1957, the sociologist Robert Merton expanded on this idea of reference groups to further develop the concept of relative deprivation, which refers to feelings of discontent that arise when individuals perceive a discrepancy between their own status, and the status of others in their social comparison group.
+In 1957, the sociologist Robert Merton expanded on this idea of reference groups to further develop the concept of relative deprivation [@relativeDeprivation], which refers to feelings of discontent that arise when individuals perceive a discrepancy between their own status, and the status of others in their social comparison group.
 Merton argues that social mobility leads to an increase in relative deprivation, since individuals who achieve higher incomes and status become more aware of the gap between their status, and those who are more affluent in their reference group.
@@
 
-In her 2019 memoir We Need To Talk About Money, Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of struggling millennials.
+In her 2019 memoir We Need To Talk About Money [@talkAboutMoney], Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of struggling millennials.
 After learning that a friend had concealed that she, in fact, didn't have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that political liberal individuals can experience as  anti-capitalist sentiment have entered the mainstream conversations:
@@
 Meanwhile, the lure of upward mobility is addictive, as even the upper class reports financial dissatisfaction.
-An article by The Cut titled "The Fleishman Effect" explores how upper-class, New York women are responding to the HBO show Fleishman Is In Trouble, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
+An article by The Cut titled "The Fleishman Effect" [@fleishmanEffect] explores how upper-class, New York women are responding to the HBO show Fleishman Is In Trouble, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
 
@@
 (Higher/About the same/Lower)" would also provide more granular detail into how someone perceives their financial status in comparison to the people closest to them.
-The Social Capital Atlas (<https://www.socialcapital.org/?dimension=EconomicConnectednessIndividual&geoLevel=county&selectedId=06037&dim1=EconomicConnectednessIndividual&dim2=CohesivenessSupportRatio&dim3=CivicEngagementVolunteeringRates&bigModalSection=&bigModalChart=scatterplot&showOutliers=false&colorBy=>) conducted a survey pursuing a similar estimand using Facebook data to explore "economic connectedness" in friend groups, which refers to the percentage of friends a respondent has who earn higher or lower incomes than them.
-
-    (fig-2021-degree-financial) shows that respondents with post-secondary education experience better changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is \"getting better\".
-Subsetting these reference groups by their area of study, and whether or not they regret their major, could provide more information on how the different returns-on-investment on different degrees change over time, as shown by The Washington Post\'s reporting on the Federal Reserve Survey of Household Economics and Decisionmaking (<https://www.washingtonpost.com/business/2022/09/02/college-major-regrets/>).
+The Social Capital Atlas [@socialCapital] conducted a survey pursuing a similar estimand using Facebook data to explore "economic connectedness" in friend groups, which refers to the percentage of friends a respondent has who earn higher or lower incomes than them.
+
+(fig-2021-degree-financial) shows that respondents with post-secondary education experience better changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is "getting better".
+Subsetting these reference groups by their area of study, and whether or not they regret their major, could provide more information on how the different returns-on-investment on different degrees change over time, as shown by The Washington Post's reporting on the Federal Reserve Survey of Household Economics and Decisionmaking [@collegeMajors].
 
@@
 
-\
-
 Additional financial variables in survey for analysis
@@
 Connect with other macroeconomic indicators - e.g. inflation, stock markets
-
-Collect more data to supplement our findings -\> link to survey <https://docs.google.com/forms/d/e/1FAIpQLSdqe87Bet1PyaCezPxN_t7A1smQCpSap6EYjG2LrTN0K_CX9A/viewform>
 
@@
 
+# Appendix
+
+Here is a link to our supplemental survey:
+
+<https://docs.google.com/forms/d/e/1FAIpQLSdqe87Bet1PyaCezPxN_t7A1smQCpSap6EYjG2LrTN0K_CX9A/viewform>
+
 # Reference

### 2023-03-10 outputs/paper/paper3.qmd
wrap graph generation into help functions to make them reusable

@@
 | Variable | New Name               | Description                                                                             | Example               |
-|----------------|----------------|------------------------|----------------|
+|------------------|------------------|-------------------|------------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 
+generate_heatmap(
+  ggplot_data = 
     survey_data_2021 |>
@@
     mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = financial_compare, y = social_class, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Comparing Financial Situation to Others",
-    y = "Self-Identified Social Class",
+    ggplot(aes(x = financial_compare, y = social_class, fill = Percentage)),
+  xlabel = "Comparing Financial Situation to Others",
+  ylabel = "Self-Identified Social Class"
 )
@@
 
+generate_compare_bar_graph(
+  ggplot_data = 
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_satisfaction, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
-
+    ggplot(aes(x = financial_satisfaction, y = pct, fill = gender)),
+  xlabel = "Financial Satisfaction",
+  ylabel = "Percentage of Responses",
+  fill_label = "Gender"
+)
+
+generate_compare_bar_graph(
+  ggplot_data =
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_change, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Gender"
-  )
-
+    ggplot(aes(x = financial_change, y = pct, fill = gender)),
+  xlabel = "Change in Financial Situation",
+  ylabel = "Percentage of Responses",
+  fill_label = "Gender"
+)
+
+generate_compare_bar_graph(
+  ggplot_data =
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = financial_compare, y = pct, fill = gender)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Comparing to Others",
-    y = "Percentage of Responses",
-    fill = "Gender"
+    ggplot(aes(x = financial_compare, y = pct, fill = gender)),
+  xlabel = "Comparing to Others",
+  ylabel = "Percentage of Responses",
+  fill_label = "Gender"
 )
@@
 
+generate_compare_bar_graph(
+  ggplot_data =
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-
+    ggplot(aes(x = age_band, y = pct, fill = financial_satisfaction)),
+  xlabel = "Age Group",
+  ylabel = "Percentage of Responses",
+  fill_label = "Financial Satisfaction"  
+)
+
+generate_compare_bar_graph(
+  ggplot_data =
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = age_band, y = pct, fill = financial_change)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Age Group"
+    ggplot(aes(x = age_band, y = pct, fill = financial_change)),
+  xlabel = "Age Group",
+  ylabel = "Percentage of Responses",
+  fill_label = "Change in Financial Situation" 
 )
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = degree, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Financial Satisfaction",
-    y = "Percentage of Responses",
-    fill = "Age Group"
-  )
-
+    ggplot(aes(x = degree, y = pct, fill = financial_satisfaction)),
+  xlabel = "Level of Education",
+  ylabel = "Percentage of Responses",
+  fill_label = "Financial Satisfaction"
+)
+
+generate_compare_bar_graph(
+  ggplot_data =
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = degree, y = pct, fill = financial_change)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Change in Financial Situation",
-    y = "Percentage of Responses",
-    fill = "Age Group"
+    ggplot(aes(x = degree, y = pct, fill = financial_change)),
+  xlabel = "Level of Education",
+  ylabel = "Percentage of Responses",
+  fill_label = "Change in Financial Situation"
 )
@@
 
+generate_heatmap(
+  ggplot_data =
     survey_data_2021 |>
@@
     mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = financial_satisfaction, y = income, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Financial Satisfaction",
-    y = "Family Income in 2016",
-  )
-
+    ggplot(aes(x = financial_satisfaction, y = income, fill = Percentage)),
+  xlabel = "Financial Satisfaction",
+  ylabel = "Family Income in 2016",
+  position = "bottom",
+  angle = 0,
+  hjust = NULL,
+  vjust = NULL
+)
+
+generate_heatmap(
+  ggplot_data =
     survey_data_2021 |>
@@
     mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = financial_change, y = income, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Comparing Financial Situation to Others",
-    y = "Self-Identified Social Class",
-  )
-
+    ggplot(aes(x = financial_change, y = income, fill = Percentage)),
+  xlabel = "Comparing Financial Situation to Others",
+  ylabel = "Family Income in 2016",
+  position = "bottom",
+  angle = 0,
+  hjust = NULL,
+  vjust = NULL
+)
 ```
@@
 
-#Legend wrap line reference: https://datavizpyr.com/fold-legend-into-two-rows-in-ggplot2/
-
+generate_time_series_graph(
+  ggplot_data =
     combined_data |>
@@
     mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = financial_satisfaction)) +
-  geom_smooth(method=lm, se=FALSE) + 
-  geom_line() +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() + 
-  theme(legend.position="bottom") + 
-  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
-  labs(
-    x = "Year",
-    y = "Percentage of Responses",
-    color = "Financial Satisfaction"
-  )
-
+    ggplot(aes(x = year, y = percent, color = financial_satisfaction)),
+  xlabel = "Year",
+  ylabel = "Percentage of Responses",
+  color_label = "Financial Satisfaction"
+)
+
+generate_time_series_graph(
+  ggplot_data =
     combined_data |>
@@
     mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = financial_compare)) +
-  geom_smooth(method=lm, se=FALSE) + 
-  geom_line() +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() + 
-  theme(legend.position="bottom") +
-  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
-  labs(
-    x = "Year",
-    y = "Percentage of Responses",
-    color = "Financial Compare"
+    ggplot(aes(x = year, y = percent, color = financial_compare)),
+  xlabel = "Year",
+  ylabel = "Percentage of Responses",
+  color_label = "Comparing Financial Situation to Others"
   )
@@
 
+generate_time_series_graph(
+  ggplot_data =
     combined_data |>
@@
     mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = financial_satisfaction )) +
-  geom_line() +
-  geom_smooth(method=lm, se=FALSE) + 
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() + 
+    ggplot(aes(x = year, y = percent, color = financial_satisfaction )),
+  ylabel = "Year",
+  xlabel = "Percentage of Responses",
+  color_label = "Financial Satisfaction"
+) +
 theme(axis.text.x = element_text(angle = 90)) +
-  facet_wrap(vars(cohort_band), nrow = 3, ncol = 3) +
-  theme(legend.position="bottom") +
-  theme(legend.text = element_text(size = 5)) +
-  labs(
-    x = "Year",
-    y = "Percentage of Responses",
-    color = "Financial Satisfaction"
-  )
+facet_wrap(vars(cohort_band), nrow = 3, ncol = 3)
+#theme(legend.text = element_text(size = 5))
 ```
@@
     filter(financial_satisfaction != "More or Less Satisfied") |>
-  ggplot(aes(x = year, y = percent, color = age_band )) +
-  geom_line() +
-  geom_smooth(method=lm, se=FALSE) + 
+    ggplot(aes(x = year, y = percent, color = age_band )),
+  xlabel = "Year",
+  ylabel = "Percentage of Responses",
+  color_label = "Age Band"
+) +
+facet_wrap(vars(financial_satisfaction), ncol = 2)
+```
+
+## Social Class
+
+Let's look specifically at the survey results for social class and how it correlates with the other variables.
+There are two survey questions related to how respondents selected the social class they belonged to: one used a classification of 4 classes (lower class, working class, middle class, and upper class), and a second survey question has an expanded 6 classes, with the additional classes of `lower middle class` and `upper middle class`.
+
+First, we are interested in the correlation between family income and social class.
+Looking at the 4 band classification of social class, we see in @fig-socialclass-1 that less than 10% of respondents classified themselves as **lower class**.
+Digging deeper into the family income of this segment in @fig-income-social , the majority of respondents (\~90%) in this class are making less than \$40K annually.
+There are 37% of respondents who classified themselves as **working class**.
+There is a wide range of their family income, with the majority (\~71%) clustered between \$10K to \$75K.
+Almost half of the respondent have classified themselves as **middle class**, with family income clustered between \$60K to \$170K (\~57%).
+It is interesting to note that those making over \$170K have the highest percentage (15%) compared across income bands for those who ranked themselves in this class.
+Lastly, out of the 4% of the respondents who responded that they belong to the **upper class**, most of them have family income over \$170K (67%).
+An interesting observation in this social class is that more than 10% of respondents making less than \$40K a year also put themselves as upper class.
+
+The 6 band social class survey expands the middle class into three categories: lower middle class, middle class, and upper middle class.
+Comparing the distribution between the two social class variables (@fig-socialclass), there are movements across the bands.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+
+survey_data_2021 |>
+  filter(is.na(social_class1)) |>
+  group_by(social_class) |>
+  count() |>
+  ungroup() |>
+  mutate(Percentage = n / sum(n))
+
+survey_data_2021 |>
+  filter(!is.na(social_class), !is.na(income)) |>
+  group_by(social_class, income) |>
+  count() |>
+  ungroup() |>
+  group_by(social_class) |>
+  mutate(Percentage = n / sum(n)) |>
+  mutate(Cum = cumsum(Percentage)) |>
+  filter (social_class == "Upper Class")
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-socialclass
+#| fig-cap: "Distribution of survey results for social class"
+#| fig-subcap: ["Social Class (4-Bands)", "Social Class (6-Bands)"]
+#| layout-ncol: 2
+
+generate_percentage_bar_graph(
+  ggplot_data = 
+    survey_data_2021 |>
+    filter(!is.na(social_class)) |>
+    ggplot(mapping = aes(x = social_class)),
+  xlabel = "Social Class (4-Bands)",
+  angle = 45,
+  vjust = 1.1,
+  hjust = 1
+)
+
+generate_percentage_bar_graph(
+  ggplot_data = 
+    survey_data_2021 |>
+    filter(!is.na(social_class1)) |>
+    ggplot(mapping = aes(x = social_class1)),
+  xlabel = "Social Class (6-Bands)",
+  angle = 45,
+  vjust = 1.1,
+  hjust = 1
+)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-income-social
+#| fig-cap: "How respondents in each family income range responded for social class"
+#| fig-subcap: ["Social Class (4-Bands)", "Social Class (6-Bands)"]
+#| layout-ncol: 2
+
+generate_heatmap(
+  ggplot_data = 
+    survey_data_2021 |>
+    filter(!is.na(social_class), !is.na(income)) |>
+    group_by(social_class, income) |>
+    count() |>
+    ungroup() |>
+    group_by(social_class) |>
+    mutate(Percentage = n / sum(n)) |>
+    ggplot(aes(x = social_class, y = income, fill = Percentage)),
+  position = "bottom",
+  xlabel = "Social Class (4-Bands)",
+  ylabel = "Family Income in 2016",
+)
+
+generate_heatmap(
+  ggplot_data = 
+    survey_data_2021 |>
+    filter(!is.na(social_class1), !is.na(income)) |>
+    group_by(social_class1, income) |>
+    count() |>
+    ungroup() |>
+    group_by(social_class1) |>
+    mutate(Percentage = n / sum(n)) |>
+    ggplot(aes(x = social_class1, y = income, fill = Percentage)),
+  position = "bottom",
+  xlabel = "Social Class (6-Bands)",
+  ylabel = "Family Income in 2016",
+)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-socialclass-financial
+#| fig-cap: "Financial satisfaction responses for each social class"
+#| fig-subcap: ["Social Class (4-Bands)", "Social Class (6-Bands)"]
+#| layout-ncol: 2
+
+survey_data_2021 |>
+  filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
+  group_by(social_class, financial_satisfaction) |>
+  count() |>
+  ungroup() |>
+  group_by(social_class) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = social_class, y = pct, fill = financial_satisfaction)) +
+  geom_bar(stat="identity", position="dodge") +
   scale_y_continuous(labels=scales::percent) +
   theme_minimal() +
-  #theme(axis.text.x = element_text(angle = 90)) +
-  facet_wrap(vars(financial_satisfaction), ncol = 2) +
   theme(legend.position="bottom") + 
-  #theme(legend.text = element_text(size = 5)) +
   labs(
-    x = "Year",
+    x = "Social Class (4-Bands)",
     y = "Percentage of Responses",
-    color = "Age Band"
-  )
-```
-
-## Social Class (still need to fix up)
-
-Let's look specifically at social class.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+    fill = "Financial Satisfaction"
+  )
+
 survey_data_2021 |>
   filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
+  group_by(social_class1, financial_satisfaction) |>
+  count() |>
+  ungroup() |>
+  group_by(social_class1) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = social_class1, y = pct, fill = financial_satisfaction)) +
+  geom_bar(stat="identity", position="dodge") +
   scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class1)
-
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
-  ggplot(aes(x = financial_satisfaction, group = social_class)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~social_class)
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(income)) |>
-  group_by(social_class, income) |>
-  count() |>
-  ungroup() |>
-  group_by(income) |>
-  mutate(Percentage = n / sum(n)) |>
-  ggplot(aes(x = social_class, y = income, fill = Percentage)) +
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  labs (
-    x = "Comparing Financial Situation to Others",
-    y = "Self-Identified Social Class",
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Social Class (6-Bands)",
+    y = "Percentage of Responses",
+    fill = "Financial Satisfaction"
   )
@@
 survey_data_2021 |>
-  filter(!is.na(family_income_bands), !is.na(social_class)) |>
-  ggplot(aes(x = social_class, group = family_income_bands)) +
+  filter(!is.na(income), !is.na(social_class)) |>
+  ggplot(aes(x = social_class, group = income)) +
   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
@@
   theme(legend.position="none") +
-  facet_grid(~family_income_bands)
-```
-
-The variable *realinc* has only \~30 values - probably something to do with bands that are in the raw survey data
+  facet_grid(~income)
+```
 

### 2023-03-10 scripts/04-helper_functions.R

@@ -8,17 +8,63 @@
 
 library(tidyverse)
 
+# Generate percentage graph
 generate_percentage_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = "Percentage of Responses", angle = NULL, hjust = NULL, vjust = NULL) {
 
-# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-ggplot_data + 
-geom_bar(aes(y = (..count..) / sum(..count..))) +
-scale_y_continuous(labels=scales::percent) +
-theme_minimal() +
-theme(axis.text.x = element_text(angle = angle, hjust = hjust, vjust = vjust)) + #rotate x-axis & smaller font
-labs (
-    x = xlabel,
-    y = ylabel
-  )
+  # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+  ggplot_data + 
+  geom_bar(aes(y = (..count..) / sum(..count..))) +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = angle, hjust = hjust, vjust = vjust)) + #rotate x-axis & smaller font
+  labs (
+      x = xlabel,
+      y = ylabel
+    )
+}
 
+# Generate heatmap
+generate_heatmap = function(ggplot_data, xlabel = NULL, ylabel = NULL, position = NULL, angle = 45, hjust = 1, vjust = 1) {
+  
+  ggplot_data + 
+  geom_tile() + 
+  scale_fill_gradient2(low="white", high="darkblue", guide="colorbar") +
+  theme(axis.text.x = element_text(angle = angle, vjust = vjust, hjust = hjust)) +
+  theme(legend.position = position) + 
+  labs (
+    x = xlabel,
+    y = ylabel,
+  )
+}
+
+# Generate comparative bar graph
+generate_compare_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = NULL, fill_label = NULL, angle = NULL, hjust = NULL, vjust = NULL) {
+
+  ggplot_data +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = xlabel,
+    y = ylabel,
+    fill = fill_label
+  )
+}
+
+# Generate time series graph with linear regression
[…]

### 2023-03-10 outputs/paper/paper3.qmd
Complete first pass at social status. Fixed up minor errors

@@
 | Variable | New Name               | Description                                                                             | Example               |
-|------------------|------------------|-------------------|------------------|
+|------|---------------|----------------------|---------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 Looking at the 4 band classification of social class, we see in @fig-socialclass-1 that less than 10% of respondents classified themselves as **lower class**.
-Digging deeper into the family income of this segment in @fig-income-social , the majority of respondents (\~90%) in this class are making less than \$40K annually.
+Digging deeper into the family income of this segment in @fig-income-social, the majority of respondents (\~90%) in this class are making less than \$40K annually.
 There are 37% of respondents who classified themselves as **working class**.
@@
 The 6 band social class survey expands the middle class into three categories: lower middle class, middle class, and upper middle class.
-Comparing the distribution between the two social class variables (@fig-socialclass), there are movements across the bands.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-survey_data_2021 |>
-  filter(is.na(social_class1)) |>
-  group_by(social_class) |>
-  count() |>
-  ungroup() |>
-  mutate(Percentage = n / sum(n))
-
-survey_data_2021 |>
-  filter(!is.na(social_class), !is.na(income)) |>
-  group_by(social_class, income) |>
-  count() |>
-  ungroup() |>
-  group_by(social_class) |>
-  mutate(Percentage = n / sum(n)) |>
-  mutate(Cum = cumsum(Percentage)) |>
-  filter (social_class == "Upper Class")
-```
+Comparing the distribution between the two social class variables (@fig-socialclass), there are some movements across the bands; for example, there are less respondents who rated themselves as lower class and upper class.
+Looking at the income levels associated with **upper class** in this 6-band classifcation, there are some peculiar observations.
+While most individuals in this social class responded with making over \$170K annually, 20% of them reported family income less than \$10K.
 
@@
 ```
+
+Looking at the correlation between social class and financial satisfaction, we saw the expected trend in @fig-socialclass-financial-1 that as individuals move up to a higher social class, they report higher financial satisfaction with their current situation.
+However, when we look at the detailed social class breakdown in @fig-socialclass-financial-2, we noticed a trend reversal for **upper class**, where close to 20% of the respondents reported not satisfied at all with their financial situation, which is much higher compared to upper middle class with only less than 2% dissatisfaction.
+We will hypothesize potential explanations for this observation in the the discussion section below on [The Fleishman Effect].
 
@@
 
-Social class over the years
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
+Looking at the percentages of respondents in different social class over the years in @fig-socialclass-overtime, we discovered a declining trend for working class and middle class, while the percentage of those who reported as lower class has been increasing over time.
+Upper class percentage stays mainly constant from 1972 to 2021 at around 3-4%.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-socialclass-overtime
+#| fig-cap: "Percentage of respondents in each social class (4-bands) over time"
+
+generate_time_series_graph(
+  ggplot_data =
     combined_data |>
     select(year, social_class) |>
+    group_by(year, social_class) |>
+    count() |>
+    filter(!is.na(social_class)) |>
     group_by(year) |>
-  count(social_class) |>
-  filter(!is.na(social_class)) |>
     mutate(percent = n / sum(n)) |>
-  ggplot(aes(x = year, y = percent, color = social_class)) +
-  geom_smooth(method=lm, se=FALSE) + 
-  geom_line() +
-  scale_y_continuous(labels=scales::percent) +
-  labs(
+    ggplot(aes(x = year, y = percent, color = social_class)),
   xlabel = "Year",
   ylabel = "Percentage of Responses",
-    color = "Social Class"
-  )
-
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-survey_data_2021 |>
-  filter(!is.na(income), !is.na(social_class)) |>
-  ggplot(aes(x = social_class, group = income)) +
-  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
-  scale_y_continuous(labels=scales::percent) +
-  ylab("Percentage of response") +
-  xlab("Financial Satisfaction") +
-  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-  theme(legend.position="none") +
-  facet_grid(~income)
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-
-# survey_data_2021 |>
-#   filter(!is.na(social_class)) |>
-#   ggplot(aes(x = social_class, y = realinc)) +
-#   geom_boxplot() +
-#   geom_jitter(alpha = 0.3, width = 0.10, height = 0)
-#   #geom_point(alpha = 0.2) +
-  #geom_violin()
+  color_label = "Social Class",
+  position = NULL,
+  nrow = NULL
+)
 ```
@@
 
-### (Economic) The Rich are getting Richer, and the Poor are getting Poorer
+## Marginalization
+
+-   Men are doing better than women
+
+-   Level of education is a big factor in determining financial satisfaction
+
+-   Generation gap?
+
+-   Even though GDP rising, we don't see significant improvement in financial satisfaction over time
+
+-   Increase of lower class % over time
+
+## (Economic) Polarization of Financial Satisfaction
 
@@
 
-### (Psychological) Fleishman Effect
+-   @fig-overtime polarization of financial situation - people are either getting better or getting worse, but not many of them are staying "average". Dissatisfaction overall is rising.
+
+-   Not much changes in social class across the years
+
+-   Those who are making lower income are less satisfied
+
+## (Psychological) Fleishman Effect
 
@@
 
-## Limitations and Future Direction
+## Future Research Directions
 
@@
 We could also see how people across different income ranges and classes experience regret about their decisions.
-
-Additional financial variables in survey for analysis
-
-Connect with US census data to check representativeness
-
-Connect with other macroeconomic indicators - e.g. inflation, stock markets
+Please see [Appendix] to view the full supplementary survey that we have put together.
+
+-   Additional financial variables in survey for analysis
+
+-   Connect with US census data to check representativeness
+
+-   Connect with other macroeconomic indicators - e.g. inflation, stock markets
 

### 2023-03-10 scripts/04-helper_functions.R

@@ -53,15 +53,15 @@ generate_compare_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = NULL,
 }
 
 # Generate time series graph with linear regression
-generate_time_series_graph = function(ggplot_data, xlabel = NULL, ylabel = NULL, color_label = NULL) {
+generate_time_series_graph = function(ggplot_data, xlabel = NULL, ylabel = NULL, color_label = NULL, position = "bottom", nrow = 2) {
   
   ggplot_data +
   geom_smooth(method=lm, se=FALSE) + 
   geom_line() +
   scale_y_continuous(labels=scales::percent) +
   theme_minimal() + 
-  theme(legend.position="bottom") + 
-  guides(color=guide_legend(nrow=2, byrow=TRUE)) + #reference: https://datavizpyr.com/fold-legend-into-two-rows-in-ggplot2/
+  theme(legend.position = position) + 
+  guides(color=guide_legend(nrow=nrow, byrow=TRUE)) + #reference: https://datavizpyr.com/fold-legend-into-two-rows-in-ggplot2/
   labs(
     x = xlabel,
     y = ylabel,

### 2023-03-11 outputs/paper/paper3.qmd
minor update to Appendix heading
added financial compare graph for social class. edited discussion

@@
 title-block-banner: true
-abstract: "This paper examines the subjective evaluations of financial satisfaction and well-being, and how they are influenced by various factors.
-Using data from the US General Social Survey, we explore how people understand their financial situations in comparison to others, subsetting by different demographic factors, against the backdrop of an increasing wealth gap between high and low earning individuals.
+abstract: "This paper examines the subjective evaluations of financial satisfaction and well-being, and how they are influenced by various demographic factors.
+Using data from the US General Social Survey, we explore how people understand their financial situations in comparison to others, according to different demographic factors, against the backdrop of an increasing wealth gap between high and low earning individuals.
 The paper discusses the sociological dynamics and cognitive biases that affect these subjective evaluations, and highlights the importance of considering these factors in understanding financial satisfaction and well-being."
@@
 | Variable | New Name               | Description                                                                             | Example               |
-|------|---------------|----------------------|---------------|
+|----------------|----------------|------------------------|----------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 However, when we look at the detailed social class breakdown in @fig-socialclass-financial-2, we noticed a trend reversal for **upper class**, where close to 20% of the respondents reported not satisfied at all with their financial situation, which is much higher compared to upper middle class with only less than 2% dissatisfaction.
-We will hypothesize potential explanations for this observation in the the discussion section below on [The Fleishman Effect].
+In @fig-socialclass-compare, only 60% of upper-class respondents see themselves as far above average compared to the general population.
+We will hypothesize potential explanations for this observation in the the discussion section below on \[The Fleishman Effect\].
 
@@
 
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-socialclass-compare
+#| fig-cap: "Financial comparison responses for each social class"
+#| fig-subcap: ["Social Class (4-Bands)", "Social Class (6-Bands)"]
+
+survey_data_2021 |>
+  filter(!is.na(social_class1), !is.na(financial_compare)) |>
+  group_by(social_class1, financial_compare) |>
+  count() |>
+  ungroup() |>
+  group_by(social_class1) |>
+  mutate(pct = percent(n / sum(n), digit = 1)) |>
+  ggplot(aes(x = social_class1, y = pct, fill = financial_compare)) +
+  geom_bar(stat="identity", position="dodge") +
+  scale_y_continuous(labels=scales::percent) +
+  theme_minimal() +
+  theme(legend.position="bottom") + 
+  labs(
+    x = "Social Class (6-Bands)",
+    y = "Percentage of Responses",
+    fill = "Financial Compare"
+  )
+
+```
+
 Looking at the percentages of respondents in different social class over the years in @fig-socialclass-overtime, we discovered a declining trend for working class and middle class, while the percentage of those who reported as lower class has been increasing over time.
@@
 
-Our results point to two trends: first, evidence to support the hypothesis that the rich are getting richer, and that the poor are getting poorer.
-And second, evidence to support the Fleishman effect \-- the idea that upper classes have a distorted sense of their financial well-being, since they compare themselves to the upper class and opt into wealthy lifestyles, rather than comparing themselves to the general population.
-
-## Marginalization
-
--   Men are doing better than women
-
--   Level of education is a big factor in determining financial satisfaction
-
--   Generation gap?
-
--   Even though GDP rising, we don't see significant improvement in financial satisfaction over time
-
--   Increase of lower class % over time
-
-## (Economic) Polarization of Financial Satisfaction
-
-(to do) Rich and richer; poor getting poorer
-
--   @fig-overtime polarization of financial situation - people are either getting better or getting worse, but not many of them are staying "average". Dissatisfaction overall is rising.
-
--   Not much changes in social class across the years
-
--   Those who are making lower income are less satisfied
-
-## (Psychological) Fleishman Effect
+Our results point to three trends: first, our visualizations illustrate a growing wealth gap over time.
+Second, the data suggests that financial well-being and satisfaction is influenced by demographic factors.
+And third, evidence to support the Fleishman effect \-- the idea that upper classes have a distorted sense of their financial well-being, since they compare themselves to the upper class, rather than comparing themselves to the general population.
+
+## Growing Wealth Gap Impacts Demographic Groups Differently
+
+Our analysis shows that the gap between high-earning and low-earning individuals in the US has been increasing since 1970, when the GSS was first recorded.
+Even though GDP is rising, we don't observe a significant improvement in financial satisfaction over time.
+In fact, twice as many respondents identify as "lower class" in 2021 than in 1970.
+
+Respondents' gender (defined in the GSS as sex at birth) and level of education have a significant influence on financial satisfaction.
+Men tend to see their financial satisfaction as improving, and report greater levels of satisfaction than women (@fig-2021-gender-financial).
+Respondents with Bachelors or Graduate degrees are more satisfied, and describe their financial situations as getting better over time (@fig-2021-degree-financial) .
+This suggests that the increased financial dissatisfaction reported over time in @fig-overtime-1 is more likely to be felt by non-male individuals with less formal education.
+
+In addition to gender and education, generational cohorts are impacted differently by the growing wealth gap.
+In @fig-overtime-cohort, the Greatest, Silent, and Boomer generations all see an increase in individuals reporting "Pretty Well Satisfied" over the course of their lives.
+In Gen X, this "Pretty Well Satisfied" line flattens, and Millennials and Gen Z see a significant decrease in high satisfaction (of course, Gen Z's visualization is correlated with a global pandemic).
+
+While satisfaction can often be correlated with income -- for instance, in @fig-2021-income-financial those who are making lower income are less satisfied-- it shouldn't be confused with income.
+The way somebody feels about the financial situation might not reflect the reality of their financial situation as held against some other objective measure.
+In @fig-overtime, we see that reported financial situations are either getting better or getting worse over time, but that dissatisfaction overall is rising (with nearly 20% of upper-class respondents reporting dissatisfaction in @fig-socialclass-financial).
+Although the levels of respondents identifying as upper and middle classes has remained relatively stable over time (@fig-socialclass-overtime), there is a general decline is people reporting being "Pretty Well Satisfied", and an increase in individuals reporting "Not Satisfied At All" (@fig-overtime).
+Who is feeling this dissatisfaction?
+Why?
+
+## Reference Groups and The "Fleishman Effect"
 
@@
 
-Our data on financial satisfaction based on income supports this hypothesis \[figure about financial satisfaction\].
+Our data on financial satisfaction based on income supports this hypothesis (@fig-socialclass-compare).
 When faceted by class instead, the question becomes more interesting.
-When asked to compare themselves to the general population, a significant portion of self-identified upper-classes see themselves as "below average".
+When asked to compare themselves to the general population, nearly 20% of self-identified upper-classes see themselves as "below average".
+While this may be due to problems with the survey recording, or respondents misunderstanding the question, the fact that only 60% of upper class individuals see themselves as being far above average suggests some sort of distortion of self-perception once that level of wealth is achieved.
 
@@
 In her 2019 memoir We Need To Talk About Money [@talkAboutMoney], Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of struggling millennials.
-After learning that a friend had concealed that she, in fact, didn't have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that political liberal individuals can experience as  anti-capitalist sentiment have entered the mainstream conversations:
-
-"Shame as it relates to wealth is somewhat more unexpected because -- despite occasional pockets of resistance and the growing anti-capitalist sentiment that has entered mainstream cultural conversations -- on the whole, we tend to venerate the rich, assigning enormous social currency to traditional markers of wealth: nice cars, massive houses, designer clothes.
-But that's the thing about shame -- it's an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to 'fit in'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
+After learning that a friend had concealed that she, in fact, didn't have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that politically liberal high-earners (or inheritors) can experience as anti-capitalist sentiments have entered the mainstream conversations:
+
+"But that's the thing about shame -- it's an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to 'fit in'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
 For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited."
 
-Concealing wealth out of shame, to Uwagba, is "actively unhelpful" for others in the same reference group, and exacerbates the frustration of realizing that achieving personal finance goals is even more out of reach than she initially thought.
-In \[\@fig-overtime\], we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z's self-reporting does correlate with a pandemic).
-This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions that distort people's understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
+Concealing wealth out of shame, to Uwagba, is "actively unhelpful" for others in the same reference group, and exacerbates the frustration of realizing that achieving personal finance goals may even more out of reach than initially believed.
+In @fig-overtime, we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z's self-reporting does correlate with a pandemic).
+This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions such as shame that distort people's understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
 Millennials aren't experiencing the same financial satisfaction that Baby Boomers did when they were the same age.
@@
 Meanwhile, the lure of upward mobility is addictive, as even the upper class reports financial dissatisfaction.
-An article by The Cut titled "The Fleishman Effect" [@fleishmanEffect] explores how upper-class, New York women are responding to the HBO show Fleishman Is In Trouble, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
+An article by The Cut titled "The Fleishman Effect" [@fleishmanEffect] explores how upper-class, New York women are responding to the HBO show *Fleishman Is In Trouble*, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
+When held against some objective measure of financial well-being, being part of a group that gifts a child with designer jewelry suggests an income and lifestyle that would suggest financial satisfaction.
+"The Fleishman Effect" occurs when these higher group standards push satisfaction out of reach.
 
@@
 
+Financial satisfaction is dependent on an individual's context, and much more subjective than a fixed measure like individual or household income.
+We hypothesize that this context may be determined not only by their current reference group, but also by their relationship to wealth, and how this relationship evolved over the course of their life.
 The study would be strengthened with data that could provide a fuller profile of respondent's reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
 
-For instance, geography can have a major influence on personal finance and reference groups.
-A respondent making \$75,000 a year living in Hattiesburg, Mississippi might feel more financially satisfied than a respondent with the same income living in San Francisco, California.  However, this could sacrifice anonymity.
-This question would still have to be broad enough to provide information on their location (i.e.
-by city), rather than something that could de-anonymize respondents when triangulated with other data (i.e.
-postal code).
-
-Another key factor in how people interpret their own wealth is the way that they narrativize their success in comparison to their upbringing.
+The way that individuals narrativize their success in comparison to their upbringing may influence the way that they interpret their financial standing.
 In our supplemental survey in Appendix X, we suggest the questions "Have you moved up a full socio-economic class compared to your upbringing?
@@
 We could also see how people across different income ranges and classes experience regret about their decisions.
+
+Geography can have a major influence on personal finance and reference groups.
+A respondent making \$75,000 a year living in Hattiesburg, Mississippi might feel more financially satisfied than a respondent with the same income living in San Francisco, California.  However, this could sacrifice anonymity.
+This question would still have to be broad enough to provide information on their location (i.e.
+by city), rather than something that could de-anonymize respondents when triangulated with other data (i.e.
+postal code).
+
 Please see [Appendix] to view the full supplementary survey that we have put together.
@@
 
-# Appendix
+# Appendix {.unnumbered}
 

### 2023-03-11 outputs/paper/paper3.qmd
updated future research directions

@@
 
+#### Supplementary Survey
+
 Financial satisfaction is dependent on an individual's context, and much more subjective than a fixed measure like individual or household income.
@@
 
+#### Cross-Referencing
+
+Connecting our results with US census data to check for representativeness in respondents answers could strengthen interpretations, and help to understand any discrepancies in self-reported data.
+This would also help us understand how the new methodology implemented by the GSS in 2021 may have influenced self-reported results.
+
+Bringing in other macroeconomic indicators, such as inflation and capital market trends, can also provide an extra dimension to understanding how wealth has evolved and been impacted over time.
+
 -   Additional financial variables in survey for analysis
-
--   Connect with US census data to check representativeness
-
--   Connect with other macroeconomic indicators - e.g. inflation, stock markets
 

### 2023-03-11 outputs/paper/paper3.qmd
title change hehe. we can change it back. just playing around

@@
 ---
-title: "Subjective Evaluations of Financial Satisfaction and Well-being: A Study of the US General Social Survey"
+title: "Investigating The Fleishman Effect : Subjective Evaluations of Financial Satisfaction in the US General Social Survey"
 title-block-banner: true

### 2023-03-12 scripts/00-simulate_data.R
Added data simulation + validation tests

@@ -1,14 +1,67 @@
 #### Preamble ####
-# Purpose: Simulates data [UPDATE ..]
-# Author: Michaela Drouillard & Christina Wei
-# Data: 21 February 2023
-# Contact: michaela.drouillard@mail.utoronto.ca & christina.wei@mail.utoronto.ca
+# Purpose: Simulates data for survey results regarding financial attitudes
+# Author: Christina Wei
+# Data: 12 March 2023
+# Contact: christina.wei@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
+# Assumptions:
+  # Assume there are survey results for each year from 1970 to 2022
+  # Assume there are 10 participant responses each year
+  # Assign financial satisfactions at random (equal weight to each answer)
+  # Assign social class at random (equal weight to each answer)
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+
+## Enumerations of data values & assumptions
+
+sim_financial_satisfaction = c("Not Satisfied", "Neutral", "Satisfied")
+sim_social_status = c("Lower Class", "Working Class", "Middle Class", "Upper Class")
+sim_gender = c("Male", "Female", "Other")
+sim_years = 1970:2022
+num_participants = 10
+total_size = num_participants * length(sim_years)
+
+## Simulation
+
+set.seed(311) #random seed
+
+simulated_data = 
+  tibble(
+    year = rep(sim_years, each = num_participants),
+    participant_id = rep(1:num_participants, times = length(sim_years)),
+    age = floor(runif(n = total_size,
+                min = 18,
+                max = 100)),
+    gender = sample(x = sim_gender,
+                    size = total_size,
+                    replace = TRUE),
+    financial_satisfaction = sample(x = sim_financial_satisfaction,
+                                    size = total_size,
+                                    replace = TRUE),
+    social_status = sample(x = sim_social_status,
+                           size = total_size,
+                           replace = TRUE),
+  )
+
+## View and test characteristics of simulated data
+
+mean(simulated_data$age)
+min(simulated_data$year) == 1970
+max(simulated_data$year) == 2022
+
+simulated_data |>
+  group_by(gender) |>
+  count()
+
+simulated_data |>
+  group_by(financial_satisfaction) |>
+  count()
+
+simulated_data |>
+  group_by(social_status) |>
+  count()

### 2023-03-12 scripts/02-data_cleaning.R

@@ -73,6 +73,7 @@ cleaned_respondent_info =
 cleaned_respondent_info =
   cleaned_respondent_info |>
   mutate("cohort_band" = case_when(
+    cohort == 9999 ~ NA,
     cohort <= 1924 ~ 'Greatest',
     cohort >= 1925 & cohort <= 1945 ~ 'Silent',
     cohort >= 1946 & cohort <= 1964 ~ 'Boomer',

### 2023-03-12 scripts/03-test_data.R

@@ -1,13 +1,186 @@
 #### Preamble ####
-# Purpose: Tests .. [UPDATE ..]
-# Author: Michaela Drouillard & Christina Wei
-# Data: 21 February 2023
-# Contact: michaela.drouillard@mail.utoronto.ca & christina.wei@mail.utoronto.ca
+# Purpose: Validate cleaned datasets for both respondent and survey data
+# Author: Christina Wei
+# Data: 12 March 2023
+# Contact: christina.wei@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 01-download_data.R; 02-data_cleaning.R
+# Pre-requisites: 
+  # 01-download_data.R
+  # 02-data_cleaning.R
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Test data ####
+#### Read in cleaned data ####
+
+cleaned_respondent_info = read_csv(
+  file = here("outputs/data/cleaned_respondent_info.csv"),
+  show_col_types = FALSE
+)
+
+cleaned_finance_survey = read_csv(
+  file = here("outputs/data/cleaned_finance_survey.csv"),
+  show_col_types = FALSE
+)
+
+#### Data Validation ####
+
+# NA threshold - less than 10% of data
+threshold = 0.1
+num_NA_threshold = nrow(cleaned_respondent_info) * threshold 
+num_NA_threshold_2021 = nrow(filter(cleaned_respondent_info, year == 2021)) * threshold
+
+## Respondent info ##
+
+# year
+class(cleaned_respondent_info$year) == "numeric"
+min(cleaned_respondent_info$year) <= 1972
+max(cleaned_respondent_info$year) >= 2021
+sum(is.na(cleaned_respondent_info$year)) == 0 #no N/As for year
+
+# cohort
+class(cleaned_respondent_info$cohort) == "numeric"
+max(cleaned_respondent_info |>
+  select(cohort) |>
+  filter(cohort != 9999)) <= as.numeric(format(Sys.Date(), format = "%Y")) - 18
+min(cleaned_respondent_info$cohort, na.rm = TRUE) >= 1972 - 100
+sum(is.na(cleaned_respondent_info$cohort)) < num_NA_threshold
+
+# cohort_band
+sum(is.na(cleaned_respondent_info$cohort_band)) ==
+  sum(is.na(cleaned_respondent_info$cohort)) +
[…]

### 2023-03-13 outputs/paper/paper3.qmd
Addressing Rohan's feedback

@@
 
+In the remainder of the paper, we first discuss the data sources, limitations and cleaning procedures in the Data section.
+The following section, Results, present trends and correlations discovered in analysis, slicing across 2021 survey data as well as investigating results over the years.
+The Discussion section shares additional insights into our data findings such as the wealth gap and Fleishman Effect, as well as outlining future research directions.
+Finally, we wrap up with the Conclusion section summarizing the main discoveries from this paper.
+
 # Data
@@
 
-Specifically, we retrieved the following demographic data.
+Specifically, we retrieved the following demographic data as seen in @tbl-uss-demographic.
 
@@
 
-For survey questions related to financial wellness, we retrieved the following data points:
+: Demographic data retrieved from USS {#tbl-uss-demographic}
+
+For survey questions related to financial wellness, we retrieved the following data points as seen in @tbl-uss-financial
 
 | Variable | New Name               | Description                                                                             | Example               |
-|----------------|----------------|------------------------|----------------|
+|------|---------------|----------------------|---------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 | class1   | social_class1          | self identified social class the respondent belong in (more granular than social_class) | Upper Middle Class    |
+
+: Financial survey data retrieved from USS {#tbl-uss-financial}
 
@@
 #| label: tbl-statistics
-#| tbl-cap: Summary statistics for number of observations in surveys
+#| tbl-cap: Summary statistics for the number of observations across the 33 surveys from 1972 to 2021
+
 combined_data |>
@@
   summarise(
-    mean = mean(n),
+    mean = round(mean(n), digits = 0),
     max = max(n),
@@
 #| label: fig-2021-demographics
-#| fig-cap: "Distribution of Respondents' Demgraphics"
+#| fig-cap: "Distribution of Respondents' Demographics"
 #| fig-subcap: ["Age","Marital Status", "Level of Education", "Family Income"]
@@
 #| layout-nrow: 2
-
-#Age
-# survey_data_2021 |>
-#   ggplot(mapping = aes(x = age)) +
-#   geom_histogram(bins = 20, color="black", fill="grey") + 
-#   theme_minimal() +
-#   labs(
-#     x = "Age",
-#     y = "Number of responses"
-#   )
-
-#source(here("scripts/04-helper_functions.R"))
 
@@
 It would be interesting to compare this to the demographic data across United States to see if this is a representative sample of the general population.
-There are also a large number of `NA` records for age responses, accounting for 8.3% of our data.
+There are 8.2% of data records that are `NA` for age responses.
 Based on GSS codebook, 5.6% respondents chose "Not Applicable" when asked about their age, while 2.7% of the participants did not provide any answers.
@@
 To analyze how respondents compare themselves to others, we look at @fig-2021-financial-3 on how they ranked themselves against American families in general.
-This graphs shows that most individuals feel like they are on par with others (39.5%), with small number of respondents feeling above average (24.8%) and below average (24.3%).
+This graph shows that most individuals feel like they are on par with others (39.5%), with small number of respondents feeling above average (24.8%) and below average (24.3%).
 However, 7.1% of respondent felt that they are far below average compared to other families in general, and 3.8% felt they are far above average.
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = social_class, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Social Class (4-Bands)",
-    y = "Percentage of Responses",
-    fill = "Financial Satisfaction"
-  )
-
+    ggplot(aes(x = social_class, y = pct, fill = financial_satisfaction)),
+  angle = 45,
+  hjust = 1,
+  xlabel = "Social Class (4-Bands)",
+  ylabel = "Percentage of Responses",
+  fill_label = "Financial Satisfaction"
+)
+
+generate_compare_bar_graph(
+  ggplot_data = 
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = social_class1, y = pct, fill = financial_satisfaction)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Social Class (6-Bands)",
-    y = "Percentage of Responses",
-    fill = "Financial Satisfaction"
+    ggplot(aes(x = social_class1, y = pct, fill = financial_satisfaction)),
+  angle = 45,
+  hjust = 1,
+  xlabel = "Social Class (6-Bands)",
+  ylabel = "Percentage of Responses",
+  fill_label = "Financial Satisfaction"
 )
@@
 #| fig-cap: "Financial comparison responses for each social class"
-#| fig-subcap: ["Social Class (4-Bands)", "Social Class (6-Bands)"]
-
+
+#fig-subcap: ["Social Class (4-Bands)", "Social Class (6-Bands)"]
+
+generate_compare_bar_graph(
+  ggplot_data = 
     survey_data_2021 |>
@@
     mutate(pct = percent(n / sum(n), digit = 1)) |>
-  ggplot(aes(x = social_class1, y = pct, fill = financial_compare)) +
-  geom_bar(stat="identity", position="dodge") +
-  scale_y_continuous(labels=scales::percent) +
-  theme_minimal() +
-  theme(legend.position="bottom") + 
-  labs(
-    x = "Social Class (6-Bands)",
-    y = "Percentage of Responses",
-    fill = "Financial Compare"
+    ggplot(aes(x = social_class1, y = pct, fill = financial_compare)),
+  angle = 45,
+  hjust = 1,
+  nrow = 2,
+  xlabel = "Social Class (6-Bands)",
+  ylabel = "Percentage of Responses",
+  fill_label = "Comparing to Others"
 )

### 2023-03-13 scripts/04-helper_functions.R

@@ -38,13 +38,15 @@ generate_heatmap = function(ggplot_data, xlabel = NULL, ylabel = NULL, position 
 }
 
 # Generate comparative bar graph
-generate_compare_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = NULL, fill_label = NULL, angle = NULL, hjust = NULL, vjust = NULL) {
+generate_compare_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = NULL, fill_label = NULL, angle = NULL, hjust = NULL, vjust = NULL, nrow = NULL) {
 
   ggplot_data +
   geom_bar(stat="identity", position="dodge") +
   scale_y_continuous(labels=scales::percent) +
   theme_minimal() +
+  theme(axis.text.x = element_text(angle = angle, vjust = vjust, hjust = hjust)) +  
   theme(legend.position="bottom") + 
+  guides(fill=guide_legend(nrow=nrow, byrow=TRUE)) +
   labs(
     x = xlabel,
     y = ylabel,

### 2023-03-13 outputs/paper/paper3.qmd
Minor formatting updates

@@
 #| label: fig-socialclass-compare
-#| fig-cap: "Financial comparison responses for each social class"
+#| fig-cap: "Financial comparison responses for each social class (6-bands)"
 
@@
   hjust = 1,
-  nrow = 2,
   xlabel = "Social Class (6-Bands)",
@@
   fill_label = "Comparing to Others"
-)
+) +
+theme(legend.position=NULL)
 
@@
 
-(fig-2021-degree-financial) shows that respondents with post-secondary education experience better changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is "getting better".
-Subsetting these reference groups by their area of study, and whether or not they regret their major, could provide more information on how the different returns-on-investment on different degrees change over time, as shown by The Washington Post's reporting on the Federal Reserve Survey of Household Economics and Decisionmaking [@collegeMajors].
+@fig-2021-degree-financial shows that respondents with post-secondary education experience better changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is "getting better".
+Obtaining more details on these reference groups by their area of study, and whether or not they regret their major, could provide more information on how the different returns-on-investment on different degrees change over time, as shown by The Washington Post's reporting on the Federal Reserve Survey of Household Economics and Decision Making [@collegeMajors].
 
@@
 
-#### Cross-Referencing
-
-Connecting our results with US census data to check for representativeness in respondents answers could strengthen interpretations, and help to understand any discrepancies in self-reported data.
+#### Incorporate Additional Data Sources
+
+Additional data sources can be used to explore additional trends in attitudes towards financial wellness.
+First, we can connecting our results with US census data to check for representativeness in respondents answers, and to understand any discrepancies in self-reported data.
 This would also help us understand how the new methodology implemented by the GSS in 2021 may have influenced self-reported results.
-
-Bringing in other macroeconomic indicators, such as inflation and capital market trends, can also provide an extra dimension to understanding how wealth has evolved and been impacted over time.
-
--   Additional financial variables in survey for analysis
+Also, bringing in other macroeconomic indicators, such as inflation and capital market trends, can provide an extra dimension to understanding how wealth has evolved and been impacted over time.
 

### 2023-03-13 outputs/paper/financial_satsifaction_in_us.qmd
changing file name to be more meaningful

--- 2023-03-13 outputs/paper/paper3.qmd
@@
 
-Specifically, we retrieved the following demographic data as seen in @tbl-uss-demographic.
+Specifically, we retrieved the following demographic data described in @tbl-uss-demographic.
 
@@
 
-For survey questions related to financial wellness, we retrieved the following data points as seen in @tbl-uss-financial
+For survey questions related to financial wellness, we retrieved the following data points described in @tbl-uss-financial.
 

### 2023-03-13 outputs/paper/financial_satsifaction_in_us.qmd
data linting applied

@@
 The paper discusses the sociological dynamics and cognitive biases that affect these subjective evaluations, and highlights the importance of considering these factors in understanding financial satisfaction and well-being."
-thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3104-3-US-Survey.git"
+thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/Financial-Wellness-in-US.git"
 author: 

### 2023-03-13 scripts/00-simulate_data.R

@@ -19,18 +19,21 @@ library(tidyverse)
 
 ## Enumerations of data values & assumptions
 
-sim_financial_satisfaction = c("Not Satisfied", "Neutral", "Satisfied")
-sim_social_status = c("Lower Class", "Working Class", "Middle Class", "Upper Class")
-sim_gender = c("Male", "Female", "Other")
-sim_years = 1970:2022
-num_participants = 10
-total_size = num_participants * length(sim_years)
+sim_financial_satisfaction <- c("Not Satisfied", "Neutral", "Satisfied")
+sim_social_status <- c("Lower Class",
+                      "Working Class",
+                      "Middle Class",
+                      "Upper Class")
+sim_gender <- c("Male", "Female", "Other")
+sim_years <- 1970:2022
+num_participants <- 10
+total_size <- num_participants * length(sim_years)
 
 ## Simulation
 
 set.seed(311) #random seed
 
-simulated_data = 
+simulated_data <-
   tibble(
     year = rep(sim_years, each = num_participants),
     participant_id = rep(1:num_participants, times = length(sim_years)),

### 2023-03-13 scripts/01-download_data.R

@@ -15,38 +15,46 @@ library(haven) # for reading DTA files
 #### Download data ####
 
 # Download & unzip all the GSS data across the years
-zip_file = "inputs/data/large_files/GSS_stata.zip"
+zip_file <- "inputs/data/large_files/GSS_stata.zip"
 
 download.file("https://gss.norc.org/documents/stata/GSS_stata.zip", zip_file)
 unzip(zip_file, exdir = "inputs/data/large_files")
 
 # read dta data and write to csv
-raw_gss_data = read_dta("inputs/data/large_files/gss7221_r3a.dta")
+raw_gss_data <- read_dta("inputs/data/large_files/gss7221_r3a.dta")
 
 
-## Filter and save data related to respondent info and survey questions of interest
+## Filter and save data related to respondent info and survey questions
 
 # Respondent information
-raw_respondent_info =
+raw_respondent_info <-
   raw_gss_data |>
-  #select (year,id,wrkstat,hrs1,hrs2,evwork,occ,prestige,wrkslf,wrkgovt,commute,industry,occ80,prestg80,indus80,indus07,occonet,found,occ10,occindv,occstatus,occtag,prestg10,prestg105plus,indus10,indstatus,indtag,marital,martype,agewed,divorce,widowed,spwrksta,sphrs1,sphrs2,spevwork,cowrksta,cowrkslf,coevwork,cohrs1,cohrs2,spocc,sppres,spwrkslf,spind,spocc80,sppres80,spind80,spocc10,spoccindv,spoccstatus,spocctag,sppres10,sppres105plus,spind10,spindstatus,spindtag,coocc10,coind10,paocc16,papres16,pawrkslf,paind16,paocc80,papres80,paind80,paocc10,paoccindv,paoccstatus,paocctag,papres10,papres105plus,paind10,paindstatus,paindtag,maocc80,mapres80,mawrkslf,maind80,maocc10,maoccindv,maoccstatus,maocctag,mapres10,mapres105plus,maind10,maindstatus,maindtag,sibs,childs,age,agekdbrn,educ,paeduc,maeduc,speduc,coeduc,codeg,degree,padeg,madeg,spdeg,major1,major2,dipged,spdipged,codipged,cosector,whenhs,whencol,sector,eftotlt,barate,gradtounder,voedcol,voednme1,voedncol,voednme2,spsector,speftotlt,spbarate,spgradtounder,cobarate,cogradtounder,coeftotlt,sex,race,res16,reg16,mobile16,famil
[…]

### 2023-03-13 scripts/02-data_cleaning.R

@@ -1,5 +1,6 @@
 #### Preamble ####
-# Purpose: Cleans data for GSS related to respondent information and surveys on financial topics
+# Purpose: Cleans data for GSS related to respondent information
+         # and surveys on financial topics
 # Author: Michaela Drouillard & Christina Wei
 # Data: 21 February 2023
 # Contact: michaela.drouillard@mail.utoronto.ca & christina.wei@mail.utoronto.ca
@@ -11,18 +12,13 @@ library(tidyverse)
 
 #### Download data ####
 
-# unzip(
-#   "inputs/data/raw_respondent_info.zip",
-#   exdir = "inputs/data/large_files"
-# )
-
-raw_respondent_info =
+raw_respondent_info <-
   read_csv(
     file = "inputs/data/raw_respondent_info.csv",
     show_col_types = FALSE
   )
 
-raw_finance_survey =
+raw_finance_survey <-
   read_csv(
     file = "inputs/data/raw_finance_survey.csv",
     show_col_types = FALSE
@@ -30,220 +26,226 @@ raw_finance_survey =
 
 ## Update demographic data ##
 
-# [work_status] Last week were you working full time, part time, going to school, keeping house, or what?
-cleaned_respondent_info =
+# [work_status] Last week were you working full time, part time,
+# going to school, keeping house, or what?
+cleaned_respondent_info <-
   raw_respondent_info |>
   mutate("work_status" = case_when(
-    wrkstat == 1 ~ 'Full Time',
-    wrkstat == 2 ~ 'Part Time',
-    wrkstat == 3 ~ 'Temporarily Not Working',
-    wrkstat == 4 ~ 'Unemployed',
-    wrkstat == 5 ~ 'Retired',
-    wrkstat == 6 ~ 'In School',
-    wrkstat == 7 ~ 'Keeping House',
-    wrkstat == 8 ~ 'Other',
+    wrkstat == 1 ~ "Full Time",
+    wrkstat == 2 ~ "Part Time",
+    wrkstat == 3 ~ "Temporarily Not Working",
+    wrkstat == 4 ~ "Unemployed",
+    wrkstat == 5 ~ "Retired",
+    wrkstat == 6 ~ "In School",
+    wrkstat == 7 ~ "Keeping House",
+    wrkstat == 8 ~ "Other",
   )) |>
-  select (-wrkstat)
+  select(-wrkstat)
 
[…]

### 2023-03-13 scripts/03-test_data.R

@@ -4,7 +4,7 @@
 # Data: 12 March 2023
 # Contact: christina.wei@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
+# Pre-requisites:
   # 01-download_data.R
   # 02-data_cleaning.R
 
@@ -13,12 +13,12 @@ library(tidyverse)
 
 #### Read in cleaned data ####
 
-cleaned_respondent_info = read_csv(
+cleaned_respondent_info <- read_csv(
   file = here("outputs/data/cleaned_respondent_info.csv"),
   show_col_types = FALSE
 )
 
-cleaned_finance_survey = read_csv(
+cleaned_finance_survey <- read_csv(
   file = here("outputs/data/cleaned_finance_survey.csv"),
   show_col_types = FALSE
 )
@@ -26,9 +26,10 @@ cleaned_finance_survey = read_csv(
 #### Data Validation ####
 
 # NA threshold - less than 10% of data
-threshold = 0.1
-num_NA_threshold = nrow(cleaned_respondent_info) * threshold 
-num_NA_threshold_2021 = nrow(filter(cleaned_respondent_info, year == 2021)) * threshold
+threshold <- 0.1
+num_na_threshold <- nrow(cleaned_respondent_info) * threshold
+num_na_threshold_2021 <-
+  nrow(filter(cleaned_respondent_info, year == 2021)) * threshold
 
 ## Respondent info ##
 
@@ -44,7 +45,7 @@ max(cleaned_respondent_info |>
   select(cohort) |>
   filter(cohort != 9999)) <= as.numeric(format(Sys.Date(), format = "%Y")) - 18
 min(cleaned_respondent_info$cohort, na.rm = TRUE) >= 1972 - 100
-sum(is.na(cleaned_respondent_info$cohort)) < num_NA_threshold
+sum(is.na(cleaned_respondent_info$cohort)) < num_na_threshold
 
 # cohort_band
 sum(is.na(cleaned_respondent_info$cohort_band)) ==
@@ -63,10 +64,10 @@ sum(!(unique(cleaned_respondent_info$cohort_band) %in%
 class(cleaned_respondent_info$age) == "numeric"
 max(cleaned_respondent_info$age, na.rm = TRUE) <= 120
 min(cleaned_respondent_info$age, na.rm = TRUE) >= 18
-sum(is.na(cleaned_respondent_info$age)) < num_NA_threshold
[…]

### 2023-03-13 scripts/04-helper_functions.R

@@ -9,44 +9,75 @@
 library(tidyverse)
 
 # Generate percentage graph
-generate_percentage_bar_graph = function(ggplot_data, xlabel = NULL, ylabel = "Percentage of Responses", angle = NULL, hjust = NULL, vjust = NULL) {
+generate_percentage_bar_graph <- function(
+    ggplot_data,
+    xlabel = NULL,
+    ylabel = "Percentage of Responses",
+    angle = NULL,
+    hjust = NULL,
+    vjust = NULL
+) {
 
-  # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
-  ggplot_data + 
+  # Reference code from
+  # https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
+  ggplot_data +
   geom_bar(aes(y = (..count..) / sum(..count..))) +
-  scale_y_continuous(labels=scales::percent) +
+  scale_y_continuous(labels = scales::percent) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = angle, hjust = hjust, vjust = vjust)) + #rotate x-axis & smaller font
-  labs (
+  theme(
+    axis.text.x = element_text(angle = angle, hjust = hjust, vjust = vjust)
+  ) + #rotate x-axis & smaller font
+  labs(
       x = xlabel,
       y = ylabel
     )
 }
 
 # Generate heatmap
-generate_heatmap = function(ggplot_data, xlabel = NULL, ylabel = NULL, position = NULL, angle = 45, hjust = 1, vjust = 1) {
-  
-  ggplot_data + 
-  geom_tile() + 
-  scale_fill_gradient2(low="white", high="darkblue", guide="colorbar") +
-  theme(axis.text.x = element_text(angle = angle, vjust = vjust, hjust = hjust)) +
-  theme(legend.position = position) + 
-  labs (
+generate_heatmap <- function(
+    ggplot_data,
+    xlabel = NULL,
+    ylabel = NULL,
+    position = NULL,
+    angle = 45,
+    hjust = 1,
+    vjust = 1
+  ) {
+
+  ggplot_data +
+  geom_tile() +
+  scale_fill_gradient2(low = "white", high = "darkblue", guide = "colorbar") +
+  theme(
+    axis.text.x = element_text(angle = angle, vjust = vjust, hjust = hjust)
+  ) +
+  theme(legend.position = position) +
+  labs(
     x = xlabel,
[…]

### 2023-03-14 outputs/paper/financial_satsifaction_in_us.qmd
cleaned up abstract

@@
 title-block-banner: true
-abstract: "This paper examines the subjective evaluations of financial satisfaction and well-being, and how they are influenced by various demographic factors.
-Using data from the US General Social Survey, we explore how people understand their financial situations in comparison to others, according to different demographic factors, against the backdrop of an increasing wealth gap between high and low earning individuals.
+abstract: "Using data from the US General Social Survey, we explore how people understand their financial situations in comparison to others, according to different demographic factors, against the backdrop of an increasing wealth gap between high and low earning individuals.
 The paper discusses the sociological dynamics and cognitive biases that affect these subjective evaluations, and highlights the importance of considering these factors in understanding financial satisfaction and well-being."

### 2023-03-14 outputs/paper/financial_satsifaction_in_us.qmd
added a reference, edited

@@
 
-Financial satisfaction and well-being are subjective self-evaluations that can be influenced by a wide variety of different factors.
-However, the COVID-19 pandemic and the ensuing inflation rates have shown that pressures that economic trends bring on affect different demographics disproportionately.
-While the pandemic was a shared global experience, individual experiences of its implications were not.
-
+Financial satisfaction is a subjective self-evaluation that can be influenced by a wide variety of different factors.
 In this paper, we use data from the US General Social Survey (GSS) from NORC [@gss] at the University of Chicago to better understand how people understand their own financial situations in comparison to others, against the background an increasing wealth gap between higher and lower earning individuals.
-We subset and visualize the data by a variety of different demographic factors, such as gender, age, level of education, along with views of how different generational cohorts report their financial well-being over time.
-
-Overall, the data supports general hypotheses that over time, the rich are getting richer, the poor are getting poorer, and that Millennials and Gen Z are financial worse off than preceding generations.
-When faceting by social class, the data provides evidence that individuals understandings of their financial well-being and satisfaction are influenced by their reference groups.
+We conduct an analysis of the data by taking into account a variety of different demographic factors, including gender, age, and level of education.
+We also compare how different generational cohorts report their financial well-being over time.
+
+Overall, the data supports general hypotheses that over time, the rich report that they're getting richer, the poor report that they're getting poorer, and Millennials and Gen Z are financially worse off than preceding generations.
+When faceting by social class, the data provides evidence that individuals' understandings of their financial well-being and satisfaction are influenced by their reference groups.
 A significant number of respondents who self-reported as upper class still reported dissatisfaction, and described their financial situation as being "below average".
@@
 Due to the COVID-19 pandemic, the GSS had to change their data collection from in-person to address-based sampling and a push-to-web methodology, meaning that most of the interview was conducted with an online self-administered questionnaire.
-When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey.
+When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey (@gssMethods).
 
@@
 Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
-Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
+Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts (@gssMethods).
 
@@
 The 2021 GSS also conducted an experiment with volunteered responses.
-In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately.
+In the past, respondents frequently volunteered information (for instance, "just right" rather than "too harsh" or "not harsh enough"), which interviewers had training for recording appropriately (@gssMethods).
 In the online format, the GSS included a volunteered response on the screen in some cases, where the variable is marked with a -V suffix, and removed the option entirely for other respondents, marking the variable with an -NV suffix.
@@
 The last income band "\>\$170K" groups all respondents making over \$170,000 into one bucket.
-It reduces the insights we can glean from the data, for example, how much more than \$170,000 are there respondents making per year.
+It reduces the insights we can glean from the data (for example, how much more than \$170,000 are the respondents making per year).
 
@@
 
--   **Financial Satisfaction** - We are interested in how people are getting along financially these days.
-So far as you and your family are concerned, would you say that you are pretty well satisfied with your present financial situation, more or less satisfied, or not satisfied at all?
--   **Comparing to Others** - Compared with American families in general, would you say your family income is far below average, below average, average, above average, or far above average?
--   **Change in Financial Situation** - During the last few years, has your financial situation been getting better, worse, or has it stayed the same?
--   **Self-Identified Social Class** - Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?
+-   **Financial Satisfaction** - "We are interested in how people are getting along financially these days.
+So far as you and your family are concerned, would you say that you are pretty well satisfied with your present financial situation, more or less satisfied, or not satisfied at all?"
+-   **Comparing to Others** - "Compared with American families in general, would you say your family income is far below average, below average, average, above average, or far above average?"
+-   **Change in Financial Situation** - "During the last few years, has your financial situation been getting better, worse, or has it stayed the same?"
+-   **Self-Identified Social Class** - "Most people see themselves as belonging to a particular class. Please tell me which social class you would say you belong to?"
 
@@
 
-Based on @fig-2021-financial-1, it looks like our society is not doing too badly with their current financial situation.
-Only 23.9% reporting "not satisfied at all" on the question asking whether they are satisfied with their financial situation.
+Based on @fig-2021-financial-1, most respondents are more or less satisfied with their current financial situation, or very satisfied with their financial situation.
+23.9% of respondents reported being "not satisfied at all" on the question asking whether they are satisfied with their financial situation.
 Similarly, most respondents reported that they are staying the same or getting better in their financial situation, with only 20.1% feeling they are getting worse financially (@fig-2021-financial-2).
 
-To analyze how respondents compare themselves to others, we look at @fig-2021-financial-3 on how they ranked themselves against American families in general.
+To analyze how respondents compare themselves to others, we look at @fig-2021-financial-3 to see how they ranked themselves against American families in general.
 This graph shows that most individuals feel like they are on par with others (39.5%), with small number of respondents feeling above average (24.8%) and below average (24.3%).
-However, 7.1% of respondent felt that they are far below average compared to other families in general, and 3.8% felt they are far above average.
+However, 7.1% of respondents felt that they are far below average compared to other families in general, and 3.8% felt they are far above average.
 
@@
 Digging further into the relationship between self identified social class and how respondents compare their financial situation with others, we visualized their correlation using @fig-2021-class-compare.
-There is a clear trend shown in this heatmap: as individuals rank themselves higher in social class, they also perceive their financial situation to be better than the general public.
+There is a clear trend shown in this heat map: as individuals rank themselves higher in social class, they also perceive their financial situation to be better than the general public.
 
@@
 Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
-When comparing their financial situations to others, more proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
+When comparing their financial situations to others, a greater proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
 These results triangulated together shows that women in our society may be more disadvantaged in their financial wellness as compared to men.
@@
 The 2021 GSS data shows that having a higher level of education is correlated with higher financial satisfaction.
-in @fig-2021-degree-financial-1 there is a trend of decreasing percentages of answers for "not satisfied at all" as level of education increases, as well as an increasing percentages of responses for "pretty well satisfied" as the level of education increases.
+In @fig-2021-degree-financial-1, there is a trend of decreasing percentages of answers for "not satisfied at all" as level of education increases, as well as an increasing percentages of responses for "pretty well satisfied" as the level of education increases.
 This finding is supported by the survey results for respondents' perceptions on the changes in the financial situation (@fig-2021-degree-financial-1).
@@
 This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation.
-On the flip side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
-There is also the general trend that as the family income increases, the percentage of being satisfied with their current situation also increases.
+Unsurprisingly, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
+There is a general trend that as the family income increases, the percentages of satisfied responses also increases.
 Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).
@@
 
-In @fig-overtime-1, responses for financial satisfaction are graphed between 1972 to 2021.
+In @fig-overtime-1, we visualize the responses for financial satisfaction between 1972 to 2021.
 Based on this graph, we noticed that the percentage of respondents feeling "more or less satisfied" with their financial situations stays flat over time at around 44%.
@@
 The percentage of the other responses all have a similarly slightly positive upward trend.
-This graph shows indications that financial satisfaction has become more divided in society between the rich and the poor.
+This graph provides evidence that financial satisfaction has become more divided in the general population between higher and lower earners.
 
@@
 
-Next we looked at how each generation cohort is doing over time.
+Next, we looked at how each generation cohort is doing over time.
 We defined the generations based on the commonly accepted guidelines [@generations].
@@
 -   **The Silent Generation (Silent)**, defined as those born between 1925 and 1945, are between the ages 27 and 47 when the survey started to collect data in 1972.
-They were more dissatisfied with their financial situation to start with, but has increased their financial satisfaction over time, with close to 50% responded with "pretty well satisfied" as of 2021.
+They were more dissatisfied with their financial situation at the beginning, but have experienced an increase in their financial satisfaction over time, with close to 50% responded with "pretty well satisfied" as of 2021.
 
@@
 They were dissatisfied with their financial situation at the beginning, with more than 30% responding that they are "not satisfied at all".
-We have seen improvements over time, with increasing percentages of "pretty well satisified" responses and decreasing percentages of "not satisfied at all" responses.
+We have seen improvements over time, with increasing percentages of "pretty well satisfied" responses and decreasing percentages of "not satisfied at all" responses.
 
 -   **Generation X (Gen X)**, defined as those born between 1965 and 1979, show up for the first time in survey data in 1983 when the eldest in this generation turned 18.
-This group has a high percentage of respondents dissatisfied with their financial situation (\~30%), while about 25% of them are pertty well satisfied.
+This group has a high percentage of respondents dissatisfied with their financial situation (\~30%), while about 25% of them are pretty well satisfied.
 Their attitudes towards financial satisfaction stay pretty much the same throughout the years.
@@
 This group starts with the highest satisfaction with their financial situation (\~45%), but then it drastically declines to only \~20% responding with "pretty well satisfied" in 2021.
+This correlates with the COVID-19 pandemic.
 
@@
 
-Let's look specifically at the survey results for social class and how it correlates with the other variables.
 There are two survey questions related to how respondents selected the social class they belonged to: one used a classification of 4 classes (lower class, working class, middle class, and upper class), and a second survey question has an expanded 6 classes, with the additional classes of `lower middle class` and `upper middle class`.

### 2023-03-14 outputs/paper/financial_satsifaction_in_us.qmd
more lil edits

@@
 An interesting observation in this social class is that more than 10% of respondents making less than \$40K a year also put themselves as upper class.
+While this could be due to respondents misinterpreting the question, it also could be that respondents earning less than \$40k also have assets like trust funds, inheritances, property, or allowances that would place them in the "upper class".
 
@@
 Comparing the distribution between the two social class variables (@fig-socialclass), there are some movements across the bands; for example, there are less respondents who rated themselves as lower class and upper class.
-Looking at the income levels associated with **upper class** in this 6-band classifcation, there are some peculiar observations.
-While most individuals in this social class responded with making over \$170K annually, 20% of them reported family income less than \$10K.
+Looking at the income levels associated with **upper class** in this 6-band classification, there are some interesting observations: while most individuals in this social class responded with making over \$170K annually, 20% of them reported family income less than \$10K.
+Again, this could be due to respondents misinterpreting the question, or respondents who can rely on financial assets other than income, such as trust funds.
 
@@
 In @fig-socialclass-compare, only 60% of upper-class respondents see themselves as far above average compared to the general population.
-We will hypothesize potential explanations for this observation in the the discussion section below on \[The Fleishman Effect\].
+We will expand on potential explanations for this phenomenon in the the discussion section.
 

### 2023-03-14 outputs/paper/financial_satsifaction_in_us.qmd
minor fixes

@@
 ---
-title: "Investigating The Fleishman Effect : Subjective Evaluations of Financial Satisfaction in the US General Social Survey"
+title: "Investigating The Fleishman Effect: Subjective Evaluations of Financial Satisfaction in the US General Social Survey"
 title-block-banner: true
@@
 
-Financial satisfaction is a subjective self-evaluation that can be influenced by a wide variety of different factors.
-In this paper, we use data from the US General Social Survey (GSS) from NORC [@gss] at the University of Chicago to better understand how people understand their own financial situations in comparison to others, against the background an increasing wealth gap between higher and lower earning individuals.
-We conduct an analysis of the data by taking into account a variety of different demographic factors, including gender, age, and level of education.
-We also compare how different generational cohorts report their financial well-being over time.
-
-Overall, the data supports general hypotheses that over time, the rich report that they're getting richer, the poor report that they're getting poorer, and Millennials and Gen Z are financially worse off than preceding generations.
-When faceting by social class, the data provides evidence that individuals' understandings of their financial well-being and satisfaction are influenced by their reference groups.
-A significant number of respondents who self-reported as upper class still reported dissatisfaction, and described their financial situation as being "below average".
+Financial decisions are among the most complex and consequential decisions we make in our lives.
+Recent survey shows that the majority of Americans are feeling anxious about their financial situation, and perceive their lives being controlled by their finances [@capitalone2020].
+Financial satisfaction and well-being assessments are subjective self-evaluations that can be influenced by a wide variety of different factors.
+However, the COVID-19 pandemic and the ensuing inflation rate increases have shown that economic pressures affect different demographics disproportionately [@statCan].
+While the pandemic was a shared global experience, individual experiences of its implications were not.
+
+Consumer Financial Protection Bureau's report on Financial Well-Being in America [@cfpb] outlined the slight increase in the average score for financial well-being in 2020 as compared to 20217.
+It also highlighted the long-standing differences in financial well-being among different groups, such as age, gender and ethnicity.
+To further understand factors influencing individual's financial well-being, we used data from the US General Social Survey (GSS) from NORC [@gss] at the University of Chicago to deconstruct financial well-being into four different perceptions: overall financial satisfaction, perceived change in financial situation, perceived family income compared to other American families, and self identified social class.
+These perceptions are then analyzed to understand their correlations demographic factors (e.g.
+gender, age, level of education), as well as how these perceptions change over time.
+
+Overall, our analysis demonstrated the trend of wealth gap widening over time, as the rich are getting richer, and the poor are getting poorer.
+Also, different generations have different evaluations of their financial situations, with Millennials and Generation Z evaluating themselves as lower in financial well-being compared to preceding generations.
+When faceting by social class, the data provides evidence that individuals understandings of their financial well-being and satisfaction are influenced by their reference groups.
+A significant number of respondents who self-reported as upper class still expressed financial dissatisfaction, described their financial situation as being "below average".
 In our discussion, we explore the different ways in which sociological dynamics and cognitive biases can affect how someone might understand their own financial situation, along with other factors specific to the survey's limitations that shape this statistic.
@@
 
-Specifically, we retrieved the following demographic data described in @tbl-uss-demographic.
+For demographic factors, we retrieved the following data as described in @tbl-uss-demographic.
 
@@
 
-For survey questions related to financial wellness, we retrieved the following data points described in @tbl-uss-financial.
+For survey questions related to financial well-being, we retrieved the following data as described in @tbl-uss-financial.
 
@@
 
-Since 1972, the General Social Survey (GSS) has been tracking trends in public opinion through in-person interviews.
+Since 1972, the GSS has been tracking trends in public opinion through in-person interviews.
 Due to the COVID-19 pandemic, the GSS had to change their data collection from in-person to address-based sampling and a push-to-web methodology, meaning that most of the interview was conducted with an online self-administered questionnaire.
-When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey (@gssMethods).
+When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey.
+This change in methodology resulted in several limitations, which are discussed in detail below.
 
@@
 
-This change in data collection methodology has limitations.
 First, data from the year 2020 is not recorded, which is inconsistent with the yearly GSS track record.
 Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
-Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts (@gssMethods).
+Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
 
@@
 Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
-The 2021 GSS had no interviewers in the web mode to appropriately record these responses, so, in order to mitigate this change in mode, no item on the GSS shows "No Answer", unless the module sponsor requested that it be included.
+The 2021 GSS had no interviewers in the web mode to appropriately record these responses.
+In order to mitigate this change in mode, no item on the 2021 GSS shows "No Answer", unless the module sponsor requested that it be included.
 The exception to this was factual questions about occupation, income, or family background, since respondents could genuinely not know the answer, rather than being indecisive on the answer.
-Instead of "No answer" or "Don't know", users could skip the question.
+Instead of "No Answer" or "Don't know", users could skip the question.
 "Skipped on web" indicates that users read the questions but skipped it.
@@
 
-Any changes in opinions observed in 2021 relative to historical trends may be due to the changes in methodology.
+Any changes in opinions observed in 2021 relative to historical trends may be impacted by the changes in methodology.
 While the incomes reported are from 2019, the emotions recorded (i.e.
-financial satisfaction, financial comparison), are from 2021, after a year when many people lost their jobs, moved cities, or had other changes in household arrangements and income.
+financial satisfaction, comparing financial situation with others), are from 2021, after a year when many people lost their jobs, moved cities, or had other changes in household arrangements and income.
 While this has less of an impact on long-term visualizations of reported satisfaction over time, analyses of reported satisfaction based on income levels using 2021 data will be influenced by this gap.
@@
 
-After downloading and filtering for the desired source data values from GSS[^1], we performed data cleaning on each one of the columns based on value definitions as defined in GSS codebook [@gssCodebook], as well as cleaning up column names.
+After downloading and filtering for the selected data variables from GSS[^1], we performed data cleaning on each one of the columns based on value definitions as defined in GSS codebook [@gssCodebook], as well as renaming each column with meaningful names.
 For example, the column `sexbirth1` has been renamed to `gender`, with the value 1 updated to "Male", and the value 2 updated to "Female" based on mapping in the codebook.
@@
 Overall, there are 68,846 observations collected across 33 surveys administered over the years.
-On average there are 2086 observations per survey, with the maximum number of 4510 observations in year 2006, and minimum number of 1372 observations in year 1990.
-The standard deviation for the number of observations per survey is 788 (see @tbl-statistics).
+On average there are 2086 observations per survey, with the largest number of observations in the 2006 survey with 4510 responses, and smallest number of observations in 1990 survey with 1372 responses.
+The standard deviation for the number of observations per survey is 789 (see @tbl-statistics).
 
@@
     min = min(n),
-    sd = sd(n)
+    sd = round(sd(n), digits = 0)
   ) |>
@@
 In 2021, there are 4032 responses recorded for the US General Social Survey.
-Out of these responses, 1730 (42.9%) of the respondents identified as male, 2198 (54.5%) identified as female, with 104 (2.6%) of them did not identify their **gender** in the survey (@tbl-2021-gender).
+Out of these responses, 1730 (42.9%) of the respondents identified as male at birth, 2198 (54.5%) identified as female at birth, with 104 (2.6%) of them did not identify their **gender** in the survey (@tbl-2021-gender).
 Other than gender, we also looked at four different dimensions of respondents' demographics: age, marital status, degree of education and family income.
@@
 The average **age** of respondents who responded in 2021 is 52 years old.
-Looking at the age distribution in @fig-2021-demographics-1, we see a good spread of values across the different categories, but with noticeably smaller number of respondents less than 29 years old, and those who are over 80 years old.
-It would be interesting to compare this to the demographic data across United States to see if this is a representative sample of the general population.
+Looking at the age distribution in @fig-2021-demographics-1, we see a spread of values across the different categories, with noticeably smaller number of respondents less than 29 years old, as well as those who are over 80 years old.
+It would be interesting to compare this to the US demographic data to determine if the 2021 USS survey respondents are representative of the general population.
 There are 8.2% of data records that are `NA` for age responses.
@@
 
-Most of the respondent provided an answer to their **marital status** (@fig-2021-demographics-2).
-About half (49.6%) of the respondents are married as of the time of completing the survey, followed by 2.41% who were never married.
-The rest of the respondents selected either divorced (16.2%), widowed (7.5%), or separated (2.4%).
+Most (99.8%) of the respondent provided an answer to their **marital status** (@fig-2021-demographics-2).
+About half (49.6%) of the respondents are married as of the time of completing the survey, followed by 24.1% who were never married.
+The rest of the respondents selected either divorced (16.2%), widowed (7.5%), or separated (2.4%) as their marital status.
 
@@
 
-Looking at the respondents' **family income** as of 2016 (@fig-2021-demographics-4), the first thing that jumps out is the number of `NA` values, accounting for 12.9% of the data.
+Looking at the respondents' **family income** as of 2016 (@fig-2021-demographics-4), there is a significant number of `NA` values, accounting for 12.9% of the data.
 Based on GSS codebook, 8.3% respondents chose "Don't Know" for their income, 2.0% refused to answer the question, and 2.7% skipped the question all together.
-Otherwise, there seems to be a good spread of the number of respondents in each of the income bands, with half of the respondents reporting family income less than \$75,000.
+For those that answered the question, there is a spread across the income bands, with half of the respondents reporting family income less than \$75,000.
 The last income band "\>\$170K" groups all respondents making over \$170,000 into one bucket.
-It reduces the insights we can glean from the data (for example, how much more than \$170,000 are the respondents making per year).
+It reduces the insights we can extrapolate from this data, for example, the maximum differences between family incomes of participants.
 
@@
 
-As estimands to assess societal attitudes towards financial wellness in United States, we looked at the results of following survey questions:
+As estimands to assess societal attitudes towards financial well-being in United States, we looked at the results of following survey questions:
 
@@
 
-Based on @fig-2021-financial-1, most respondents are more or less satisfied with their current financial situation, or very satisfied with their financial situation.
-23.9% of respondents reported being "not satisfied at all" on the question asking whether they are satisfied with their financial situation.
-Similarly, most respondents reported that they are staying the same or getting better in their financial situation, with only 20.1% feeling they are getting worse financially (@fig-2021-financial-2).
+Based on @fig-2021-financial-1, most respondents selected "stayed the same" or "getting better" for their satisfaction with the current financial situation, with only 23.9% reporting "not satisfied at all".
+Similarly for changes in financial situation, most respondents reported that they are staying the same or getting better in their financial situation, with only 20.1% feeling they are getting worse financially (@fig-2021-financial-2).
 
@@
 
-Now that we have a general understanding of the respondent demographics and their attitudes towards financial wellness, let's take a look at the intersections between the attitudes and the demographics.
+Now that we have a general understanding of the respondent demographics and their attitudes towards financial well-being, let's take a look at the intersections between their attitudes and demographic factors.
 
@@
 Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
-When comparing their financial situations to others, a greater proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
-These results triangulated together shows that women in our society may be more disadvantaged in their financial wellness as compared to men.
+When comparing their financial situations to others, a higher proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
+These results triangulated together indicate that women in our society may be more disadvantaged in their financial well-being as compared to men.
 
@@
 
-Are we achieving better financial wellness as we get older?
+Are we achieving better financial well-being as we get older?
 We found evidence to support this in the 2021 GSS data, as @fig-2021-age-financial-1 demonstrates that the percentage of respondents who selected "pretty well satisfied" over time increases as age increases.
@@
   ylabel = "Percentage of Responses",
-  fill_label = "Financial Satisfaction"  
+  fill_label = "Financial Satisfaction",
+  nrow = 2
 )
@@
   ylabel = "Percentage of Responses",
-  fill_label = "Change in Financial Situation" 
+  fill_label = "Change in Financial Situation",
+  nrow = 2
 )
@@
 In @fig-2021-degree-financial-1, there is a trend of decreasing percentages of answers for "not satisfied at all" as level of education increases, as well as an increasing percentages of responses for "pretty well satisfied" as the level of education increases.
-This finding is supported by the survey results for respondents' perceptions on the changes in the financial situation (@fig-2021-degree-financial-1).
+This finding is supported by the survey results for respondents' perceptions on the changes in their financial situation (@fig-2021-degree-financial-2).
 As the level of education increases, there is a noticeable increase in the percentages of answers for "getting better" in their financial situation.
@@
   ylabel = "Percentage of Responses",
-  fill_label = "Financial Satisfaction"
+  fill_label = "Financial Satisfaction",
+  angle = 45,
+  hjust = 1,
+  nrow = 2
 )
@@
   ylabel = "Percentage of Responses",
-  fill_label = "Change in Financial Situation"
+  fill_label = "Change in Financial Situation",
+  angle = 45,
+  hjust = 1,
+  nrow = 2
 )
@@
 
-As a hypothesis, we would expect attitudes towards financial wellness is highly correlated with respondents' financial income.
+Overall, we would expect attitudes towards financial well-being is highly correlated with respondents' financial income.
 This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation.
-Unsurprisingly, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
-There is a general trend that as the family income increases, the percentages of satisfied responses also increases.
+On the other side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
+There is also the general trend that as the family income increases, respondents perceived themselves as doing better financial as compared to others.
 Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).
@@
 
-After investigating different demographic factors that are correlated to respondents' attitudes towards financial wellness, we want to understand how these attitudes evolve over time.
+After investigating different demographic factors that are correlated to respondents' attitudes towards financial well-being, let's analyze how these attitudes change over time.
 
@@
 #| fig-cap: "Attitude towards financial wellness over the years"
-#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
+#| fig-subcap: ["Financial Satisfaction", "Comparing Financial Situation to Others"]
 #| layout-ncol: 2
@@
   ylabel = "Percentage of Responses",
-  color_label = "Comparing Financial Situation to Others"
+  color_label = "Comparing to Others",
+  nrow = 3
   )
@@
 
-Next, we looked at how each generation cohort is doing over time.
+Next we looked at how each generation cohort is doing over time.
 We defined the generations based on the commonly accepted guidelines [@generations].
-Based on @fig-overtime-cohort we found the following trends:
+The trends for each generation cohort are discussed below (@fig-overtime-cohort).
 
 -   **The Greatest Generation (Greatest)**, defined as those born between 1910 and 1924, are between the ages 48 and 64 when the survey started to collect data in 1972.
-They are generally getting more satisfied with their financial situation over time, as shown by increasing "pretty well satisfied" and decreasing "not satisfied at all" response percentages.
+They are generally getting more satisfied with their financial situation over time, as shown by the increasing "pretty well satisfied" and decreasing "not satisfied at all" response percentages.
 
@@
 This group has a high percentage of respondents dissatisfied with their financial situation (\~30%), while about 25% of them are pretty well satisfied.
-Their attitudes towards financial satisfaction stay pretty much the same throughout the years.
+Their attitudes towards financial satisfaction stayed at around the same level throughout the years.
 
 -   **Millennials,** defined as those born between 1980 and 1994, show up for the first time in survey data around year 2000.
-This group starts with a higher satisfaction with their financial situation (\~30%), but it decreases over the years with only \~20% responding with "pretty well satisfied" in 2021.
+This group started with a higher satisfaction with their financial situation (\~30%), but it decreases over the years with only \~20% responding with "pretty well satisfied" in 2021.
 
@@
 
-In addition to looking at cohort over time, we also want to know how respondents at the same age groups are perceiving their financial situation over time.
+In addition to investigating generation cohorts over time, we also want to understand how respondents in the same age groups are perceiving their financial situation over time.
 Looking at @fig-overtime-age, we see a discouraging picture where "not satisfied at all" with their financial situations is trending upwards for almost all age groups, while being "pretty well satisfied" is trending downwards.
@@
 
-There are two survey questions related to how respondents selected the social class they belonged to: one used a classification of 4 classes (lower class, working class, middle class, and upper class), and a second survey question has an expanded 6 classes, with the additional classes of `lower middle class` and `upper middle class`.
+This section discusses the survey results related to social class and how it correlates with other variables.
+There are two survey questions related to how respondents selected the social class they belonged to: one used a classification of 4 classes (`lower class`, `working class`, `middle class`, and `upper class`), and a second survey question using an expanded classification of 6 classes, with the addition of `lower middle class` and `upper middle class`.
 
@@
 Comparing the distribution between the two social class variables (@fig-socialclass), there are some movements across the bands; for example, there are less respondents who rated themselves as lower class and upper class.
-Looking at the income levels associated with **upper class** in this 6-band classification, there are some interesting observations: while most individuals in this social class responded with making over \$170K annually, 20% of them reported family income less than \$10K.
+Looking at the income levels associated with **upper class** in this 6-band classification (@fig-income-social-2), there are some interesting observations: while most individuals in this social class responded with making over \$170K annually, 20% of them reported family income less than \$10K.
 Again, this could be due to respondents misinterpreting the question, or respondents who can rely on financial assets other than income, such as trust funds.
@@
 
-Looking at the correlation between social class and financial satisfaction, we saw the expected trend in @fig-socialclass-financial-1 that as individuals move up to a higher social class, they report higher financial satisfaction with their current situation.
+Analyzing the correlation between social class and financial satisfaction, we saw the expected trend in @fig-socialclass-financial-1 that as individuals move up to a higher social class, they report higher financial satisfaction with their current situation.
 However, when we look at the detailed social class breakdown in @fig-socialclass-financial-2, we noticed a trend reversal for **upper class**, where close to 20% of the respondents reported not satisfied at all with their financial situation, which is much higher compared to upper middle class with only less than 2% dissatisfaction.
 In @fig-socialclass-compare, only 60% of upper-class respondents see themselves as far above average compared to the general population.
-We will expand on potential explanations for this phenomenon in the the discussion section.
+We will hypothesize potential explanations for this observation in the the discussion section below.
 
@@
 
-Looking at the percentages of respondents in different social class over the years in @fig-socialclass-overtime, we discovered a declining trend for working class and middle class, while the percentage of those who reported as lower class has been increasing over time.
+As for the percentages of respondents in different social class over the years in @fig-socialclass-overtime, we discovered a declining trend for working class and middle class, while the percentage of those who reported as lower class has been increasing over time.
 Upper class percentage stays mainly constant from 1972 to 2021 at around 3-4%.

### 2023-03-14 outputs/paper/financial_satsifaction_in_us.qmd
Merging a few more changes

@@
 Due to the COVID-19 pandemic, the GSS had to change their data collection from in-person to address-based sampling and a push-to-web methodology, meaning that most of the interview was conducted with an online self-administered questionnaire.
-When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey.
+When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey [@gssMethods].
 This change in methodology resulted in several limitations, which are discussed in detail below.
@@
 Because the GSS data collection spans 2020 and 2021, the reference year for the 2021 GSS for questions where, for instance, respondents are asked for their income "last year", is 2019.
-Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts.
+Second, in the 2021 GSS, people below the age of 30, people without high school degrees, and Black respondents completed less surveys, proportionally, than their 2018 counterparts [@gssMethods].
 
@@
 Interviewers in previous rounds of the GSS had specific training in how to record "Don't Know" responses, which occurred when respondents either refused to answer a question or did not know the answer to a question.
-The 2021 GSS had no interviewers in the web mode to appropriately record these responses.
+The 2021 GSS had no interviewers in the web mode to appropriately record these responses [@gssMethods].
 In order to mitigate this change in mode, no item on the 2021 GSS shows "No Answer", unless the module sponsor requested that it be included.
@@
 Also, there are more female respondent reporting that their financial situation is getting worse (22.7%) compared to male respondents (16.7%) (@fig-2021-gender-financial-2).
-When comparing their financial situations to others, a higher proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
+When comparing their financial situations to others, a greater proportion of male respondents (30.2%) rated themselves as above average as compared to female respondents (21.3%) (@fig-2021-gender-financial-3).
 These results triangulated together indicate that women in our society may be more disadvantaged in their financial well-being as compared to men.
@@
 This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation.
-On the other side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
-There is also the general trend that as the family income increases, respondents perceived themselves as doing better financial as compared to others.
+Unsurprisingly, the majority of respondents making \$170K and above are pretty well satisfied with their current situation.
+There is a general trend that as the family income increases, the percentages of satisfied responses also increases.
 Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).
@@
 In @fig-socialclass-compare, only 60% of upper-class respondents see themselves as far above average compared to the general population.
-We will hypothesize potential explanations for this observation in the the discussion section below.
+We will expand on potential explanations for this phenomenon in the the discussion section.
 

### 2023-03-14 outputs/paper/financial_satsifaction_in_us.qmd
edited discussion

@@
 | Variable | New Name               | Description                                                                             | Example               |
-|----------------|----------------|------------------------|----------------|
+|-----------------|-----------------|----------------------|-----------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 Our results point to three trends: first, our visualizations illustrate a growing wealth gap over time.
-Second, the data suggests that financial well-being and satisfaction is influenced by demographic factors.
-And third, evidence to support the Fleishman effect \-- the idea that upper classes have a distorted sense of their financial well-being, since they compare themselves to the upper class, rather than comparing themselves to the general population.
-
-## Growing Wealth Gap Impacts Demographic Groups Differently
-
-Our analysis shows that the gap between high-earning and low-earning individuals in the US has been increasing since 1970, when the GSS was first recorded.
 Even though GDP is rising, we don't observe a significant improvement in financial satisfaction over time.
 In fact, twice as many respondents identify as "lower class" in 2021 than in 1970.
-
-Respondents' gender (defined in the GSS as sex at birth) and level of education have a significant influence on financial satisfaction.
+Second, the data suggests that financial well-being and satisfaction is influenced by demographic factors.
+And third, evidence to support the Fleishman Effect \-- a term that refers to self-identified upper class respondents who have a distorted sense of their financial well-being, since they compare themselves to other elites.
+
+## Demographic Factors' Influence on Financial Satisfaction
+
+According to the findings of our study, the impact of macro-economic trends like the widening wealth gap on individuals varies based on their demographic characteristics.
+Specifically, women and those with less formal education tend to report increased dissatisfaction over time.
 Men tend to see their financial satisfaction as improving, and report greater levels of satisfaction than women (@fig-2021-gender-financial).
 Respondents with Bachelors or Graduate degrees are more satisfied, and describe their financial situations as getting better over time (@fig-2021-degree-financial) .
-This suggests that the increased financial dissatisfaction reported over time in @fig-overtime-1 is more likely to be felt by non-male individuals with less formal education.
 
@@
 In @fig-overtime-cohort, the Greatest, Silent, and Boomer generations all see an increase in individuals reporting "Pretty Well Satisfied" over the course of their lives.
-In Gen X, this "Pretty Well Satisfied" line flattens, and Millennials and Gen Z see a significant decrease in high satisfaction (of course, Gen Z's visualization is correlated with a global pandemic).
+In Gen X, this "Pretty Well Satisfied" line flattens, and Millennials and Gen Z see a significant decrease in high satisfaction (of course, Gen Z's visualization is correlated with a global pandemic, and represents respondents at the beginning of their careers).
 
@@
 The way somebody feels about the financial situation might not reflect the reality of their financial situation as held against some other objective measure.
-In @fig-overtime, we see that reported financial situations are either getting better or getting worse over time, but that dissatisfaction overall is rising (with nearly 20% of upper-class respondents reporting dissatisfaction in @fig-socialclass-financial).
+For instance, nearly 20% of upper-class respondents report dissatisfaction in @fig-socialclass-financial.
+In @fig-overtime, we see that reported financial situations are either getting better or getting worse over time, but that dissatisfaction overall is rising.
 Although the levels of respondents identifying as upper and middle classes has remained relatively stable over time (@fig-socialclass-overtime), there is a general decline is people reporting being "Pretty Well Satisfied", and an increase in individuals reporting "Not Satisfied At All" (@fig-overtime).
-Who is feeling this dissatisfaction?
+Who is feeling all of this dissatisfaction?
 Why?
@@
 
-Our data on financial satisfaction based on income supports this hypothesis (@fig-socialclass-compare).
-When faceted by class instead, the question becomes more interesting.
-When asked to compare themselves to the general population, nearly 20% of self-identified upper-classes see themselves as "below average".
-While this may be due to problems with the survey recording, or respondents misunderstanding the question, the fact that only 60% of upper class individuals see themselves as being far above average suggests some sort of distortion of self-perception once that level of wealth is achieved.
+Our data on financial satisfaction based on income supports this hypothesis.
+When asked to compare themselves to the general population, nearly 20% of self-identified upper-classes see themselves as "below average" (@fig-socialclass-compare).
+While this may be due to problems with the survey recording, or respondents misinterpreting the question, the fact that only 60% of upper class individuals see themselves as being far above average suggests some sort of distortion of self-perception once that level of wealth is achieved.
 
@@
 Could survey respondents have misinterpreted what "upper class" means?
-Or, are there more factors at play when people consider their own financial situations \-- for instance, the child of millionaires who themselves make \$40k a year might see themselves as culturally and socially upper class, yet below average in terms of financial comparison.
-
-In her 2019 memoir We Need To Talk About Money [@talkAboutMoney], Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of struggling millennials.
+Or, are there more factors at play when people consider their own financial situations \-- for instance, the child of millionaires who themselves make \$40k a year might see themselves as culturally and socially upper class, yet below average in terms of income.
+
+In her 2019 memoir *We Need To Talk About Money* [@talkAboutMoney], Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of hustling milennials struggling to achieve the same checkpoints of financial success as previous generations.
 After learning that a friend had concealed that she, in fact, didn't have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that politically liberal high-earners (or inheritors) can experience as anti-capitalist sentiments have entered the mainstream conversations:
 
-"But that's the thing about shame -- it's an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to 'fit in'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
-For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited."
+> But that's the thing about shame -- it's an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to 'fit in'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
+For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited.
 
@@
 In @fig-overtime, we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z's self-reporting does correlate with a pandemic).
-This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions such as shame that distort people's understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
+This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions such as shame that may distort people's understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
 Millennials aren't experiencing the same financial satisfaction that Baby Boomers did when they were the same age.
 
-Meanwhile, the lure of upward mobility is addictive, as even the upper class reports financial dissatisfaction.
-An article by The Cut titled "The Fleishman Effect" [@fleishmanEffect] explores how upper-class, New York women are responding to the HBO show *Fleishman Is In Trouble*, which is about the stress that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
+Some upper-class individuals, specifically those belonging to generational cohorts experiencing financial distress, might therefore be persuade to perform and internalize a level of financial dissatisfaction that doesn't actually correspond to their wealth.
+On the other hand, financial dissatisfaction in the upper-classes can also come from a reference group effect, where their frame of comparison is no longer the average American.
+
+An article by The Cut titled "The Fleishman Effect" [@fleishmanEffect] explores how upper-class, New York women are responding to the FX television show *Fleishman Is In Trouble*, which is about, among other things, the pressures that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
 When held against some objective measure of financial well-being, being part of a group that gifts a child with designer jewelry suggests an income and lifestyle that would suggest financial satisfaction.
@@
 
+While the GSS data cannot provide a full enough profile of its respondents to determine the exact social and cultural conditions that their perceptions are shaped by, it does suggest that financial satisfaction is an individualized and subjective experience, and that demographic categories and class dynamics have an impact on responses.
+
 ## Future Research Directions
@@
 We hypothesize that this context may be determined not only by their current reference group, but also by their relationship to wealth, and how this relationship evolved over the course of their life.
-The study would be strengthened with data that could provide a fuller profile of respondent's reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
-
-The way that individuals narrativize their success in comparison to their upbringing may influence the way that they interpret their financial standing.
-In our supplemental survey in Appendix X, we suggest the questions "Have you moved up a full socio-economic class compared to your upbringing?
+The study would be strengthened by data that can provide a fuller profile of respondent's reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
+
+We hypothesize that the way that individuals narrativize their success in comparison to their upbringing may influence the way that they interpret their financial standing.
+In our supplemental survey in the Appendix, we suggest the questions "Have you moved up a full socio-economic class compared to your upbringing?
 (Y/N/Don't Know", "Do you have at least one parent who moved up a full socio-economic class?" (Y/N/Don't Know), and "Is your current career in the same field as your parents?" (Y/N/Don't Know) to better understand respondent's family histories.
@@
 
-@fig-2021-degree-financial shows that respondents with post-secondary education experience better changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is "getting better".
-Obtaining more details on these reference groups by their area of study, and whether or not they regret their major, could provide more information on how the different returns-on-investment on different degrees change over time, as shown by The Washington Post's reporting on the Federal Reserve Survey of Household Economics and Decision Making [@collegeMajors].
+@fig-2021-degree-financial shows that respondents with post-secondary education experience more positive changes in financial situation, with 50% of respondents with graduate degrees reporting that their financial situation is "getting better".
+Obtaining more detail on these respondents' areas of study, and whether or not they regret their major, could provide more information on how the returns-on-investment of different degrees change over time.
+This data can be cross-referenced with *The Washington Post*'s analysis on which majors are the most regretted, which draws on data from Federal Reserve Survey of Household Economics and Decision Making [@collegeMajors].
 
@@
 (Y/N/Don't know)" to the survey.
-Adding this question to our study would provide more detailed insight on reference groups, since we could see how people who enrolled in each major compare themselves to others, and, with more information on family history, we could also gather insight on which fields individuals from different social classes pursue in higher education.
+Adding this question to our study would provide more detailed insight on reference groups, since we could see how people who enrolled in each major compare themselves to others, and, with the additional information on their family's class history, we could also gather insight on which fields individuals from different social classes might pursue in higher education.
 We could also see how people across different income ranges and classes experience regret about their decisions.
@@
 Geography can have a major influence on personal finance and reference groups.
-A respondent making \$75,000 a year living in Hattiesburg, Mississippi might feel more financially satisfied than a respondent with the same income living in San Francisco, California.  However, this could sacrifice anonymity.
+A respondent making \$75,000 a year living in Hattiesburg, Mississippi might feel more financially satisfied than a respondent with the same income living in San Francisco, California.
+However, collecting data on respondents' location could sacrifice anonymity.
 This question would still have to be broad enough to provide information on their location (i.e.
@@
 
-In conclusion, financial satisfaction and well-being are subjective and influenced by various factors.
 This paper used data from the US General Social Survey to examine how people perceive their financial situations in comparison to others, particularly in light of the increasing wealth gap between high and low earning individuals.
-
-The results support the hypothesis that the rich are getting richer, the poor are getting poorer, and that younger generations are financially worse off than preceding cohorts.
-The data also suggests that people's reference groups influence their understanding of their financial well-being and satisfaction.
+We found that significant trends in financial satisfaction and comparison based on demographic categories such as generational cohort, level of education, age, gender, and social class.The results support the hypothesis that the rich are getting richer (or at least, feel like it), the poor are getting poorer (or at least, feel like it), and that younger generations are financially worse off than preceding cohorts.
+The data also suggests that people's reference groups influence their understanding of their financial well-being and satisfaction, with upper-classes experiencing counter-intuitively significant rates of financial dissatisfaction.
 Overall, this paper highlights the complexity of financial satisfaction and well-being, and the importance of considering sociological dynamics and cognitive biases in understanding these subjective evaluations.
+Further research with additional data on factors that shape a respondents' self-narrative, such as upbringing, family histories of social mobility, or the areas of study pursued in post-secondary education could add to our understandings of reference groups as they relate to wealth and perceived success.
 

### 2023-03-15 outputs/paper/financial_satsifaction_in_us.qmd
edited discussion again
Final Edits

@@
 | Variable | New Name               | Description                                                                             | Example               |
-|-----------------|-----------------|----------------------|-----------------|
+|-----------------|-----------------|---------------------|-----------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
@@
 
-Our results point to three trends: first, our visualizations illustrate a growing wealth gap over time.
+Our results point to three different trends.
+First, our visualizations illustrate a growing wealth gap over time.
 Even though GDP is rising, we don't observe a significant improvement in financial satisfaction over time.
-In fact, twice as many respondents identify as "lower class" in 2021 than in 1970.
-Second, the data suggests that financial well-being and satisfaction is influenced by demographic factors.
-And third, evidence to support the Fleishman Effect \-- a term that refers to self-identified upper class respondents who have a distorted sense of their financial well-being, since they compare themselves to other elites.
-
-## Demographic Factors' Influence on Financial Satisfaction
+Also, there is a rising trend in the percentage of respondents identifying as "lower class" from 1970 to 2021.
+Second, the data suggests that perceptions of financial well-being are different across demographic profiles, such as gender and levels of education.
+And third, we found evidence to support the Fleishman Effect \-- a term that refers to self-identified upper class respondents who have a distorted sense of their financial well-being, since they compare themselves to other elites.
+We hypothesize that a variety of different social, cultural, and personal pressures can influence how individuals perform their wealth or conform to wider trends of dissatisfaction.
+
+## Financial Satisfaction Differences Across Demographic Profiles
 
@@
 
-While satisfaction can often be correlated with income -- for instance, in @fig-2021-income-financial those who are making lower income are less satisfied-- it shouldn't be confused with income.
-The way somebody feels about the financial situation might not reflect the reality of their financial situation as held against some other objective measure.
-For instance, nearly 20% of upper-class respondents report dissatisfaction in @fig-socialclass-financial.
+While financial satisfaction is often correlated with income -- for instance, in @fig-2021-income-financial those who are making lower income are less satisfied -- they shouldn't be confused as the same concept.
+The way somebody feels about their financial well-being might not reflect the objective assessment of their financial wealth.
+For instance, nearly 20% of upper-class respondents report dissatisfaction in @fig-socialclass-financial-2.
 In @fig-overtime, we see that reported financial situations are either getting better or getting worse over time, but that dissatisfaction overall is rising.
@@
 Who is feeling all of this dissatisfaction?
-Why?
+And why?
 
@@
 
-Our data on financial satisfaction based on income supports this hypothesis.
-When asked to compare themselves to the general population, nearly 20% of self-identified upper-classes see themselves as "below average" (@fig-socialclass-compare).
-While this may be due to problems with the survey recording, or respondents misinterpreting the question, the fact that only 60% of upper class individuals see themselves as being far above average suggests some sort of distortion of self-perception once that level of wealth is achieved.
-
+Our analysis supports the hypothesis that individuals compare their financial situations to different reference groups.
+When asked to compare themselves to the general population, nearly 20% of self-identified upper class individuals see themselves as "below average" (@fig-socialclass-compare).
+While this may be due to problems with the survey recording, or respondents misinterpreting the question, the fact that only 60% of upper class individuals see themselves as being far above average suggests some sort of distortion of self-perception at that level of social class.
 Is the upper class simply delusional?
 Could survey respondents have misinterpreted what "upper class" means?
-Or, are there more factors at play when people consider their own financial situations \-- for instance, the child of millionaires who themselves make \$40k a year might see themselves as culturally and socially upper class, yet below average in terms of income.
-
-In her 2019 memoir *We Need To Talk About Money* [@talkAboutMoney], Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues were able to afford luxuries like property ownership while signalling, by complaining about mortgages and budgeting, that they were in the same cohort of hustling milennials struggling to achieve the same checkpoints of financial success as previous generations.
-After learning that a friend had concealed that she, in fact, didn't have a mortgage, and that her parents had given her enough money to buy her apartment outright, Uwagba writes about the cultural shame that politically liberal high-earners (or inheritors) can experience as anti-capitalist sentiments have entered the mainstream conversations:
-
-> But that's the thing about shame -- it's an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to 'fit in'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
-For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited.
+Or, are there more factors at play when people consider their own financial situations?
+For instance, the children of millionaires making \$40K a year might see themselves belonging culturally and socially to the upper class, but rating themselves below average for their financial situation.
+
+In her 2019 memoir *We Need To Talk About Money* [@talkAboutMoney], Otegha Uwagba writes about coming of age as a Black, Oxford-educated millennial woman from a low-income family, and how her perception of her own financial well-being was distorted by the way her upper class colleagues.
+Even though these colleagues were able to afford luxuries like property ownership, they complained about mortgages and budgeting, signalling that they were in the same cohort of hustling Millennials struggling to achieve checkpoints of financial success defined by previous generations.
+After learning that a friend had concealed that she didn't need a mortgage because her parents had given her enough money to buy her apartment, Uwagba writes about the cultural shame that politically liberal high-earners (or inheritors) can experience as anti-capitalist sentiments have entered the mainstream conversations:
+
+> *"But that's the thing about shame -- it's an emotion that arises from our most base human instincts: the fear of being judged negatively by others; our desire to 'fit in'; the discomfort of feeling exposed; and those emotional responses transcend income levels.
+For wealthy people who also happen to be politically liberal -- and so value the notion of social equity -- being rich, and the advantages that wealth presents can be a source of moral discomfort, particularly when said wealth is inherited."*
 
 Concealing wealth out of shame, to Uwagba, is "actively unhelpful" for others in the same reference group, and exacerbates the frustration of realizing that achieving personal finance goals may even more out of reach than initially believed.
-In @fig-overtime, we see that, beginning with Gen X, Gen Xers, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z's self-reporting does correlate with a pandemic).
-This complicates this idea of reference group comparisons \-- while intra-group comparisons are full of cognitive biases and emotions such as shame that may distort people's understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
+
+In @fig-overtime, we see that Gen X, Millennials, and Gen Z are experiencing more financial dissatisfaction over time compared to other generations (although, Gen Z's self-reporting does correlate with a pandemic).
+This complicates this idea of reference group comparisons: while intra-group comparisons are full of cognitive biases and emotions such as shame that may distort people's understandings of their own and their colleagues financial well-being, younger generations can no longer use their own age and educational achievements as a reference group.
 Millennials aren't experiencing the same financial satisfaction that Baby Boomers did when they were the same age.
-
-Some upper-class individuals, specifically those belonging to generational cohorts experiencing financial distress, might therefore be persuade to perform and internalize a level of financial dissatisfaction that doesn't actually correspond to their wealth.
+Some upper-class individuals, specifically those belonging to generational cohorts experiencing financial distress, might therefore be persuaded to perform and internalize a level of financial dissatisfaction that doesn't actually correspond to their wealth.
 On the other hand, financial dissatisfaction in the upper-classes can also come from a reference group effect, where their frame of comparison is no longer the average American.
@@
 An article by The Cut titled "The Fleishman Effect" [@fleishmanEffect] explores how upper-class, New York women are responding to the FX television show *Fleishman Is In Trouble*, which is about, among other things, the pressures that high-earning women feel when trying to achieve satisfaction in a city that houses one of the wealthiest reference groups in the world, where "the go-to bat mitzvah gift of the moment is a Cartier bracelet, for which moms are expected to pitch in for a group present".
-When held against some objective measure of financial well-being, being part of a group that gifts a child with designer jewelry suggests an income and lifestyle that would suggest financial satisfaction.
+When held against objective measures of financial wealth, being part of a group that gifts a child with designer jewelry suggests an income and lifestyle that would suggest financial satisfaction.
 "The Fleishman Effect" occurs when these higher group standards push satisfaction out of reach.
 
-While the GSS data cannot provide a full enough profile of its respondents to determine the exact social and cultural conditions that their perceptions are shaped by, it does suggest that financial satisfaction is an individualized and subjective experience, and that demographic categories and class dynamics have an impact on responses.
+While the GSS data does not provide expanded profiles of its respondents to determine the social and cultural conditions that their perceptions are shaped by, it does suggest that financial satisfaction is an individualized and subjective experience, and that demographic categories and class dynamics have an impact on responses.
 
@@
 
-#### Supplementary Survey
-
-Financial satisfaction is dependent on an individual's context, and much more subjective than a fixed measure like individual or household income.
-We hypothesize that this context may be determined not only by their current reference group, but also by their relationship to wealth, and how this relationship evolved over the course of their life.
+### Supplementary Survey
+
+Financial well-being is dependent on an individual's context, and much more subjective than quantitative measures like individual or household incomes.
+We hypothesize that this context may be determined not only by their current reference group, but also by their relationship to wealth, and how this relationship evolved over the course of their lives.
 The study would be strengthened by data that can provide a fuller profile of respondent's reference groups, and other factors that influence their motivations and attitudes towards personal finance and upward mobility.
@@
 We hypothesize that the way that individuals narrativize their success in comparison to their upbringing may influence the way that they interpret their financial standing.
-In our supplemental survey in the Appendix, we suggest the questions "Have you moved up a full socio-economic class compared to your upbringing?
-(Y/N/Don't Know", "Do you have at least one parent who moved up a full socio-economic class?" (Y/N/Don't Know), and "Is your current career in the same field as your parents?" (Y/N/Don't Know) to better understand respondent's family histories.
-Asking "Are your friends in the same socio-economic class, higher, or lower?
-(Higher/About the same/Lower)" would also provide more granular detail into how someone perceives their financial status in comparison to the people closest to them.
+In our supplemental survey in the Appendix, we suggest including questions to better understand respondent's family histories such as: "Have you moved up a full socio-economic class compared to your upbringing?"; "Do you have at least one parent who moved up a full socio-economic class?"; and "Is your current career in the same field as your parents?".
+Also, we would like to understand how individuals perceives their financial status in comparison to the people closest to them by asking questions such as "Are your friends in the same socio-economic class, higher, or lower?".
 The Social Capital Atlas [@socialCapital] conducted a survey pursuing a similar estimand using Facebook data to explore "economic connectedness" in friend groups, which refers to the percentage of friends a respondent has who earn higher or lower incomes than them.
@@
 This data can be cross-referenced with *The Washington Post*'s analysis on which majors are the most regretted, which draws on data from Federal Reserve Survey of Household Economics and Decision Making [@collegeMajors].
-
-We propose adding the questions "If you went to college or university, what was your major?" "Do you regret your major?
-(Y/N/Don't know)" to the survey.
-Adding this question to our study would provide more detailed insight on reference groups, since we could see how people who enrolled in each major compare themselves to others, and, with the additional information on their family's class history, we could also gather insight on which fields individuals from different social classes might pursue in higher education.
-We could also see how people across different income ranges and classes experience regret about their decisions.
+As such, we propose adding the questions like "If you went to college or university, what was your major?" and "Do you regret your major?" to the survey.
+
+These additional questions in our survey would provide more detailed insight on reference groups of the respondents.
+We can analyze how people enrolled in each major compare themselves to others, and, with the additional information on their family's class history, we could also gather insight on which fields individuals from different social classes might pursue in higher education.
+It will also be interesting to investigate how people across different income ranges and classes experience regret about their decisions.
 
@@
 
-#### Incorporate Additional Data Sources
-
-Additional data sources can be used to explore additional trends in attitudes towards financial wellness.
+### Incorporate Additional Data Sources
+
+Additional data sources can be used to explore additional trends in attitudes towards financial well-being.
 First, we can connecting our results with US census data to check for representativeness in respondents answers, and to understand any discrepancies in self-reported data.
@@
 This paper used data from the US General Social Survey to examine how people perceive their financial situations in comparison to others, particularly in light of the increasing wealth gap between high and low earning individuals.
-We found that significant trends in financial satisfaction and comparison based on demographic categories such as generational cohort, level of education, age, gender, and social class.The results support the hypothesis that the rich are getting richer (or at least, feel like it), the poor are getting poorer (or at least, feel like it), and that younger generations are financially worse off than preceding cohorts.
+We found that significant trends in financial satisfaction and comparison based on demographic categories such as generational cohort, level of education, age, gender, and social class.The results support the hypothesis that the rich are getting richer, the poor are getting poorer, and that younger generations are financially worse off than preceding cohorts.
 The data also suggests that people's reference groups influence their understanding of their financial well-being and satisfaction, with upper-classes experiencing counter-intuitively significant rates of financial dissatisfaction.

### 2023-03-15 outputs/paper/financial_satsifaction_in_us.qmd
table format change

@@
 | Variable | New Name               | Description                                                                             | Example               |
-|-----------------|-----------------|---------------------|-----------------|
+|--------|---------------|---------------------|---------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |

### 2023-03-15 outputs/paper/financial_satisfaction_in_us.qmd

--- 2023-03-15 outputs/paper/financial_satsifaction_in_us.qmd
@@
 | Variable | New Name               | Description                                                                             | Example               |
-|--------|---------------|---------------------|---------------|
+|----------------|----------------|-----------------------|----------------|
 | satfin   | financial_satisfaction | Respondent's satisfaction with their current financial situation                        | Pretty Well Satisfied |
